### 2012-06-14 Paper/Paper.tex
first commit

\documentclass[runningheads,a4paper]{llncs}

\usepackage{amssymb}
\setcounter{tocdepth}{3}
\usepackage{graphicx}

\usepackage{url}
\urldef{\mailcf}\path|{berlin,borazio}@ess.informatik.tu-darmstadt.de, tobias.grosse-puppendahl@igd.fraunhofer.de|
\newcommand{\keywords}[1]{\par\addvspace\baselineskip
\noindent\keywordname\enspace\ignorespaces#1}

\begin{document}

\mainmatter  % start of an individual contribution

% first the title is needed
\title{Enhancing accelerometer-based activity recognition with capacitive proximity sensing}

% a short form should be given in case it is too long for the running head
\titlerunning{Enhancing accelerometer-based activity recognition}

% the name(s) of the author(s) follow(s) next
\author{Eugen Berlin, Marko Borazio, Tobias Grosse-Puppendahl}
%
\authorrunning{Enhancing accelerometer-based activity recognition}
% (feature abused for this document to repeat the title also on left hand pages)

% the affiliations are given next; don't give your e-mail address
% unless you accept that it will be published
\institute{TU Darmstadt, Karolinenplatz 5, 64289 Darmstadt, Germany\\
Fraunhofer IGD, Fraunhoferstr. 5, 64283 Darmstadt, Germany\\
\mailcf\\
}

\toctitle{Enhancing accelerometer-based activity recognition with capacitive proximity sensing}

\tocauthor{Eugen Berlin, Marko Borazio, Tobias Grosse-Puppendahl}
\maketitle
\begin{abstract}
Zwei flinke Boxer jagen die quirlige Eva und ihren Mops durch Sylt. Franz jagt im komplett verwahrlosten Taxi quer durch Bayern. Zwölf Boxkämpfer jagen Viktor quer über den großen Sylter Deich. Vogel Quax zwickt Johnys Pferd Bim. Sylvia wagt quick den Jux bei Pforzheim. Polyfon zwitschernd aßen Mäxchens Vögel Rüben, Joghurt und Quark. "Fix, Schwyz! " quäkt Jürgen blöd vom Paß. Victor jagt zwölf Boxkämpfer quer über den großen Sylter Deich. Falsches Üben von Xylophonmusik quält jeden größeren Zwerg. 
\keywords{AAL, capacitive proximity sensors, classification, user context}
\end{abstract}

\section{Introduction}
Current wearable activity recognizing systems are able to unobtrusively recognize a person’s activities throughout the whole day. These systems often apply accelerometer-based data recordings to identify the activities performed. However, it is a great challenge to identify fine-grained activities just by using this single modality. In this paper, we use an accelerometer-based activity recognition system, the wrist-worn HedgeHog device \cite{HedgeHog}. We investigate the possibility of integrating a capacitive proximity sensor unobtrusively into the wristband to obtain higher recognition rates for several activities in daily life.
Capacitive proximity sensors can indirectly measure the distance and nature of a grounded object within reach. This means that the measurement result depends on the object’s distance, its size and the material it is made of. A wristband worn capacitive proximity sensor requires a shield that eliminates the influence of the grounded arm directly underneath the sensor. Using this setup, we can detect the proximity to a grounded object in the environments for distances up to 20 cm. Especially for mobile devices, it is required that the sensor only draws a very small amount of power. Thus, other proximity sensing input modalities like ultrasound or optical measurements are not applicable for this type of mobile application.

[MB; EB: Introduction to HedgeHog device]

There are several intuitive examples for which a combination of accelerometer-based activity recognition with a capacitive proximity sensor reveals its strength. For example, it may be conducted which material is placed underneath the hand. The capacitive proximity sensor would return a different measurement result for a hand placed on a couch covered with fabric than a hand placed a wooden table. Moreover, the approximate distance to objects can be exploited to identify activities like grasping into a locker or a refridgerator to prepare food.

\section{Related Work}

[MB; EB: Related Work activity recognition]

In the field of capacitive proximity sensing, three different measurement modes shown in Figure \ref{fig:modes} were identified by Smith et al. \cite{Smith1999}: transmit mode, shunt mode and loading mode. Transmit mode is based on a varying electric potential coupled to an object that can be measured by a capacitive proximity sensor next to that object. Shunt mode applies two electrodes, a transmit and a receive electrode, that can measure capacitance changes produced by objects disturbing the electric field between the two electrodes. In loading mode, a single electrode builds up an electric field to any grounded object in the environment. By measuring the capacitance, conclusions can be made upon the proximity and nature of an object. In our work we apply loading mode since it requires only a single electrode that can be integrated invisibly into the wristband.

A great variety of capacitive sensors and measurement techniques exists \cite{Smith1999}. The most common sensing principle, the loading mode, is based on running numerous charge and discharge cycles of the virtual capacitor that is created by the electrode and the environment. Depending on the charge and discharge times, one can infer the corresponding capacitance. This sensing principle is applied by Wimmer et al. in [wimmer2006] who presented a toolkit for capacitive proximity sensing.

In previous works, capacitive proximity sensors were applied in various fields of human-computer-interaction. For example, Wimmer et al. and Grosse-Puppendahl et al. presented gesture recognition systems \cite{Wimmer,Grosse-puppendahl2012} as well as smart furniture that can sense and classify human postures \cite{Wimmer,Grosse-puppendahl2011}. Cheng et al. have investigated the possibility of using capacitive sensors for activity recognition by measuring shape changes of muscles and skin \cite{Cheng2010}. To our knowledge, capacitive proximity sensors have not been embedded into a wearable device to enhance the performance of activity recognition.

\section{Hardware}
\subsection{HedgeHog}
The HedgeHog sensor is a custom designed and built wearable data logger aiming at long-term deployments in activity recognition scenarios. Due to its small size (30x25x20mm) and weight, this sensor, being worn at the wrist, is an unobtrusive way to record relevant motion data.

The HedgeHog sensor is built around the low-power Microchip microcontroller (PIC18F46J50) featuring an accelerometer sensor (ADXL345) to capture human motion, light and ambient temperature sensors and a microSD flash card for locally storing the sensor data. The sensor is powered by a $200mAh$ lithium polymer battery, which allows two weeks of continuous recording on a single battery charge. A plastic case is used to nicely package and protect the sensor to be worn at the wrist.

\includegraphics[width=.75\textwidth]{sensor.png}
The accelerometer sensor is being sampled at $100Hz$, resulting in 10ms equidistant measurements. For efficiency reasons, the sensor data is being run length encoded before being stored locally.

The HedgeHog can be extended with further sensors tailoring different application scenarios. For our scenario, we have added a capacitive proximity sensor, which is described in detail in the next section.

\subsection{Capacitive Proximity Sensor}
power drain: 1mA when active
image : photo
image : board layout and schematics !

The capacitive proximity sensor for the wristband performs measurements in loading mode. The sensing circuit schematic is shown in Figure \ref{}. It is based on a timer that controls the charging and discharging cycles of the virtual capacitor. The timer toggles from charge to discharge at the time when a threshold voltage at the capacitor is reached. This results in an astable operation with succeeding charge/discharge cycles. When the capacitance of the virtual capacitor increases, the charging time will also increase and vice-versa. Therefore, the capacitance is inversely proportional to the number of charging cycles in a given time span. In order to guard the sensor from measuring the capacitance to underlying objects, a shield is placed directly underneath the measuring electrode. The shield is driven with the same potential as the sensing electrode, such that the the capacitance between the two electrodes is neglectable. Using this shielding method, the measured capacitance will only be slightly affected by the grounded underlying arm.

The proximity sensor board generates a clock signal with varying frequency depending on the proximity to body parts, objects and their characteristics. The HedgeHog therefore needs to measure the incoming frequency, which is implemented by counting the signal’s edges over the gate time of approximately 9.5ms.

pseudocode:
    sample the accelerometer sensor

start frequency counter

go to sleep for 9.5 ms

wake-up (timer interrupt)

stop frequency counter

read frequency count

\section{Experiment}

\subsection{Setup and Scenario}

The expirement setup aims to depict a typical scenario in daily life. Especially in the field of Ambient Assisted Living (AAL), it is desired to monitor activities like drinking, preparing lunch and sleeping. A fine-grained monitoring of such activities may help elderly or people suffering from mental diseases to maintain a healthy day/night rhythm and take action if irregularities occur.

The recorded test set contains the following activities: opening door, sitting on a couch, lying on a couch, grasping into a locker, putting something from a shelf, sitting at the table, preparing a bread with marmalade, eating bread, mould water, drinking water and sleeping. Some of those activities are very hard to recognize without context information, just using wrist-worn accelerometer data. For example, sitting at the table and sitting on a couch are very similar activities. The data of a capacitive proximity sensor can extend this data basis by providing that is related to the material an object is made of.

\begin{figure}[htbp]
	\centering
		\includegraphics[trim=2cm 0.5cm 2cm 0.5cm,width=1.00\textwidth]{Images/drinking.pdf}
	\caption{Drinking activity}
	\label{fig:drinking}
\end{figure}

\begin{figure}[htbp]
	\centering
		\includegraphics[trim=2cm 0.5cm 2cm 0.5cm,width=1.00\textwidth]{Images/making-bread.pdf}
	\caption{Preparing bread}
	\label{fig:making-bread}
\end{figure}

[EB: Hier plot von drinking und grasping into locker]

Figure \ref{fig:drinking} depicts the data that was recorded for the activity drinking water. The accelerometer data shows that there periodic up- and down-movements while the capacitive proximity sensor delivers data that is associated to the proximity of the table. The second Figure \ref{fig:opendoor} illustrates a grasp into a locker while preparing lunch. It is obvious that the locker’s bottom surface that is located very closely to the wristband has an influence on the proximity sensor.

In order to evaluate if capacitive proximity sensors in wrist-bands can enhance the performance of activity recognition, we have conducted an evaluation with ten test persons. All test persons received a plot with the activities they were supposed to perform. After manual labeling, we used this test-set as ground truth and performed a cross-validation on an support-vector machine (SVM) classifier once with and once without including the data of the capacitive proximity sensor into the feature set.

[EB MB: kurze Einführung in SVMs]

The classifier was trained with various features that were extracted from windows with the length of 5 seconds. In order to suppress noise contained in the capacitive proximity sensing data, we applied a Gaussian kernel with a size of [TODO] samples. The final feature set contained the arithmetic mean and standard variance of both accelerometer and proximity data. Moreover, we included the median, min and max features for both modalities.

\subsection{Evaluation Results}

\section{Conclusion and Outlook}

\section*{Acknowledgments}
We would like to thank the students of TU Darmstadt and the employees of Fraunhofer IGD that lent us their sitting and lying abilities for the evaluation.

\newpage

\bibliographystyle{splncs03}
\bibliography{references}

\end{document}

### 2012-06-14 Auswertung/plot_results.m

@@ -1,0 +1,86 @@
+%recordings = [csvread('feugen.csv')];
+%classes = [csvread('ae.csv')];
+recordings = nagi;
+classes = ann_nagi;
+
+sum_acc = sum(recordings(:,2));
+num_recordings = size(recordings);
+num_recordings = num_recordings(1);
+
+recordings_new = zeros(sum_acc, 7);
+classes_new = zeros(sum_acc, 1);
+
+% expand recordings by acc
+k = 1;
+for i = 1:num_recordings
+    for j = 1:recordings(i, 2)
+       recordings_new(k, :) = recordings(i, :);
+       classes_new(k, :) = classes(i, :);
+       k = k + 1;
+    end
+end
+
+recordings = recordings_new;
+time = [0:1:sum_acc-1] ./ 100;
+time = time';
+    
+% walking
+%start = 55370;
+%stop = 62320;
+
+% making bread
+%start = 39837;
+%stop = 42168;
+
+% drink
+%start = 50550;
+%stop = 60890;
+
+% all
+start = + 10000;
+stop = sum_acc - 10000;
+
+delta = stop - start;
+window_size = 20;
+
+acceleration_x = recordings(:,3);
+acceleration_y = recordings(:,4);
+acceleration_z = recordings(:,5);
+
+cap = recordings(:,6);
+
+% smooth all recorded data by average filter
+average_kernel = ones(100, 1) * 0.0125;
+%cap = conv(cap, average_kernel, 'same');
+%acceleration_x = conv(acceleration_x, average_kernel, 'same');
+%acceleration_y = conv(acceleration_y, average_kernel, 'same');
+%acceleration_z = conv(acceleration_z, average_kernel, 'same');
+
+% resulting values
+res_cap = cap(start:stop);
+res_acceleration_x = acceleration_x(start:stop);
+res_acceleration_y = acceleration_y(start:stop);
+res_acceleration_z = acceleration_z(start:stop);
+% res_time = time(start:stop); % for continuos time
+res_time = ((0:(delta)) * 0.01)';
+
+subplot(2,1,1);
+set(gca,'LooseInset',get(gca,'TightInset'));
+set (gca, 'box', 'on');
+plot(res_time, res_cap, '-m', 'LineWidth', 1);
+axis([res_time(1) res_time(delta) (min(res_cap)-1)  (max(res_cap)+1)]);
+ylabel('frequency count');
[…]

### 2012-06-14 Auswertung/run_svm.m

@@ -1,0 +1,178 @@
+features_file_no_cap = 'features_no_cap.arff';
+features_file_with_cap = 'features_with_cap.arff';
+
+%classes = [csvread('ae.csv'); csvread('am.csv')];
+%recordings = [csvread('feugen.csv'); csvread('fmarko.csv')];
+classes = ann_eugen;
+recordings = eugen;
+
+window_size = 100;
+window_overlap = 50;
+number_features = 8;
+
+number_recordings = size(recordings);
+number_recordings = number_recordings(1);
+extracted_windows = 0; 
+i = 1;
+
+clear extracted_features;
+
+while i < (number_recordings)
+    % this determines the window's size
+    current_class = classes(i);
+    extract_features = true;
+   
+    % determine window size by acc
+    end_of_window = i;
+    sum_acc = 0;
+    while sum_acc < window_size
+        end_of_window = end_of_window + 1;
+        
+        if end_of_window > number_recordings
+           break; 
+        end
+        
+        sum_acc = sum_acc + recordings(end_of_window, 2);
+    end
+    
+    % this is a fixed window size, alternative to acc
+    % end_of_window = i + window_size - 1;
+    
+    % determine wether the window is valid
+    for j = i:end_of_window
+        % if j is out of bounds, stop it
+        if j > number_recordings
+            extract_features = false;
+            break;
+        end
+        
+        compare_class = classes(j);
+        
+        %if the class changes, do not extract features
+        if current_class ~= compare_class
+            extract_features = false;
+            break;
+        end
+    end
+    
+    % if the window is ok, do the feature extraction
+    if extract_features == true && current_class ~= 0
+        accel_1 = recordings(i:end_of_window, 3);
+        accel_2 = recordings(i:end_of_window, 4);
+        accel_3 = recordings(i:end_of_window, 5);
+        cap = recordings(i:end_of_window, 6);
+        
+        results = zeros(number_features,1);
+        results(1) = mean(accel_1);
+        results(2) = mean(accel_2);
+        results(3) = mean(accel_3);
+        results(4) = var(accel_1);
+        results(5) = var(accel_2);
+        results(6) = var(accel_3);
+        results(7) = mean(cap);
+        results(8) = var(cap);
+        results(9) = current_class;
+        
+        extracted_windows = extracted_windows + 1;
+        
[…]

### 2012-06-14 Auswertung/write_features.m

@@ -1,0 +1,149 @@
+features_file_no_cap = 'features_no_cap.arff';
+features_file_with_cap = 'features_with_cap.arff';
+
+%classes = [csvread('ae.csv'); csvread('am.csv')];
+%recordings = [csvread('feugen.csv'); csvread('fmarko.csv')];
+classes = ann_enzo;
+recordings = enzo;
+
+window_size = 100;
+window_overlap = 50;
+number_features = 8;
+
+number_recordings = size(recordings);
+number_recordings = number_recordings(1);
+extracted_windows = 0; 
+i = 1;
+
+clear extracted_features;
+
+while i < (number_recordings)
+    % this determines the window's size
+    current_class = classes(i);
+    extract_features = true;
+   
+    % determine window size by acc
+    end_of_window = i;
+    sum_acc = 0;
+    while sum_acc < window_size
+        end_of_window = end_of_window + 1;
+        
+        if end_of_window > number_recordings
+           break; 
+        end
+        
+        sum_acc = sum_acc + recordings(end_of_window, 2);
+    end
+    
+    % this is a fixed window size, alternative to acc
+    % end_of_window = i + window_size - 1;
+    
+    % determine wether the window is valid
+    for j = i:end_of_window
+        % if j is out of bounds, stop it
+        if j > number_recordings
+            extract_features = false;
+            break;
+        end
+        
+        compare_class = classes(j);
+        
+        %if the class changes, do not extract features
+        if current_class ~= compare_class
+            extract_features = false;
+            break;
+        end
+    end
+    
+    % if the window is ok, do the feature extraction
+    if extract_features == true && current_class ~= 0
+        accel_1 = recordings(i:end_of_window, 3);
+        accel_2 = recordings(i:end_of_window, 4);
+        accel_3 = recordings(i:end_of_window, 5);
+        cap = recordings(i:end_of_window, 6);
+        
+        results = zeros(number_features,1);
+        results(1) = mean(accel_1);
+        results(2) = mean(accel_2);
[…]

### 2012-06-14 SVM/SVM-KMExample/pathdef.m

@@ -1,0 +1,338 @@
+function p = pathdef
+%PATHDEF Search path defaults.
+%   PATHDEF returns a string that can be used as input to MATLABPATH
+%   in order to set the path.
+
+  
+%   Copyright 1984-2007 The MathWorks, Inc.
+%   $Revision: 1.4.2.2 $ $Date: 2007/06/07 14:45:14 $
+
+
+% DO NOT MODIFY THIS FILE.  IT IS AN AUTOGENERATED FILE.  
+% EDITING MAY CAUSE THE FILE TO BECOME UNREADABLE TO 
+% THE PATHTOOL AND THE INSTALLER.
+
+p = [...
+%%% BEGIN ENTRIES %%%
+     '/media/nas_ess/projects/Ubicomp_2011/twitches/SVM-KMExample:', ...
+     '/media/nas_ess/projects/Ubicomp_2011/twitches/SVM-KM:', ...
+     '/media/nas_ess/projects/Ubicomp_2011/twitches/SVM-KM/AdaptScalSVM:', ...
+     '/media/nas_ess/projects/Ubicomp_2011/twitches/SVM-KM/FeatureSelection:', ...
+     '/media/nas_ess/projects/Ubicomp_2011/twitches/SVM-KM/featselreg:', ...
+     '/media/nas_ess/projects/Ubicomp_2011/twitches/SVM-KM/kbp:', ...
+     '/media/nas_ess/projects/Ubicomp_2011/twitches/SVM-KM/libsvminterface:', ...
+     '/media/nas_ess/projects/Ubicomp_2011/twitches/SVM-KM/regpath:', ...
+     '/media/nas_ess/projects/Ubicomp_2011/twitches/SVM-KM/svmroc:', ...
+     '/media/nas_ess/projects/Ubicomp_2011/twitches/SVM-KM/utilities:', ...
+     '/media/nas_ess/projects/Ubicomp_2011/twitches/SVM-KM/waveletkernelmultires:', ...
+     '/media/nas_ess/projects/Ubicomp_2011/night/somtoolbox:', ...
+     '/media/nas_ess/projects/dataset_converter/v0.46:', ...
+     matlabroot,'/toolbox/HMM:', ...
+     '/home/marko/HMMall:', ...
+     matlabroot,'/toolbox/videoio:', ...
+     matlabroot,'/toolbox/mmwrite:', ...
+     matlabroot,'/toolbox/mmreader:', ...
[…]

### 2012-06-15 Paper/Paper.tex
some changes
-references
-hardware figure
-title,institute and some text work

@@
 \usepackage{graphicx}
+\usepackage{booktabs}
 
 \usepackage{url}
-\urldef{\mailcf}\path|{berlin,borazio}@ess.informatik.tu-darmstadt.de, tobias.grosse-puppendahl@igd.fraunhofer.de|
-\newcommand{\keywords}[1]{\par\addvspace\baselineskip
-\noindent\keywordname\enspace\ignorespaces#1}
+% \urldef{\mailcf}\path|{berlin,borazio}@ess.informatik.tu-darmstadt.de, tobias.grosse-puppendahl@igd.fraunhofer.de|
+% \newcommand{\keywords}[1]{\par\addvspace\baselineskip
+% \noindent\keywordname\enspace\ignorespaces#1}
 
@@
 % first the title is needed
-\title{Enhancing accelerometer-based activity recognition with capacitive proximity sensing}
+\title{Enhancing Accelerometer-based Activity Recognition with Capacitive Proximity Sensing}
 
 % a short form should be given in case it is too long for the running head
-\titlerunning{Enhancing accelerometer-based activity recognition}
+\titlerunning{Enhancing Accelerometer-based Activity Recognition}
 
 % the name(s) of the author(s) follow(s) next
-\author{Eugen Berlin, Marko Borazio, Tobias Grosse-Puppendahl}
+\author{Tobias Grosse-Puppendahl\inst{1} \and Eugen Berlin\inst{2} \and Marko Borazio\inst{2} }
 %
-\authorrunning{Enhancing accelerometer-based activity recognition}
+\authorrunning{Enhancing Accelerometer-based Activity Recognition}
 % (feature abused for this document to repeat the title also on left hand pages)
@@
 % unless you accept that it will be published
-\institute{TU Darmstadt, Karolinenplatz 5, 64289 Darmstadt, Germany\\
+\institute{
 	Fraunhofer IGD, Fraunhoferstr. 5, 64283 Darmstadt, Germany\\
-\mailcf\\
+	\email{tobias.grosse-puppendahl@igd.fraunhofer.de}
+	\and
+	Technische Universit\"at Darmstadt, Hochschulstr. 10, 64289 Darmstadt, Germany\\
+	\email{\{berlin,borazio\}@ess.tu-darmstadt.de}
 }
 
-\toctitle{Enhancing accelerometer-based activity recognition with capacitive proximity sensing}
+% \toctitle{Enhancing Accelerometer-based Activity Recognition with Capacitive Proximity Sensing}
+% \tocauthor{Eugen Berlin, Marko Borazio, Tobias Grosse-Puppendahl}
 
-\tocauthor{Eugen Berlin, Marko Borazio, Tobias Grosse-Puppendahl}
 \maketitle
@@
 Falsches Üben von Xylophonmusik quält jeden größeren Zwerg.
-\keywords{AAL, capacitive proximity sensors, classification, user context}
+\keywords{ambient assisted living, capacitive proximity sensors, activity recognition, user context}
 \end{abstract}
@@
 \section{Introduction}
-Current wearable activity recognizing systems are able to unobtrusively recognize a person’s activities throughout the whole day.
-These systems often apply accelerometer-based data recordings to identify the activities performed.
+Current wearable activity recognition systems are able to unobtrusively capture and recognize a person's activities throughout the whole day.
+These systems often rely on inertial sensor data that is captured by wearable sensors attached to the body to identify the activities performed.
 However, it is a great challenge to identify fine-grained activities just by using this single modality.
-In this paper, we use an accelerometer-based activity recognition system, the wrist-worn HedgeHog device \cite{HedgeHog}.
-We investigate the possibility of integrating a capacitive proximity sensor unobtrusively into the wristband to obtain higher recognition rates for several activities in daily life.
+In this paper, we use an accelerometer-based activity recognition system, the wrist-worn activity logger called HedgeHog\footnote{\url{http://www.ess.tu-darmstadt.de/hedgehog}} and investigate the possibility of integrating a capacitive proximity sensor unobtrusively into the wristband to obtain higher recognition rates for several activities in daily life.
+
 Capacitive proximity sensors can indirectly measure the distance and nature of a grounded object within reach.
-This means that the measurement result depends on the object’s distance, its size and the material it is made of.
+This means that the measurement result depends on the object's distance, its size and the material it is made of.
 A wristband worn capacitive proximity sensor requires a shield that eliminates the influence of the grounded arm directly underneath the sensor.
@@
 The capacitive proximity sensor would return a different measurement result for a hand placed on a couch covered with fabric than a hand placed a wooden table.
-Moreover, the approximate distance to objects can be exploited to identify activities like grasping into a locker or a refridgerator to prepare food.
+Moreover, the approximate distance to objects can be exploited to identify activities like grasping into a locker or a refrigerator  to prepare food.
 
@@
 [MB; EB: Related Work activity recognition]
+
+Berlin et al.
+\cite{berlin_laerhoven_tei_2010} have augmented an accelerometer based sensor node with an RFID bracelet to be able to detect objects a human is interacting with.
+Thus, activities can use this additional modality
 
@@
 The HedgeHog sensor is a custom designed and built wearable data logger aiming at long-term deployments in activity recognition scenarios.
-Due to its small size (30x25x20mm) and weight, this sensor, being worn at the wrist, is an unobtrusive way to record relevant motion data.
+Due to its small form-factor (30x25x20mm) and weight, this sensor, being worn at the wrist, is an unobtrusive way to record relevant motion data.
 
 The HedgeHog sensor is built around the low-power Microchip microcontroller (PIC18F46J50) featuring an accelerometer sensor (ADXL345) to capture human motion, light and ambient temperature sensors and a microSD flash card for locally storing the sensor data.
-The sensor is powered by a $200mAh$ lithium polymer battery, which allows two weeks of continuous recording on a single battery charge.
-A plastic case is used to nicely package and protect the sensor to be worn at the wrist.
+The sensor is powered by a $200mAh$ lithium polymer battery, which allows for two weeks of continuous recording on a single battery charge.
+A plastic case is used to nicely package and protect the sensor to be worn at the wrist (as shown in Fig.
+\ref{fig:sensornode}).
 
-\includegraphics[width=.75\textwidth]{sensor.png}
-The accelerometer sensor is being sampled at $100Hz$, resulting in 10ms equidistant measurements. For efficiency reasons, the sensor data is being run length encoded before being stored locally.
+\begin{figure}
+	\centering
+	\includegraphics[width=\textwidth]{Images/hardware_sensor.jpg}
+	\caption{The inertial data logger featuring a low-power microcontroller, a 3 axis accelerometer, a microSD flash card for storing the sensor data and a USB connector for accessing the stored data (on the left) is being powered by a small lithium polymer battery and is packaged into a plastic case to be worn at the wrist (a version with a OLED display on the right).}
+	\label{fig:sensornode}
+\end{figure}
+
+The 3D accelerometer sensor is being sampled at $100Hz$, resulting in 10ms equidistant measurements.
+For efficiency reasons, the sensor data is run-length encoded before being stored locally to the microSD card.
 
@@
 The proximity sensor board generates a clock signal with varying frequency depending on the proximity to body parts, objects and their characteristics.
-The HedgeHog therefore needs to measure the incoming frequency, which is implemented by counting the signal’s edges over the gate time of approximately 9.5ms.
+The HedgeHog therefore needs to measure the incoming frequency, which is implemented by counting the signal's edges over the gate time of approximately 9.5ms.
 
@@
 
-The expirement setup aims to depict a typical scenario in daily life.
+The experiment setup aims to depict a typical scenario in daily life.
 Especially in the field of Ambient Assisted Living (AAL), it is desired to monitor activities like drinking, preparing lunch and sleeping.
@@
 
-The recorded test set contains the following activities: opening door, sitting on a couch, lying on a couch, grasping into a locker, putting something from a shelf, sitting at the table, preparing a bread with marmalade, eating bread, mould water, drinking water and sleeping.
+The recorded test set contains the following activities: opening door, sitting on a couch, lying on a couch, grasping into a locker, putting something from a shelf, sitting at the table, preparing a bread with marmalade, eating the bread, pouring water, drinking the water and sleeping.
 Some of those activities are very hard to recognize without context information, just using wrist-worn accelerometer data.
@@
 
-\begin{figure}[htbp]
+\begin{figure}
 	\centering
@@
 The accelerometer data shows that there periodic up- and down-movements while the capacitive proximity sensor delivers data that is associated to the proximity of the table.
-The second Figure \ref{fig:opendoor} illustrates a grasp into a locker while preparing lunch.
-It is obvious that the locker’s bottom surface that is located very closely to the wristband has an influence on the proximity sensor.
+Figure \ref{fig:opendoor} illustrates a grasp into a locker while preparing lunch.
+It is obvious that the locker's bottom surface that is located very closely to the wristband has an influence on the proximity sensor.
 
@@
 \section*{Acknowledgments}
-We would like to thank the students of TU Darmstadt and the employees of Fraunhofer IGD that lent us their sitting and lying abilities for the evaluation.
+% We would like to thank the students of TU Darmstadt and the employees of Fraunhofer IGD that lent us their sitting and lying abilities for the evaluation. 
+We would like to thank our evaluation participants from the TU Darmstadt and the Fraunhofer IGD.
 

### 2012-06-15 Paper/Paper.tex
added some drawings

@@
 
-\section{Related Work}
+\section{Related work}
 
@@
 Thus, activities can use this additional modality
+
+\begin{figure}[htbp]
+	\centering
+		\includegraphics[width=1.00\textwidth]{Images/modes.png}
+	\caption{Capacitive proximity sensing modes \cite{Smith1996}}
+	\label{fig:modes}
+\end{figure}
 
@@
 and Grosse-Puppendahl et al.
-presented gesture recognition systems \cite{Wimmer,Grosse-puppendahl2012} as well as smart furniture that can sense and classify human postures \cite{Wimmer,Grosse-puppendahl2011}.
+presented gesture recognition systems \cite{Wimmer,Grosse-puppendahl2012} as well as smart furniture that can sense human activities \cite{Wimmer} and classify human postures \cite{Grosse-puppendahl2011}.
 Cheng et al.
@@
 For efficiency reasons, the sensor data is run-length encoded before being stored locally to the microSD card.
-
 The HedgeHog can be extended with further sensors tailoring different application scenarios.
@@
 
-\subsection{Capacitive Proximity Sensor}
+\subsection{Capacitive proximity sensor}
 power drain: 1mA when active
@@
 The sensing circuit schematic is shown in Figure \ref{}.
-It is based on a timer that controls the charging and discharging cycles of the virtual capacitor.
+It is based on a timer that controls the charging and discharging cycles of the virtual capacitor that is built by the wristband electrode and the surrounding environment.
 The timer toggles from charge to discharge at the time when a threshold voltage at the capacitor is reached.
@@
 
-The proximity sensor board generates a clock signal with varying frequency depending on the proximity to body parts, objects and their characteristics.
-The HedgeHog therefore needs to measure the incoming frequency, which is implemented by counting the signal's edges over the gate time of approximately 9.5ms.
+The proximity sensor board generates a clock signal with varying frequency depending on the charge and discharge cycles.
+The HedgeHog measures the resulting capacitance by counting the signal's edges over a gate time of approximately 9.5ms.
 
-pseudocode:
-    sample the accelerometer sensor
-
-start frequency counter
-
-go to sleep for 9.5 ms
-
-wake-up (timer interrupt)
-
-stop frequency counter
-
-read frequency count
-
+\begin{figure}[htbp]
+	\centering
+		\includegraphics[trim=1cm 1cm 1cm 1cm,width=1.00\textwidth]{Images/pseudocode.pdf}
+	\caption{Measurement procedure}
+	\label{fig:pseudocode}
+\end{figure}
 \section{Experiment}
@@
 The recorded test set contains the following activities: opening door, sitting on a couch, lying on a couch, grasping into a locker, putting something from a shelf, sitting at the table, preparing a bread with marmalade, eating the bread, pouring water, drinking the water and sleeping.
-Some of those activities are very hard to recognize without context information, just using wrist-worn accelerometer data.
+Some of those activities are very hard to recognize without any further contextual information, when the data is limited to a single modality like a 3D accelerometer.
 For example, sitting at the table and sitting on a couch are very similar activities.

### 2012-06-15 Paper/Paper.tex
more changes - captions - some text improvements

--- 2012-06-15 Paper/Paper.tex
@@
 \titlerunning{Enhancing Accelerometer-based Activity Recognition}
+% (feature abused for this document to repeat the title also on left hand pages)
+\authorrunning{Enhancing Accelerometer-based Activity Recognition}
 
 % the name(s) of the author(s) follow(s) next
-\author{Tobias Grosse-Puppendahl\inst{1} \and Eugen Berlin\inst{2} \and Marko Borazio\inst{2} }
-%
-\authorrunning{Enhancing Accelerometer-based Activity Recognition}
-% (feature abused for this document to repeat the title also on left hand pages)
+% \author{Tobias Grosse-Puppendahl\inst{1} \and Eugen Berlin\inst{2} \and Marko Borazio\inst{2} }
+\author{Author 1 \and Author 2 \and Author 3 }
 
@@
 % unless you accept that it will be published
+% \institute{
+% 	Fraunhofer IGD, Fraunhoferstr. 5, 64283 Darmstadt, Germany\\
+% 	\email{tobias.grosse-puppendahl@igd.fraunhofer.de}
+% 	\and
+% 	Technische Universit\"at Darmstadt, Hochschulstr. 10, 64289 Darmstadt, Germany\\
+% 	\email{\{berlin,borazio\}@ess.tu-darmstadt.de}
+% }
+
 \institute{
-	Fraunhofer IGD, Fraunhoferstr. 5, 64283 Darmstadt, Germany\\
-	\email{tobias.grosse-puppendahl@igd.fraunhofer.de}
+	Anonymized for review \\
+	\email{ananymous@insitution.edu}
 	\and
-	Technische Universit\"at Darmstadt, Hochschulstr. 10, 64289 Darmstadt, Germany\\
-	\email{\{berlin,borazio\}@ess.tu-darmstadt.de}
+	Anonymized for review \\
+	\email{ananymous@insitution.edu}
 }
@@
 However, it is a great challenge to identify fine-grained activities just by using this single modality.
-In this paper, we use an accelerometer-based activity recognition system, the wrist-worn activity logger called HedgeHog\footnote{\url{http://www.ess.tu-darmstadt.de/hedgehog}} and investigate the possibility of integrating a capacitive proximity sensor unobtrusively into the wristband to obtain higher recognition rates for several activities in daily life.
+In this paper, we have obtained and use a wrist-worn activity data logger\footnote{HedgeHog Activity Logger -- \url{http://www.ess.tu-darmstadt.de/hedgehog}} and investigate the possibility of integrating a capacitive proximity sensor unobtrusively into the wristband to obtain higher recognition rates for several activities in daily life.
 
@@
 
-Berlin et al.
-\cite{berlin_laerhoven_tei_2010} have augmented an accelerometer based sensor node with an RFID bracelet to be able to detect objects a human is interacting with.
-Thus, activities can use this additional modality
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+The field of activity recognition relying on wearable sensors mostly mostly considers inertial data from the participants body to infer performed activities, such as in the works of [bla,foo].
+The acceleration data is often augmented with data from sensors such as gyroscopes, magnetometers, ambient light or temperature, aiming at extracting more motion and environmental context of the user.
+[references?!]
+
+The works of \cite{intel_bracelet} and ...
+show that detecting objects the user touches or uses throughout the day can be very helpful for the activity recognition, including activities of daily living (maja(?)), maintenance tasks (?), gardening activities, etc.
+For example, Berlin et al.
+\cite{berlin_laerhoven_tei_2010} have augmented a wrist-worn accelerometer sensor node with an RFID bracelet to be able to detect objects a human is interacting with.
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 
@@
 This sensing principle is applied by Wimmer et al.
-in [wimmer2006] who presented a toolkit for capacitive proximity sensing.
+in \cite{wimmer2006} who presented a toolkit for capacitive proximity sensing.
 
@@
 \section{Hardware}
-\subsection{HedgeHog}
+This section presents the two components of the required hardware, the wrist-worn activity data logger tailored to capture acceleration data, and the capacitive proximity sensor used for distance measurements.
+
+\subsection{Activity Data Logger}
 The HedgeHog sensor is a custom designed and built wearable data logger aiming at long-term deployments in activity recognition scenarios.
@@
 
-The HedgeHog sensor is built around the low-power Microchip microcontroller (PIC18F46J50) featuring an accelerometer sensor (ADXL345) to capture human motion, light and ambient temperature sensors and a microSD flash card for locally storing the sensor data.
+The sensor node itself is built around the low-power Microchip microcontroller (PIC18F46J50) featuring an accelerometer sensor (ADXL345) to capture human motion, light and ambient temperature sensors and a microSD flash card for locally storing the sensor data.
 The sensor is powered by a $200mAh$ lithium polymer battery, which allows for two weeks of continuous recording on a single battery charge.
-A plastic case is used to nicely package and protect the sensor to be worn at the wrist (as shown in Fig.
+A USB port is used to configure the sensor (e.g.
+setting the sensitivity of the accelerometer), to access the stored sensor data, and to recharge the battery.
+A plastic case nicely packages and protects the sensor to be worn at the wrist (as shown in Fig.
 \ref{fig:sensornode}).
@@
 
-The HedgeHog can be extended with further sensors tailoring different application scenarios.
+Since the main microcontroller on the sensor node has unoccupied pins, it can be extended with further sensors tailoring different application scenarios.
 For our scenario, we have added a capacitive proximity sensor, which is described in detail in the next section.
@@
 
+\begin{verbatim}
 pseudocode:
   sample the accelerometer sensor
-
   start frequency counter
-
   go to sleep for 9.5 ms
-
   wake-up (timer interrupt)
-
   stop frequency counter
-
   read frequency count
+\end{verbatim}
 
 \section{Experiment}
+
+This section presents the experimental setup including the activities and the participants, as well as the findings that were obtained during the evaluation.
 
@@
 	\centering
-		\includegraphics[trim=2cm 0.5cm 2cm 0.5cm,width=1.00\textwidth]{Images/drinking.pdf}
-	\caption{Drinking activity}
+		\includegraphics[trim=1.9cm 0.5cm 1.8cm 0.5cm,clip,width=\textwidth]{Images/drinking.pdf}
+	\caption{An example of the ``drinking'' activity with raw 3D acceleration data (lower plot) and proximity sensor data (upper plot).
+The participant took five nips from the glass, while leaving the hands lying on the table in between.
+These motions can be easily detected in the acceleration as well as proximity data.}
 	\label{fig:drinking}
@@
 	\centering
-		\includegraphics[trim=2cm 0.5cm 2cm 0.5cm,width=1.00\textwidth]{Images/making-bread.pdf}
-	\caption{Preparing bread}
+		\includegraphics[trim=1.9cm 0.5cm 1.8cm 0.5cm,clip,width=\textwidth]{Images/making-bread.pdf}
+	\caption{An example of the ``preparing bread'' activity, where the participants had to put marmalade on a toast.
+Note, that the acceleration data does not seem to have any characteristic patterns, while the proximity sensor indicates a table, plate, or other objects in ``unmittelbarer nähe''.}
 	\label{fig:making-bread}

### 2012-06-15 Paper/Paper.tex

@@
 
+\begin{figure}[htbp]
+	\centering
+% 		\includegraphics[width=1.00\textwidth]{Images/modes.png}
+	\caption{Capacitive proximity sensing modes \cite{Smith1996}}
+	\label{fig:modes}
+\end{figure}
+
 In the field of capacitive proximity sensing, three different measurement modes shown in Figure \ref{fig:modes} were identified by Smith et al.
@@
 For efficiency reasons, the sensor data is run-length encoded before being stored locally to the microSD card.
-
-Since the main microcontroller on the sensor node has unoccupied pins, it can be extended with further sensors tailoring different application scenarios.
+The HedgeHog can be extended with further sensors tailoring different application scenarios.
 For our scenario, we have added a capacitive proximity sensor, which is described in detail in the next section.
@@
 In order to guard the sensor from measuring the capacitance to underlying objects, a shield is placed directly underneath the measuring electrode.
-The shield is driven with the same potential as the sensing electrode, such that the the capacitance between the two electrodes is neglectable.
+The shield is driven with the same potential as the sensing electrode, such that the the capacitance between the two electrodes is negligible.
 Using this shielding method, the measured capacitance will only be slightly affected by the grounded underlying arm.
 
-The proximity sensor board generates a clock signal with varying frequency depending on the proximity to body parts, objects and their characteristics.
-The HedgeHog therefore needs to measure the incoming frequency, which is implemented by counting the signal's edges over the gate time of approximately 9.5ms.
+The proximity sensor board generates a clock signal with varying frequency depending on the charge and discharge cycles.
+The HedgeHog measures the resulting capacitance by counting the signal's edges over a gate time of approximately 9.5ms.
 
-\begin{verbatim}
-pseudocode:
-  sample the accelerometer sensor
-  start frequency counter
-  go to sleep for 9.5 ms
-  wake-up (timer interrupt)
-  stop frequency counter
-  read frequency count
-\end{verbatim}
-
+\begin{figure}[htbp]
+	\centering
+% 		\includegraphics[trim=1cm 1cm 1cm 1cm,clip,width=\textwidth]{Images/pseudocode.pdf}
+	\caption{Measurement procedure}
+	\label{fig:pseudocode}
+\end{figure}
 \section{Experiment}
@@
 	\caption{An example of the ``preparing bread'' activity, where the participants had to put marmalade on a toast.
-Note, that the acceleration data does not seem to have any characteristic patterns, while the proximity sensor indicates a table, plate, or other objects in ``unmittelbarer nähe''.}
+Note, that the acceleration data does not seem to have any characteristic patterns, while the proximity sensor indicates a table, plate, or other objects in immediate distance.}
 	\label{fig:making-bread}

### 2012-06-18 Paper/Paper.tex
added images and text

@@
 	Anonymized for review \\
-	\email{ananymous@insitution.edu}
+	\email{anonymous@insitution.edu}
 	\and
 	Anonymized for review \\
-	\email{ananymous@insitution.edu}
+	\email{anonymous@insitution.edu}
 }
@@
 	\centering
-% 		\includegraphics[width=1.00\textwidth]{Images/modes.png}
-	\caption{Capacitive proximity sensing modes \cite{Smith1996}}
+		\includegraphics[width=1.00\textwidth]{Images/modes.pdf}
+	\caption{Three different capacitive proximity sensing modes can be distinguished.
+The sensing electrodes build up an electric field to objects in the environment, illustrated with a cloud.
+\cite{Smith1996}}
 	\label{fig:modes}
@@
 This sensing principle is applied by Wimmer et al.
-in \cite{wimmer2006} who presented a toolkit for capacitive proximity sensing.
+in \cite{Wimmer2007} who presented a toolkit for capacitive proximity sensing.
 
@@
 \subsection{Capacitive Proximity Sensor}
-power drain: 1mA when active
+
+The capacitive proximity sensor performs measurements in loading mode.
+Two electrodes are integrated into the wristband, one sensing electrode and one shielding electrode that aims to eliminate the influence of the arm that is a very proximate grounded object to the sensing electrode.
+The sensor draws a supply current of 1 mA at 3.3 V when active.
+This low power consumption qualifies the sensor for wearable proximity sensing applications.
+
 image : photo
-image : board layout and schematics !
 
-The capacitive proximity sensor for the wristband performs measurements in loading mode.
-The sensing circuit schematic is shown in Figure \ref{}.
-It is based on a timer that controls the charging and discharging cycles of the virtual capacitor that is built by the wristband electrode and the surrounding environment.
+\begin{figure}[htbp]
+	\centering
+		\includegraphics[width=1.00\textwidth]{Images/schematic.pdf}
+	\caption{The sensing circuit is based on a timer with an operational amplifier that acts as a voltage follower.}
+	\label{fig:schematic}
+\end{figure}
+
+The sensing circuit schematic is shown in Figure \ref{fig:schematic}.
+It is based on a timer that controls the charging and discharging cycles of the virtual capacitor that is built by the sensing electrode and the surrounding environment.
 The timer toggles from charge to discharge at the time when a threshold voltage at the capacitor is reached.
@@
 Therefore, the capacitance is inversely proportional to the number of charging cycles in a given time span.
-In order to guard the sensor from measuring the capacitance to underlying objects, a shield is placed directly underneath the measuring electrode.
+In order to guard the sensor from measuring the capacitance to underlying objects, a shield electrode is placed directly underneath the measuring electrode.
 The shield is driven with the same potential as the sensing electrode, such that the the capacitance between the two electrodes is negligible.
-Using this shielding method, the measured capacitance will only be slightly affected by the grounded underlying arm.
-
-The proximity sensor board generates a clock signal with varying frequency depending on the charge and discharge cycles.
-The HedgeHog measures the resulting capacitance by counting the signal's edges over a gate time of approximately 9.5ms.
+Using this shielding method the measured capacitance will only be slightly affected by the grounded underlying arm.
 
@@
 	\centering
-% 		\includegraphics[trim=1cm 1cm 1cm 1cm,clip,width=\textwidth]{Images/pseudocode.pdf}
+ 		\includegraphics[trim=1cm 1cm 1cm 1cm,clip,width=\textwidth]{Images/pseudocode.pdf}
 	\caption{Measurement procedure}
@@
 \end{figure}
+
+Figure \ref{fig:pseudocode} illustrates the operations required for a measurement cycle.
+The proximity sensor board generates a clock signal with varying frequency depending on the charge and discharge cycles.
+The HedgeHog measures the resulting capacitance by counting the signal's edges over a gate time of approximately 9.5ms.
+During that counting phase, the HedgeHog is sent to sleep in order to reduce power consumption.
+In the following, the HedgeHog applies run-length encoding on the measured data to reduce overhead and periodically logs the data to the integrated SD card.
+
 \section{Experiment}
@@
 
-[EB: Hier plot von drinking und grasping into locker]
-
 Figure \ref{fig:drinking} depicts the data that was recorded for the activity drinking water.
 The accelerometer data shows that there periodic up- and down-movements while the capacitive proximity sensor delivers data that is associated to the proximity of the table.
-Figure \ref{fig:opendoor} illustrates a grasp into a locker while preparing lunch.
-It is obvious that the locker's bottom surface that is located very closely to the wristband has an influence on the proximity sensor.
+Figure \ref{fig:making-bread} illustrates the preperation of a bread with marmelade while preparing lunch.
+It is obvious that the capacitive proximity sensor's data provides valuable information about the distance of the wristband to the table.
 
@@
 
+We have shown that current wearable inertial sensing devices for activity recognition can be significantly enhanced by integrating a capacitive proximity sensor into the wristband.
+Regarding the classication accuracy, we obtained an overall increase in correctly classified instances of TODO percent.
+Some classes, for instance ``preparing bread'' and ``drinking'', could benefit very much from proximity-related sensor data.
+For such classes the classication performance could be boosted by up to TODO percent.
+
+The classification results could be improved by the usage of more than one sensing electrode in the wristband.
+For example the wristband could integrate four electrodes that are placed on each side of the arm.
+However, this will lead to smaller electrode surfaces that lead to a decrease of sensing distance.
+We think that these considerations are an important subject of future work.
+The sensor's power consumption can be decreased by shorter measurement windows and the choice of more energy efficient hardware components.
+Measuring pulse width lengths instead of counting the rising edges of the sensor's signal may reduce the required time needed for a measurement.
+
+Capacitive proximity sensors represent a suitable new input modality for future activity recognition systems.
+The low power consumption as well as the invisible integration of a sensor into the wristband meets an essential requirement of wearable applications.
+Especially in AAL environments, these systems can help monitoring the course of chronic diseases by recognizing activities of daily life.
+This may improve the quality of life of persons affected and their caregivers.
+
 \section*{Acknowledgments}
-% We would like to thank the students of TU Darmstadt and the employees of Fraunhofer IGD that lent us their sitting and lying abilities for the evaluation. 
-We would like to thank our evaluation participants from the TU Darmstadt and the Fraunhofer IGD.
+We would like to thank our evaluation participants from TU Darmstadt and Fraunhofer IGD.
 

### 2012-06-18 Auswertung/plot_results.m
new plots

@@ -1,7 +1,7 @@
 %recordings = [csvread('feugen.csv')];
 %classes = [csvread('ae.csv')];
-recordings = nagi;
-classes = ann_nagi;
+recordings = eugen;
+classes = ann_eugen;
 
 sum_acc = sum(recordings(:,2));
 num_recordings = size(recordings);
@@ -29,16 +29,16 @@ time = time';
 %stop = 62320;
 
 % making bread
-%start = 39837;
-%stop = 42168;
+start = 39837;
+stop = 42168;
 
 % drink
-%start = 50550;
-%stop = 60890;
+start = 50550;
+stop = 60890;
 
 % all
-start = + 10000;
-stop = sum_acc - 10000;
+%start = + 10000;
+%stop = sum_acc - 10000;
 
 delta = stop - start;
 window_size = 20;
@@ -50,11 +50,11 @@ acceleration_z = recordings(:,5);
 cap = recordings(:,6);
 
 % smooth all recorded data by average filter
-average_kernel = ones(100, 1) * 0.0125;
-%cap = conv(cap, average_kernel, 'same');
-%acceleration_x = conv(acceleration_x, average_kernel, 'same');
-%acceleration_y = conv(acceleration_y, average_kernel, 'same');
-%acceleration_z = conv(acceleration_z, average_kernel, 'same');
+average_kernel = ones(10, 1) * 0.1;
+cap = conv(cap, average_kernel, 'same');
+acceleration_x = conv(acceleration_x, average_kernel, 'same');
+acceleration_y = conv(acceleration_y, average_kernel, 'same');
+acceleration_z = conv(acceleration_z, average_kernel, 'same');
 
 % resulting values
 res_cap = cap(start:stop);

### 2012-06-18 Auswertung/write_features.m

@@ -3,12 +3,17 @@ features_file_with_cap = 'features_with_cap.arff';
 
 %classes = [csvread('ae.csv'); csvread('am.csv')];
 %recordings = [csvread('feugen.csv'); csvread('fmarko.csv')];
-classes = ann_enzo;
-recordings = enzo;
+classes = ann_marko;
+recordings = marko;
 
 window_size = 100;
-window_overlap = 50;
+window_overlap = window_size/2;
 number_features = 8;
+
+recordings(:, 3) = smooth(recordings(:, 3),10);
+recordings(:, 4) = smooth(recordings(:, 4),10);
+recordings(:, 5) = smooth(recordings(:, 5),10);
+recordings(:, 6) = smooth(recordings(:, 6),10);
 
 number_recordings = size(recordings);
 number_recordings = number_recordings(1);
@@ -22,7 +27,7 @@ while i < (number_recordings)
     current_class = classes(i);
     extract_features = true;
    
-    % determine window size by acc
+    % determine window size by counter
     end_of_window = i;
     sum_acc = 0;
     while sum_acc < window_size

### 2012-06-18 Paper/Paper.tex
typos fixed

@@
 The accelerometer data shows that there periodic up- and down-movements while the capacitive proximity sensor delivers data that is associated to the proximity of the table.
-Figure \ref{fig:making-bread} illustrates the preperation of a bread with marmelade while preparing lunch.
+Figure \ref{fig:making-bread} illustrates the preparation of a bread with marmalade while preparing lunch.
 It is obvious that the capacitive proximity sensor's data provides valuable information about the distance of the wristband to the table.
@@
 We have shown that current wearable inertial sensing devices for activity recognition can be significantly enhanced by integrating a capacitive proximity sensor into the wristband.
-Regarding the classication accuracy, we obtained an overall increase in correctly classified instances of TODO percent.
+Regarding the classification accuracy, we obtained an overall increase in correctly classified instances of TODO percent.
 Some classes, for instance ``preparing bread'' and ``drinking'', could benefit very much from proximity-related sensor data.
-For such classes the classication performance could be boosted by up to TODO percent.
+For such classes the classification performance could be boosted by up to TODO percent.
 

### 2012-06-19 Auswertung/write_features.m
added classifications svm

@@ -3,8 +3,8 @@ features_file_with_cap = 'features_with_cap.arff';
 
 %classes = [csvread('ae.csv'); csvread('am.csv')];
 %recordings = [csvread('feugen.csv'); csvread('fmarko.csv')];
-classes = ann_marko;
-recordings = marko;
+classes = ann_eugen;
+recordings = eugen;
 
 window_size = 100;
 window_overlap = window_size/2;
@@ -61,7 +61,7 @@ while i < (number_recordings)
     end
     
     % if the window is ok, do the feature extraction
-    if extract_features == true && current_class ~= 0
+    if extract_features == true
         accel_1 = recordings(i:end_of_window, 3);
         accel_2 = recordings(i:end_of_window, 4);
         accel_3 = recordings(i:end_of_window, 5);
@@ -114,7 +114,7 @@ fprintf(fileID, '@ATTRIBUTE mean_accel_3 NUMERIC\r\n');
 fprintf(fileID, '@ATTRIBUTE var_accel_1 NUMERIC\r\n');
 fprintf(fileID, '@ATTRIBUTE var_accel_2 NUMERIC\r\n');
 fprintf(fileID, '@ATTRIBUTE var_accel_3 NUMERIC\r\n');
-fprintf(fileID, '@ATTRIBUTE class {class1,class2,class3,class4,class5,class6,class7,class8,class9}\r\n');
+fprintf(fileID, '@ATTRIBUTE class {class0,class1,class2,class3,class4,class5,class6,class7,class8,class9}\r\n');
 fprintf(fileID, '\r\n');
 
 fprintf(fileID, '@DATA\r\n');
@@ -141,7 +141,7 @@ fprintf(fileID, '@ATTRIBUTE var_accel_2 NUMERIC\r\n');
 fprintf(fileID, '@ATTRIBUTE var_accel_3 NUMERIC\r\n');
 fprintf(fileID, '@ATTRIBUTE mean_cap NUMERIC\r\n');
 fprintf(fileID, '@ATTRIBUTE var_cap NUMERIC\r\n');
-fprintf(fileID, '@ATTRIBUTE class {class1,class2,class3,class4,class5,class6,class7,class8,class9}\r\n');
+fprintf(fileID, '@ATTRIBUTE class {class0,class1,class2,class3,class4,class5,class6,class7,class8,class9}\r\n');
 fprintf(fileID, '\r\n');
 
 fprintf(fileID, '@DATA\r\n');

### 2012-06-19 Paper/Paper.tex
updated two images in the paper
updated two images
changed tex

@@
 Thus, other proximity sensing input modalities like ultrasound or optical measurements are not applicable for this type of mobile application.
-
-[MB; EB: Introduction to HedgeHog device]
 
@@
 
-image : photo
-
 \begin{figure}[htbp]
@@
 		\includegraphics[width=1.00\textwidth]{Images/schematic.pdf}
-	\caption{The sensing circuit is based on a timer with an operational amplifier that acts as a voltage follower.}
+	\caption{The sensing circuit is based on a timer with an operational amplifier that acts as a voltage follower.
+The wire that is labeled with "`antenna"' leads to the sensing electrode, whereas the wire labeled with "`guard"' leads to the shield electrode.}
 	\label{fig:schematic}
@@
 A fine-grained monitoring of such activities may help elderly or people suffering from mental diseases to maintain a healthy day/night rhythm and take action if irregularities occur.
+Figure blubb shows the modified HedgeHog activity logger that has been extended with a capacitive proximity sensor.
+The wristband has two electrodes, a sensing electrode underneath a slightly bigger shield electrode.
+
+[Hier Figure Aufbau]
 
@@
 	\centering
-		\includegraphics[trim=1.9cm 0.5cm 1.8cm 0.5cm,clip,width=\textwidth]{Images/drinking.pdf}
+		\includegraphics[width=\textwidth]{../Auswertung/images/eugen_7.pdf}
 	\caption{An example of the ``drinking'' activity with raw 3D acceleration data (lower plot) and proximity sensor data (upper plot).
@@
 	\centering
-		\includegraphics[trim=1.9cm 0.5cm 1.8cm 0.5cm,clip,width=\textwidth]{Images/making-bread.pdf}
+		\includegraphics[width=\textwidth]{../Auswertung/images/eugen_5.pdf}
 	\caption{An example of the ``preparing bread'' activity, where the participants had to put marmalade on a toast.
@@
 
-In order to evaluate if capacitive proximity sensors in wrist-bands can enhance the performance of activity recognition, we have conducted an evaluation with ten test persons.
+In order to evaluate if capacitive proximity sensors in wrist-bands can enhance the performance of activity recognition, we have conducted an evaluation with seven test persons.
 All test persons received a plot with the activities they were supposed to perform.
-After manual labeling, we used this test-set as ground truth and performed a cross-validation on an support-vector machine (SVM) classifier once with and once without including the data of the capacitive proximity sensor into the feature set.
+After manual labeling, we used this test-set as ground truth and performed a 4-fold cross-validation on an support-vector machine (SVM) classifier once with and once without including the data of the capacitive proximity sensor into the feature set.
+We used an SVM classifier because of its high relevance in activity recognition and its fast performance.
 
-[EB MB: kurze Einführung in SVMs]
-
-The classifier was trained with various features that were extracted from windows with the length of 5 seconds.
-In order to suppress noise contained in the capacitive proximity sensing data, we applied a Gaussian kernel with a size of [TODO] samples.
+The classifier was trained with various features that were extracted from windows with the length of 1 second.
+It turned out that greater window sizes do not provide better results.
+In order to suppress noise contained in the capacitive proximity sensing data, we applied a moving average filter with a kernel size of 10.
 The final feature set contained the arithmetic mean and standard variance of both accelerometer and proximity data.
-Moreover, we included the median, min and max features for both modalities.
+These simple feature types represent standard features applied in activity recognition.
 
@@
 
-\newpage
+% \newpage
 

### 2012-06-20 Paper/Paper.tex
added lots of references - references for activity recognition - hardware sensor image replaced by a new one

@@
 
-[MB; EB: Related Work activity recognition]
-
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 
-The field of activity recognition relying on wearable sensors mostly mostly considers inertial data from the participants body to infer performed activities, such as in the works of [bla,foo].
-The acceleration data is often augmented with data from sensors such as gyroscopes, magnetometers, ambient light or temperature, aiming at extracting more motion and environmental context of the user.
-[references?!]
+The field of activity recognition relying on wearable sensors mostly considers inertial data from the participants body to infer performed activities, such as in the works of \cite{Ravi2005,Bao2004,Srinivasan2010,Amft2005}.
+The acceleration data is often augmented with data from sensors such as gyroscopes \cite{Holleczek_2010}, magnetometers \cite{Altun_2010}, ambient light \cite{Borazio2012} or ambient and skin temperature \cite{Krause_2003}, aiming at extracting more motion and environmental context of the user.
+Researchers in \cite{wyss2010recognition}, for example, use heart rate information additionally to the accelerometer to detect military related activities like lifting and lowering loads or even digging.
+In \cite{Ward_2006}, workshop assembly activities were detected by augmenting acceleration sensors with microphones.
 
-The works of \cite{intel_bracelet} and ...
-show that detecting objects the user touches or uses throughout the day can be very helpful for the activity recognition, including activities of daily living (maja(?)), maintenance tasks (?), gardening activities, etc.
-For example, Berlin et al.
-\cite{berlin_laerhoven_tei_2010} have augmented a wrist-worn accelerometer sensor node with an RFID bracelet to be able to detect objects a human is interacting with.
+The works of \cite{Fishkin_2005} and \cite{patterson2005fga} show that detecting objects the user touches or uses can be very helpful for activity recognition, including activities of daily living \cite{phealth:maja} and \cite{Philipose_2004}, activity tracking in car manufacturing \cite{Stiefmeier08}, etc.
+Berlin et al.
+\cite{berlin_laerhoven_tei_2010} have augmented a wrist-worn accelerometer sensor node with an RFID bracelet to be able to detect objects a human is interacting with, aiming both at household as well as gardening activities.
+
+% physical activities with accelerometers detected via motifs \cite{minnen_2006}
+
+% detection of eating and drinking activities from acceleration sensors \cite{Amft2005}
 
@@
 	\centering
-	\includegraphics[width=\textwidth]{Images/hardware_sensor.jpg}
-	\caption{The inertial data logger featuring a low-power microcontroller, a 3 axis accelerometer, a microSD flash card for storing the sensor data and a USB connector for accessing the stored data (on the left) is being powered by a small lithium polymer battery and is packaged into a plastic case to be worn at the wrist (a version with a OLED display on the right).}
+	\includegraphics[width=\textwidth]{Images/hardware_sensor_2.jpg}
+	\caption{The inertial data logger featuring a low-power microcontroller, a 3 axis accelerometer, a microSD flash card for storing the sensor data and a USB connector for accessing the stored data (on the right) is being powered by a small lithium polymer battery and is packaged into a plastic case to be worn at the wrist (a version with a OLED display on the left).}
 	\label{fig:sensornode}

### 2012-06-20 Paper/Paper.tex
results
image added and references fixed  - wristband with electrodes and sensors  - removed urls from references

@@
 	\caption{The sensing circuit is based on a timer with an operational amplifier that acts as a voltage follower.
-The wire that is labeled with "`antenna"' leads to the sensing electrode, whereas the wire labeled with "`guard"' leads to the shield electrode.}
+The wire that is labeled with ``antenna'' leads to the sensing electrode, whereas the wire labeled with ``guard'' leads to the shield electrode.}
 	\label{fig:schematic}
@@
 
-\begin{figure}[htbp]
+\begin{figure}
+	\centering
+		\includegraphics[width=1.00\textwidth]{Images/capacitive_sensor_wristband_2.jpg}
+	\caption{The hardware prototype at a glance: HedgeHog activity logger at the lower right, the capacitive sensor unit at the lower left, and the wristband with the sensing and the shield electrodes on-top each other covered with adhesive tape for isolation purposes.}
+	\label{fig:cap_sensor}
+\end{figure}
+
+Figure \ref{fig:cap_sensor} shows the the wrist-worn prototype used in the evaluation experiments, with the HedgeHog as the main data logger, the capacitive sensor circuit and the wristband holding the sensing and shielding electrodes.
+
+\begin{figure}
 	\centering
  		\includegraphics[trim=1cm 1cm 1cm 1cm,clip,width=\textwidth]{Images/pseudocode.pdf}
-	\caption{Measurement procedure}
+	\caption{Measurement procedure carried out by the HedgeHog sensor: using the microcontroller's Timer0 module in counting mode, the oscillating signal generated by the capacitive sensor circuit can be measured by counting the frequency pulses over a predefined gate time (approximately 9.5ms).}
 	\label{fig:pseudocode}

### 2012-06-21 Paper/Paper.tex
writing paper - related work on activity recognition - structure overview and section labels

@@
 % \urldef{\mailcf}\path|{berlin,borazio}@ess.informatik.tu-darmstadt.de, tobias.grosse-puppendahl@igd.fraunhofer.de|
-% \newcommand{\keywords}[1]{\par\addvspace\baselineskip
+% \newcommand{\keywords}[1]{\par\addvspace\+baselineskip
 % \noindent\keywordname\enspace\ignorespaces#1}
@@
 
+The remainder of this paper is  is structured as follows: First, Section \ref{sect:related} is dedicated to related work work in activity recognition, in particular considering approaches with multiple sensor modalities.
+Section \ref{sect:hardware} will present the hardware based on which a wrist-worn sensor prototype was built, fusing together an inertial data logger and a capacitive sensor integrated in a wristband.
+The experimental setup, the scripted scenario with daily activities and the activity recognition evaluation results, showing the performance boost of the capacitive sensor unit, are given in Section \ref{sect:experiment}.
+The paper is wrapped up with a conclusions section enumerating the findings of the paper, as well as pointing out future research potential.
+
 \section{Related Work}
-
+\label{sect:related}
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 
-The works of \cite{Fishkin_2005} and \cite{patterson2005fga} show that detecting objects the user touches or uses can be very helpful for activity recognition, including activities of daily living \cite{phealth:maja} and \cite{Philipose_2004}, activity tracking in car manufacturing \cite{Stiefmeier08}, etc.
-Berlin et al.
-\cite{berlin_laerhoven_tei_2010} have augmented a wrist-worn accelerometer sensor node with an RFID bracelet to be able to detect objects a human is interacting with, aiming both at household as well as gardening activities.
+The works of \cite{Fishkin_2005} and \cite{patterson2005fga} show that detecting objects the user touches or uses can be very helpful for activity recognition.
+By using RFID readers that are embedded in gloves or bracelets at the wrist and RFID tags attached to various objects of interest, one can detect the object grasped and used by the user, thus aiding the activity recognition in various application scenarios, such as activities of daily living \cite{phealth:maja} and \cite{Philipose_2004}, activity tracking in car manufacturing \cite{Stiefmeier08}, or household and gardening activities \cite{berlin_laerhoven_tei_2010}.
 
-% physical activities with accelerometers detected via motifs \cite{minnen_2006}
+Our approach to enhance the inertial data from a wearable sensor is quite similar to the RFID scenarios just mentioned, as it also relies on a single wearable sensor and thus is aiming at an unobtrusive deployment.
+The main difference lies in the fact that we do not need accurate detection of tagged objects, but are only interested in the proximity to various objects which, as we expect, could boost activity recognition.
 
-% detection of eating and drinking activities from acceleration sensors \cite{Amft2005}
+% While the RFID gloves or bracelets might not have reached user acceptance, and deploying tags in the user's home might appear cumbersome and still quite intrusive, our approach considers a single wrist-worn sensor unit consisting of an accelerometer and proximity sensor.
+% Berlin et al.
+\cite{berlin_laerhoven_tei_2010} have also fused and benchmarked a wrist-worn accelerometer sensor node with an RFID bracelet to be able to detect objects a human is interacting with, aiming both at household as well as gardening activities.
 
@@
 		\includegraphics[width=1.00\textwidth]{Images/modes.pdf}
-	\caption{Three different capacitive proximity sensing modes can be distinguished.
-The sensing electrodes build up an electric field to objects in the environment, illustrated with a cloud.
-\cite{Smith1996}}
+	\caption{Three different capacitive proximity sensing modes can be distinguished \cite{Smith1996}.
+The sensing electrodes build up an electric field to objects in the environment, illustrated with a cloud.}
 	\label{fig:modes}
@@
 \section{Hardware}
+\label{sect:hardware}
 This section presents the two components of the required hardware, the wrist-worn activity data logger tailored to capture acceleration data, and the capacitive proximity sensor used for distance measurements.
@@
 The shield is driven with the same potential as the sensing electrode, such that the the capacitance between the two electrodes is negligible.
-Using this shielding method the measured capacitance will only be slightly affected by the grounded underlying arm.
+Using this shielding method the measured capacitance will only be slightly
+affected by the grounded underlying arm.
 
@@
 \section{Experiment}
+\label{sect:experiment}
 
@@
 \subsection{Evaluation Results}
+\label{sect:evaluation}
 
 \section{Conclusion and Outlook}
+\label{sect:conclusions}
 
@@
 \section*{Acknowledgments}
+
 We would like to thank our evaluation participants from TU Darmstadt and Fraunhofer IGD.

### 2012-06-21 Paper/Paper.tex
changed text

--- 2012-06-20 Paper/Paper.tex
@@
 \begin{abstract}
-Zwei flinke Boxer jagen die quirlige Eva und ihren Mops durch Sylt.
-Franz jagt im komplett verwahrlosten Taxi quer durch Bayern.
-Zwölf Boxkämpfer jagen Viktor quer über den großen Sylter Deich.
-Vogel Quax zwickt Johnys Pferd Bim.
-Sylvia wagt quick den Jux bei Pforzheim.
-Polyfon zwitschernd aßen Mäxchens Vögel Rüben, Joghurt und Quark.
-"Fix, Schwyz!
-" quäkt Jürgen blöd vom Paß.
-Victor jagt zwölf Boxkämpfer quer über den großen Sylter Deich.
-Falsches Üben von Xylophonmusik quält jeden größeren Zwerg.
+Activity recognition solely with an accelerometer is a common investigated research topic and enables the detection of basic activities like sit, walk or stand.
+Recent work in this area adds different sensing modalities to the inertial data to collect more information of the users environment and to boost activity recognition.
+We present a sensor set-up that consists of an accelerometer and a capacitive proximity sensor that senses the users activity based on the combined sensor values.
+We show that our proposed approach improves the recognition rate significantly for detecting activities of daily living (e.g.
+open door, prepare dinner, etc.) by adding capacitive proximity sensor data to the inertial data.
 \keywords{ambient assisted living, capacitive proximity sensors, activity recognition, user context}
@@
 \section{Introduction}
+
+Sensing a persons activity is being researched for several years now, raising even more interest in it by the advanced technology of mobile phones, which include multiple sensors and therefore enable activity recognition with such a device \cite{brezmes2009activity}.
 Current wearable activity recognition systems are able to unobtrusively capture and recognize a person's activities throughout the whole day.
-These systems often rely on inertial sensor data that is captured by wearable sensors attached to the body to identify the activities performed.
-However, it is a great challenge to identify fine-grained activities just by using this single modality.
-In this paper, we have obtained and use a wrist-worn activity data logger\footnote{HedgeHog Activity Logger -- \url{http://www.ess.tu-darmstadt.de/hedgehog}} and investigate the possibility of integrating a capacitive proximity sensor unobtrusively into the wristband to obtain higher recognition rates for several activities in daily life.
-
-Capacitive proximity sensors can indirectly measure the distance and nature of a grounded object within reach.
+These systems often rely on inertial sensor data that is captured by wearable sensors embedded in a mobile device \cite{brezmes2009activity} or attached to the body \cite{Ravi2005} to identify the activities performed.
+Usually, single sensor modalities are used or duplicated to detect the activities.
+However, it is a great challenge to identify fine-grained activities just by using a single modality like the accelerometer.
+Capacitive proximity sensors on the other hand can indirectly measure the distance and nature of a grounded object within reach.
 This means that the measurement result depends on the object's distance, its size and the material it is made of.
-A wristband worn capacitive proximity sensor requires a shield that eliminates the influence of the grounded arm directly underneath the sensor.
-Using this setup, we can detect the proximity to a grounded object in the environments for distances up to 20 cm.
-Especially for mobile devices, it is required that the sensor only draws a very small amount of power.
-Thus, other proximity sensing input modalities like ultrasound or optical measurements are not applicable for this type of mobile application.
+In this work, we show how an accelerometer and a capacitive proximity sensor can be used to improve activity recognition in activities of daily living.
+For this we obtained a wrist-worn activity data logger\footnote{HedgeHog Activity Logger -- \url{http://www.ess.tu-darmstadt.de/hedgehog}} and integrate a capacitive proximity sensor unobtrusively into the wristband.
 
@@
 
+A wristband worn capacitive proximity sensor requires a shield that eliminates the influence of the grounded arm directly underneath the sensor.
+Using this setup, we can detect the proximity to a grounded object in the environments for distances up to 20 cm.
+Especially for mobile devices, it is required that the sensor only draws a very small amount of power.
+Thus, other proximity sensing input modalities like ultrasound or optical measurements are not applicable for this type of mobile application.
+
+This paper is structured as follows: We give a short insight in related work in chapter 2, followed by describing the hardware set-up used for this study in chapter 3.
+The experiment and evaluation will be described in chapter 4, showing how data has been obtained and how activity recognition is being improved.
+This work is concluded in chapter 5 with an outlook to future work.
+
 \section{Related Work}
@@
 
-The field of activity recognition relying on wearable sensors mostly considers inertial data from the participants body to infer performed activities, such as in the works of \cite{Ravi2005,Bao2004,Srinivasan2010,Amft2005}.
+Activity recognition research relying on wearable sensors mostly considers inertial data from the participants body to infer performed activities, such as in the works of \cite{Ravi2005,Bao2004,Srinivasan2010,Amft2005}.
 The acceleration data is often augmented with data from sensors such as gyroscopes \cite{Holleczek_2010}, magnetometers \cite{Altun_2010}, ambient light \cite{Borazio2012} or ambient and skin temperature \cite{Krause_2003}, aiming at extracting more motion and environmental context of the user.
@@
 
-[Hier Figure Aufbau]
-
-The recorded test set contains the following activities: opening door, sitting on a couch, lying on a couch, grasping into a locker, putting something from a shelf, sitting at the table, preparing a bread with marmalade, eating the bread, pouring water, drinking the water and sleeping.
+The recorded test set contains the following activities: opening door, sitting on a couch, lying on a couch, grasping into a locker, sitting at the table, preparing a bread with marmalade, eating the bread, drinking water and sleeping.
 Some of those activities are very hard to recognize without any further contextual information, when the data is limited to a single modality like a 3D accelerometer.
 For example, sitting at the table and sitting on a couch are very similar activities.
-The data of a capacitive proximity sensor can extend this data basis by providing that is related to the material an object is made of.
+We aim to show that the data basis can be significantly improved by the additional input modality.
 
-\begin{figure}
-	\centering
-		\includegraphics[width=\textwidth]{../Auswertung/images/eugen_7.pdf}
-	\caption{An example of the ``drinking'' activity with raw 3D acceleration data (lower plot) and proximity sensor data (upper plot).
-The participant took five nips from the glass, while leaving the hands lying on the table in between.
-These motions can be easily detected in the acceleration as well as proximity data.}
-	\label{fig:drinking}
-\end{figure}
-
-\begin{figure}
-	\centering
-		\includegraphics[width=\textwidth]{../Auswertung/images/eugen_5.pdf}
-	\caption{An example of the ``preparing bread'' activity, where the participants had to put marmalade on a toast.
-Note, that the acceleration data does not seem to have any characteristic patterns, while the proximity sensor indicates a table, plate, or other objects in immediate distance.}
-	\label{fig:making-bread}
-\end{figure}
-
-Figure \ref{fig:drinking} depicts the data that was recorded for the activity drinking water.
-The accelerometer data shows that there periodic up- and down-movements while the capacitive proximity sensor delivers data that is associated to the proximity of the table.
-Figure \ref{fig:making-bread} illustrates the preparation of a bread with marmalade while preparing lunch.
-It is obvious that the capacitive proximity sensor's data provides valuable information about the distance of the wristband to the table.
-
-In order to evaluate if capacitive proximity sensors in wrist-bands can enhance the performance of activity recognition, we have conducted an evaluation with seven test persons.
-All test persons received a plot with the activities they were supposed to perform.
-After manual labeling, we used this test-set as ground truth and performed a 4-fold cross-validation on an support-vector machine (SVM) classifier once with and once without including the data of the capacitive proximity sensor into the feature set.
+In order to evaluate if capacitive proximity sensors in wrist-bands can enhance the performance of activity recognition, we have conducted an evaluation with 7 test persons.
+All test persons received a basic plot with the activities they were supposed to perform.
+They were not given any instructions about the way they are supposed to perform the activities.
+After manual labeling, we used this test-set as ground truth and performed a 4-fold cross-validation on an support-vector machine (SVM) classifier on each user, once with and once without including the data of the capacitive proximity sensor into the feature set.
 We used an SVM classifier because of its high relevance in activity recognition and its fast performance.
@@
 In order to suppress noise contained in the capacitive proximity sensing data, we applied a moving average filter with a kernel size of 10.
-The final feature set contained the arithmetic mean and standard variance of both accelerometer and proximity data.
+The final feature set contained the arithmetic mean, min, max, median and standard variance of both accelerometer and proximity data.
 These simple feature types represent standard features applied in activity recognition.
@@
 \subsection{Evaluation Results}
+
+In the following, the performed activities will be analyzed in detail, stating out the influence of the capacitive proximity sensor on the classification result.
+In general, the usage of data provided by the new modality showed improvements in recognition rates by 2.4 - 14.4 \%.
+
+\begin{figure}[htbp]
+	\centering
+		\includegraphics[width=1.00\textwidth]{../Auswertung/images/tobias_1.pdf}
+	\caption{While entering the door, the wristband approaches the door knob twice, while opening and closing.
+Regarding the capacitive proximity sensor, this situation can be observed at the beginning and at the end of the activity.}
+	\label{fig:tobias_1}
+\end{figure}
+
+The opening door activity has very poor recongition rates without the including the proximity sensor.
+The average f-measure could be increased from of 34.2 \% to 46.8 \%.
+A plot of the activity is given in Figure \ref{fig:tobias_1}.
+The capacitive proximity sensor shows two approaches to the door knob, one for opening the door (2 s) and one for closing the door (9 s).
+The acceleration sensor captures relevant data in the time in which the person moves into the room and the hand changes from the outer to the inner door knob (5 - 7 s).
+The recorded data for this activity shows strong correlations between all test persons.
+
+\begin{figure}[htbp]
+	\centering
+		\includegraphics[width=1.00\textwidth]{../Auswertung/images/eugen_2.pdf}
+	\caption{Example of a ``sitting'' activity in which user changed his position very frequently.
+Throughout the whole activity the values of the proximity sensor stay constant due to the close proximity to the couch's fabric.}
+	\label{fig:eugen_2}
+\end{figure}
+
+After closing the door, the test persons were supposed to sit down on the couch.
+It turned out that there are great variations of the sitting posture and the corresponding hand positions.
+Many users tapped with their fingers or hands while sitting, and changed their sitting positions very frequently as shown in Figure \ref{fig:eugen_2}.
+In this case, it is obvious that the data from the acceleration sensor is very difficult to interprete as there are numerous changes in the axial orientation of the sensor.
+However, the capacitive proximity sensor is able to indicate when a hand is placed on the surface of the couch.
+Especially for this test person, the f-measure increased from 27.4 \% to 66.7 \%, while the average f-measure improved from 68.1 \% to 77.6 \%.
+
+In the following, the test persons were supposed to lay down on the couch.
+Again, there were great variations in performing this activity between all participants.
+For example, some of them crossed their hands under their head.
+For that class, the average f-measure could only be increased by 2.4 \%, from 82.1 to 84.5 \%.
+The participants were also asked to put food and dishes from a shelf and a locker on a table.
+The capacitive proximity sensor captured the proximity to the shelf and to the table.
+The overall f-measure for this activity is quite low, but improved by 6.7 \% from 50.4 \% to 57.1\%.
+
+\begin{figure}[htpb]
+	\centering
+		\includegraphics[width=\textwidth]{../Auswertung/images/eugen_5.pdf}
+	\caption{An example of the ``preparing bread'' activity, where the participants had to put marmalade on a toast.
+The proximity sensor indicates the closeness to the table, while the acceleration sensor shows reoccurring movements.}
+	\label{fig:eugen_5}
+\end{figure}
+
+Figure \ref{fig:eugen_5} shows an example instance of preparing a bread with marmelade.
+It is notable that the acceleration data does not seem to provide any characteristic patterns, while the proximity sensor indicates a table, plate, or other objects in immediate distance.
+This activity showed a high improvement in the average f-measure by 10 \%, from 49.0 \% to 59.8 \%, when the capacitive proximity sensor is taken into account.
+
+\begin{figure}[htpb]
+	\centering
+		\includegraphics[width=\textwidth]{../Auswertung/images/eugen_6.pdf}
+	\caption{An eating person who took 5 bites from the bread. After each bite, the hand is placed on the table.}
+	\label{fig:eugen_6}
+\end{figure}
+
+Considering the eating activity, the influence of the capacitive proximity sensor is very low.
+Some of the participants ate their bread leaving their hand close to the mouth, while others moved their hand up and down putting their bread aside on the plate.
+A good example of an eating activity is shown in Figure \ref{fig:eugen_6} where the participant took a few bites from the bread.
+The average f-measure for that activity increased slightly from 80.3 \% to 83.8 \%.
+
+\begin{figure}[htpb]
+	\centering
+		\includegraphics[width=\textwidth]{../Auswertung/images/eugen_7.pdf}
+	\caption{An example of the ``drinking'' activity of a test person.
+The participant first pours some water into the glass and then takes three drams of water.
+After each dram, he returns his arm to the table which can be observed in the characteristic patterns of the proximity sensor.}
+	\label{fig:eugen_7}
+\end{figure}
+
+The activity drinking water is depicted in Figure \ref{fig:eugen_7}.
+The participant took five nips from the glass, while leaving the hands lying on the table in between.
+These motions can be easily detected in the acceleration as well as proximity data.
+The accelerometer data shows that there periodic up- and down-movements while the capacitive proximity sensor delivers data that is associated to the proximity of the table.
+
+\begin{figure}[htbp]
+	\centering
+		\includegraphics[width=1.00\textwidth]{../Auswertung/images/marko_8.pdf}
+	\caption{An examplary instance of the class ``walking''.
+The acceleration sensor and the proximity sensor show periodic reoccuring patterns that are related to the pendulum-like arm movement and the proximity to the person's body during those movements.}
+	\label{fig:marko_8}
+\end{figure}
+
+Regarding the walking activity one can identify periodic changes in the acceleration as well as in the measured capacitance, illustrated in Figure \ref{fig:marko_8}.
+While walking, the capacitance between the wristband and the leg increases when the wristband is located close to the body and decrease when the wristband moves away.
+
+\begin{figure}[htbp]
+	\centering
+		\includegraphics[width=1.00\textwidth]{../Auswertung/images/eugen_9.pdf}
+	\caption{During the sleeping activity the data from both sensors remains constant. One draw conclusions about the coverage of the arm with either couchons, blankets or the proximity to the mattress.}
+	\label{fig:eugen_9}
+\end{figure}
+
+The sleeping activity was captured with an average f-measure of 83.4 \% that increased to 88.4 \% when using the capacitive proximity sensor.
+In this situation the capacitive proximity sensor is able to capture the surrounding couchons and blankets.
+The accelerometer data and the proximity data is very constant during that activity.
+
+\begin{table}[htbp]
+	\centering	\begin{tabular}{p{0.48cm}p{0.48cm}p{0.48cm}p{0.48cm}p{0.48cm}p{0.48cm}p{0.48cm}p{0.48cm}p{0.48cm}|p{0.48cm}p{0.48cm}p{0.48cm}p{0.48cm}p{0.48cm}p{0.48cm}p{0.48cm}p{0.48cm}p{0.48cm}|p{2.2cm}}
+a & b & c & d & e & f & g & h & i & a & b & c & d & e & f & g & h & i & $\leftarrow$ classified as\\
+\hline
+10 & 3 & 0 & 2 & 0 & 0 & 0 & 0 & 0 		& 9 & 5 & 0 & 0 & 1 & 0 & 0 & 0 & 0 & a = open door\\
+\hline
+2 & 259 & 4 & 2 & 4 & 2 & 4 & 0 & 0 	& 1 & 248 & 5 & 6 & 4 & 2 & 5 & 0 & 0 & b = sit\\
+\hline
+0 & 5 & 213 & 2 & 2 & 1 & 8 & 0 & 4 	& 0 & 3 & 212 & 3 & 2 & 1 & 6 & 0 & 8 & c = lying\\
+\hline
+3 & 5 & 2 & 52 & 9 & 3 & 1 & 1 & 1 		& 0 & 8 & 2 & 44 & 7 & 8 & 1 & 1 & 3 & d = get things\\
+\hline
+1 & 9 & 3 & 13 & 46 & 6 & 3 & 0 & 0 	& 1 & 7 & 1 & 11 & 37 & 9 & 2 & 0 & 10 & e = make bread\\
+\hline
+0 & 5 & 1 & 1 & 5 & 62 & 5 & 0 & 1 		& 0 & 3 & 2 & 3 & 8 & 58 & 5 & 0 & 1 & f = eat\\
+\hline
+0 & 8 & 4 & 2 & 1 & 5 & 68 & 0 & 0 		& 0 & 8 & 3 & 0 & 6 & 4 & 67 & 0 & 1 & g = drink\\
+\hline
+1 & 0 & 0 & 1 & 0 & 0 & 0 & 12 & 0 		& 1 & 0 & 0 & 1 & 0 & 0 & 0 & 13 & 0 & h = walk\\
+\hline
+1 & 0 & 0 & 7 & 0 & 0 & 0 & 0 & 325 	& 0 & 1 & 2 & 10 & 6 & 1 & 1 & 0 & 312 & i = sleep\\
+		\end{tabular}
+	\caption{Comparison of two Confusion matrices for an exemplary user. The left part includes the capacitive proximity sensor, the right part represents classification results without the sensor.}
+	\label{tab:ConfusionMatrixForAnExemplaryUser}
+\end{table}
+
+Table \ref{tab:confmatrix}
 

### 2012-06-21 Paper/Paper.tex
table and references updated

@@
 \begin{abstract}
-Activity recognition solely with an accelerometer is a common investigated research topic and enables the detection of basic activities like sit, walk or stand.
+Activity recognition solely with an accelerometer is a common investigated research topic and enables the detection of basic activities like sitting, walking or standing.
 Recent work in this area adds different sensing modalities to the inertial data to collect more information of the users environment and to boost activity recognition.
-We present a sensor set-up that consists of an accelerometer and a capacitive proximity sensor that senses the users activity based on the combined sensor values.
+This work presents a sensor prototype which consists of an accelerometer and a capacitive proximity sensor that senses the users activity based on the combined sensor values.
 We show that our proposed approach improves the recognition rate significantly for detecting activities of daily living (e.g.
-open door, prepare dinner, etc.) by adding capacitive proximity sensor data to the inertial data.
+sitting, preparing dinner, eating, drinking, etc.) by adding capacitive proximity sensor data to the inertial data.
 \keywords{ambient assisted living, capacitive proximity sensors, activity recognition, user context}
@@
 Capacitive proximity sensors on the other hand can indirectly measure the distance and nature of a grounded object within reach.
-This means that the measurement result depends on the object's distance, its size and the material it is made of.
+This means that the measurement result depends on the
+object's distance, its size and the material it is made of.
 In this work, we show how an accelerometer and a capacitive proximity sensor can be used to improve activity recognition in activities of daily living.
@@
 
-This paper is structured as follows: We give a short insight in related work in chapter 2, followed by describing the hardware set-up used for this study in chapter 3.
+% This paper is structured as follows: We give a short insight in related work in chapter 2, followed by describing the hardware set-up used for this study in chapter 3.
 The experiment and evaluation will be described in chapter 4, showing how data has been obtained and how activity recognition is being improved.
@@
 
+The remainder of this paper is  is structured as follows: First, Section \ref{sect:related} is dedicated to related work work in activity recognition, in particular considering approaches with multiple sensor modalities.
+Section \ref{sect:hardware} will present the hardware based on which a wrist-worn sensor prototype was built, fusing together an inertial data logger and a capacitive sensor integrated in a wristband.
+The experimental setup with the scripted scenario of daily activities and the activity recognition evaluation results, showing the performance boost of the capacitive sensor unit, are given in Section \ref{sect:experiment}.
+The paper is wrapped up with a conclusions section enumerating the findings of the paper, as well as pointing out future research potential.
+
 \section{Related Work}
-
+\label{sect:related}
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 
-\begin{figure}[htpb]
+\begin{figure}
 	\centering
@@
 
+\subsection{Evaluation Results}
+\label{sect:evaluation}
+
 The sleeping activity was captured with an average f-measure of 83.4 \% that increased to 88.4 \% when using the capacitive proximity sensor.
@@
 
-\begin{table}[htbp]
-	\centering	\begin{tabular}{p{0.48cm}p{0.48cm}p{0.48cm}p{0.48cm}p{0.48cm}p{0.48cm}p{0.48cm}p{0.48cm}p{0.48cm}|p{0.48cm}p{0.48cm}p{0.48cm}p{0.48cm}p{0.48cm}p{0.48cm}p{0.48cm}p{0.48cm}p{0.48cm}|p{2.2cm}}
-a & b & c & d & e & f & g & h & i & a & b & c & d & e & f & g & h & i & $\leftarrow$ classified as\\
-\hline
-10 & 3 & 0 & 2 & 0 & 0 & 0 & 0 & 0 		& 9 & 5 & 0 & 0 & 1 & 0 & 0 & 0 & 0 & a = open door\\
-\hline
-2 & 259 & 4 & 2 & 4 & 2 & 4 & 0 & 0 	& 1 & 248 & 5 & 6 & 4 & 2 & 5 & 0 & 0 & b = sit\\
-\hline
-0 & 5 & 213 & 2 & 2 & 1 & 8 & 0 & 4 	& 0 & 3 & 212 & 3 & 2 & 1 & 6 & 0 & 8 & c = lying\\
-\hline
-3 & 5 & 2 & 52 & 9 & 3 & 1 & 1 & 1 		& 0 & 8 & 2 & 44 & 7 & 8 & 1 & 1 & 3 & d = get things\\
-\hline
-1 & 9 & 3 & 13 & 46 & 6 & 3 & 0 & 0 	& 1 & 7 & 1 & 11 & 37 & 9 & 2 & 0 & 10 & e = make bread\\
-\hline
-0 & 5 & 1 & 1 & 5 & 62 & 5 & 0 & 1 		& 0 & 3 & 2 & 3 & 8 & 58 & 5 & 0 & 1 & f = eat\\
-\hline
-0 & 8 & 4 & 2 & 1 & 5 & 68 & 0 & 0 		& 0 & 8 & 3 & 0 & 6 & 4 & 67 & 0 & 1 & g = drink\\
-\hline
-1 & 0 & 0 & 1 & 0 & 0 & 0 & 12 & 0 		& 1 & 0 & 0 & 1 & 0 & 0 & 0 & 13 & 0 & h = walk\\
-\hline
-1 & 0 & 0 & 7 & 0 & 0 & 0 & 0 & 325 	& 0 & 1 & 2 & 10 & 6 & 1 & 1 & 0 & 312 & i = sleep\\
+\begin{table}
+	\caption{Comparison of two Confusion matrices for an exemplary user. The left part includes the capacitive proximity sensor, the right part represents classification results without the sensor.}
+
+	\setlength{\tabcolsep}{2.4pt}
+	\centering
+
+	\begin{scriptsize}
+	\begin{tabular}%{p{0.48cm}p{0.48cm}p{0.48cm}p{0.48cm}p{0.48cm}p{0.48cm}p{0.48cm}p{0.48cm}p{0.48cm}|p{0.48cm}p{0.48cm}p{0.48cm}p{0.48cm}p{0.48cm}p{0.48cm}p{0.48cm}p{0.48cm}p{0.48cm}|p{2.2cm}}
+	{|ccccccccc|p{0.01cm}|ccccccccc|p{0.01cm}|l}
+		\multicolumn{9}{c}{with proximity} &\multicolumn{1}{c}{}& 
+			\multicolumn{9}{c}{without proximity} &\multicolumn{1}{c}{}&  \\ \cmidrule{1-9} \cmidrule{11-19}
+
+		\multicolumn{1}{c}{a} & b & c & d & e & f & g & h & \multicolumn{1}{c}{i} &\multicolumn{1}{c}{}& 
+			\multicolumn{1}{c}{a} & b & c & d & e & f & g & h & \multicolumn{1}{c}{i} 	&\multicolumn{1}{c}{}& $\leftarrow$ classified as\\
+		\cline{1-9} \cline{11-19} \cline{21-21}
+		\textbf{10} & 3 & 0 & 2 & 0 & 0 & 0 & 0 & 0 	&& \textbf{9} & 5 & 0 & 0 & 1 & 0 & 0 & 0 & 0 		&& a = open door	\\ 
+		2 & \textbf{259} & 4 & 2 & 4 & 2 & 4 & 0 & 0 	&& 1 & \textbf{248} & 5 & 6 & 4 & 2 & 5 & 0 & 0 	&& b = sitting		\\ 
+		0 & 5 & \textbf{213} & 2 & 2 & 1 & 8 & 0 & 4 	&& 0 & 3 & \textbf{212} & 3 & 2 & 1 & 6 & 0 & 8 	&& c = lying		\\ 
+		3 & 5 & 2 & \textbf{52} & 9 & 3 & 1 & 1 & 1 	&& 0 & 8 & 2 & \textbf{44} & 7 & 8 & 1 & 1 & 3 		&& d = get things	\\ 
+		1 & 9 & 3 & 13 & \textbf{46} & 6 & 3 & 0 & 0 	&& 1 & 7 & 1 & 11 & \textbf{37} & 9 & 2 & 0 & 10 	&& e = make bread	\\ 
+		0 & 5 & 1 & 1 & 5 & \textbf{62} & 5 & 0 & 1 	&& 0 & 3 & 2 & 3 & 8 & \textbf{58} & 5 & 0 & 1 		&& f = eating		\\ 
+		0 & 8 & 4 & 2 & 1 & 5 & \textbf{68} & 0 & 0 	&& 0 & 8 & 3 & 0 & 6 & 4 & \textbf{67} & 0 & 1 		&& g = drinking		\\
+		1 & 0 & 0 & 1 & 0 & 0 & 0 & \textbf{12} & 0 	&& 1 & 0 & 0 & 1 & 0 & 0 & 0 & \textbf{13} & 0 		&& h = walking		\\
+		1 & 0 & 0 & 7 & 0 & 0 & 0 & 0 & \textbf{325} 	&& 0 & 1 & 2 & 10 & 6 & 1 & 1 & 0 & \textbf{312} 	&& i = sleeping		\\ 
+		\cline{1-9} \cline{11-19} 
 	\end{tabular}
-	\caption{Comparison of two Confusion matrices for an exemplary user. The left part includes the capacitive proximity sensor, the right part represents classification results without the sensor.}
+	\end{scriptsize}
+
 	\label{tab:ConfusionMatrixForAnExemplaryUser}
@@
 
-Table \ref{tab:confmatrix}
+Table \ref{tab:ConfusionMatrixForAnExemplaryUser}
 
 \section{Conclusion and Outlook}
+\label{sect:conclusions}
 

### 2012-06-21 Paper/Paper.tex
lines added

--- 2012-06-21 Paper/Paper.tex
@@
 It is notable that the acceleration data does not seem to provide any characteristic patterns, while the proximity sensor indicates a table, plate, or other objects in immediate distance.
-This activity showed a high improvement in the average f-measure by 10 \%, from 49.0 \% to 59.8 \%, when the capacitive proximity sensor is taken into account.
+This activity showed a high improvement in the average f-measure by 10 \%, from 49.0 \% to 59.8 \%, when the data delivered by the capacitive proximity sensor is taken into account.
 
@@
 The activity drinking water is depicted in Figure \ref{fig:eugen_7}.
-The participant took five nips from the glass, while leaving the hands lying on the table in between.
+The participant took a few nips from the glass, while leaving the hands lying on the table in between.
 These motions can be easily detected in the acceleration as well as proximity data.
@@
 \end{figure}
+=======
+\subsection{Evaluation Results}
+\label{sect:evaluation}
 
@@
 \section{Conclusion and Outlook}
+\label{sect:conclusions}
 

### 2012-06-21 Paper/Paper.tex
recent changes to the paper

--- 2012-06-21 Paper/Paper.tex
@@
 Capacitive proximity sensors on the other hand can indirectly measure the distance and nature of a grounded object within reach.
-This means that the measurement result depends on the
-object's distance, its size and the material it is made of.
+This means that the measurement result depends on the object's distance, its size and the material it is made of.
 In this work, we show how an accelerometer and a capacitive proximity sensor can be used to improve activity recognition in activities of daily living.
-For this we obtained a wrist-worn activity data logger\footnote{HedgeHog Activity Logger -- \url{http://www.ess.tu-darmstadt.de/hedgehog}} and integrate a capacitive proximity sensor unobtrusively into the wristband.
+% For this we obtained a wrist-worn activity data logger\footnote{HedgeHog Activity Logger -- \url{http://www.ess.tu-darmstadt.de/hedgehog}} and integrate a capacitive proximity sensor unobtrusively into the wristband.
+For this we obtained an open-hardware and open-source wrist-worn activity data logger \cite{hedgehog} and integrate a capacitive proximity sensor unobtrusively into the wristband.
 
@@
  		\includegraphics[trim=1cm 1cm 1cm 1cm,clip,width=\textwidth]{Images/pseudocode.pdf}
-	\caption{Measurement procedure carried out by the HedgeHog sensor: using the microcontroller's Timer0 module in counting mode, the oscillating signal generated by the capacitive sensor circuit can be measured by counting the frequency pulses over a predefined gate time (approximately 9.5ms).}
+	\caption{Measurement procedure carried out by the HedgeHog sensor: using the microcontroller's Timer0 module in counting mode, the oscillating signal generated by the capacitive sensor circuit can be measured by counting the frequency pulses over a predefined gate time of approximately 9.5ms.}
 	\label{fig:pseudocode}
@@
 The HedgeHog measures the resulting capacitance by counting the signal's edges over a gate time of approximately 9.5ms.
-During that counting phase, the HedgeHog is sent to sleep in order to reduce power consumption.
-In the following, the HedgeHog applies run-length encoding on the measured data to reduce overhead and periodically logs the data to the integrated SD card.
+During that counting phase, the microcontroller is sent to sleep in order to reduce power consumption.
+In the following, the HedgeHog applies run-length encoding on the measured data to reduce overhead and periodically logs the data to the integrated microSD card.
 
@@
 A fine-grained monitoring of such activities may help elderly or people suffering from mental diseases to maintain a healthy day/night rhythm and take action if irregularities occur.
-Figure blubb shows the modified HedgeHog activity logger that has been extended with a capacitive proximity sensor.
+Figure \ref{fig:cap_sensor} shows the modified HedgeHog activity logger that has been extended with a capacitive proximity sensor.
 The wristband has two electrodes, a sensing electrode underneath a slightly bigger shield electrode.
@@
 
-The classifier was trained with various features that were extracted from windows with the length of 1 second.
-It turned out that greater window sizes do not provide better results.
+% \begin{table}
+% 	\centering
+% 	\caption{Some details on the participants and the activities which were performed during the experiment.}
+% 	\begin{tabular}{rp{.2cm}cc}
+% 		number of participants	& & 7		&	\\
+% 		average age 			& & ?		&	\\
+% 		male-to-female ratio 	& & 6/1 	&	\\
+% 	\end{tabular}
+% 	\hspace*{2cm}
+% 	\begin{tabular}{lc}
+% 		performed activities	& 	\\ \toprule
+% 		open door				& 	\\
+% 		walking					& 	\\
+% 		sitting					& 	\\
+% 		lying					& 	\\
+% 		get things				& 	\\
+% 		making bread			& 	\\
+% 		eating					& 	\\
+% 		drinking 				& 	\\
+% 		sleeping				& 	\\
+% 	\end{tabular}
+% \end{table}
+
+The classifier was trained with basic features that were extracted from a sliding window of 1 second width.
+Our first tests have showed that greater window sizes do not provide better results.
 In order to suppress noise contained in the capacitive proximity sensing data, we applied a moving average filter with a kernel size of 10.
-The final feature set contained the arithmetic mean, min, max, median and standard variance of both accelerometer and proximity data.
+The final feature set contained the arithmetic mean, min, max, median and standard variance for each accelerometer axis and the capacitive proximity signal.
 These simple feature types represent standard features applied in activity recognition.
@@
 \begin{table}
-	\caption{Comparison of two Confusion matrices for an exemplary user. The left part includes the capacitive proximity sensor, the right part represents classification results without the sensor.}
+	\caption{Activity recognition evaluation revealing the positive impact of the capacitive proximity sensor.
+Here, we are comparing SVM classification presented as confusion matrices for an exemplary user, without the proximity data on the left, and with the proximity data on the right.}
 
@@
 	{|ccccccccc|p{0.01cm}|ccccccccc|p{0.01cm}|l}
-		\multicolumn{9}{c}{with proximity} &\multicolumn{1}{c}{}& 
-			\multicolumn{9}{c}{without proximity} &\multicolumn{1}{c}{}&  \\ \cmidrule{1-9} \cmidrule{11-19}
+		\multicolumn{9}{c}{without proximity data} &\multicolumn{1}{c}{}& 
+			\multicolumn{9}{c}{with proximity data} &\multicolumn{1}{c}{}&  \\ \cmidrule{1-9} \cmidrule{11-19}
 
@@
 		\cline{1-9} \cline{11-19} \cline{21-21}
-		\textbf{10} & 3 & 0 & 2 & 0 & 0 & 0 & 0 & 0 	&& \textbf{9} & 5 & 0 & 0 & 1 & 0 & 0 & 0 & 0 		&& a = open door	\\ 
-		2 & \textbf{259} & 4 & 2 & 4 & 2 & 4 & 0 & 0 	&& 1 & \textbf{248} & 5 & 6 & 4 & 2 & 5 & 0 & 0 	&& b = sitting		\\ 
-		0 & 5 & \textbf{213} & 2 & 2 & 1 & 8 & 0 & 4 	&& 0 & 3 & \textbf{212} & 3 & 2 & 1 & 6 & 0 & 8 	&& c = lying		\\ 
-		3 & 5 & 2 & \textbf{52} & 9 & 3 & 1 & 1 & 1 	&& 0 & 8 & 2 & \textbf{44} & 7 & 8 & 1 & 1 & 3 		&& d = get things	\\ 
-		1 & 9 & 3 & 13 & \textbf{46} & 6 & 3 & 0 & 0 	&& 1 & 7 & 1 & 11 & \textbf{37} & 9 & 2 & 0 & 10 	&& e = make bread	\\ 
-		0 & 5 & 1 & 1 & 5 & \textbf{62} & 5 & 0 & 1 	&& 0 & 3 & 2 & 3 & 8 & \textbf{58} & 5 & 0 & 1 		&& f = eating		\\ 
-		0 & 8 & 4 & 2 & 1 & 5 & \textbf{68} & 0 & 0 	&& 0 & 8 & 3 & 0 & 6 & 4 & \textbf{67} & 0 & 1 		&& g = drinking		\\
+% 		\textbf{10} & 3 & 0 & 2 & 0 & 0 & 0 & 0 & 0 	&& \textbf{9} & 5 & 0 & 0 & 1 & 0 & 0 & 0 & 0 		&& a = open door	\\ 
+% 		2 & \textbf{259} & 4 & 2 & 4 & 2 & 4 & 0 & 0 	&& 1 & \textbf{248} & 5 & 6 & 4 & 2 & 5 & 0 & 0 	&& b = sitting		\\ 
+% 		0 & 5 & \textbf{213} & 2 & 2 & 1 & 8 & 0 & 4 	&& 0 & 3 & \textbf{212} & 3 & 2 & 1 & 6 & 0 & 8 	&& c = lying		\\ 
+% 		3 & 5 & 2 & \textbf{52} & 9 & 3 & 1 & 1 & 1 	&& 0 & 8 & 2 & \textbf{44} & 7 & 8 & 1 & 1 & 3 		&& d = get things	\\ 
+% 		1 & 9 & 3 & 13 & \textbf{46} & 6 & 3 & 0 & 0 	&& 1 & 7 & 1 & 11 & \textbf{37} & 9 & 2 & 0 & 10 	&& e = make bread	\\ 
+% 		0 & 5 & 1 & 1 & 5 & \textbf{62} & 5 & 0 & 1 	&& 0 & 3 & 2 & 3 & 8 & \textbf{58} & 5 & 0 & 1 		&& f = eating		\\ 
+% 		0 & 8 & 4 & 2 & 1 & 5 & \textbf{68} & 0 & 0 	&& 0 & 8 & 3 & 0 & 6 & 4 & \textbf{67} & 0 & 1 		&& g = drinking		\\
+% 		1 & 0 & 0 & 1 & 0 & 0 & 0 & \textbf{12} & 0 	&& 1 & 0 & 0 & 1 & 0 & 0 & 0 & \textbf{13} & 0 		&& h = walking		\\
+% 		1 & 0 & 0 & 7 & 0 & 0 & 0 & 0 & \textbf{325} 	&& 0 & 1 & 2 & 10 & 6 & 1 & 1 & 0 & \textbf{312} 	&& i = sleeping		\\ 
+		\textbf{9} & 5 & 0 & 0 & 1 & 0 & 0 & 0 & 0 		&& \textbf{10} & 3 & 0 & 2 & 0 & 0 & 0 & 0 & 0 		&& a = open door	\\ 
+		1 & \textbf{248} & 5 & 6 & 4 & 2 & 5 & 0 & 0 	&& 2 & \textbf{259} & 4 & 2 & 4 & 2 & 4 & 0 & 0 	&&  b = sitting		\\ 
+		0 & 3 & \textbf{212} & 3 & 2 & 1 & 6 & 0 & 8 	&& 0 & 5 & \textbf{213} & 2 & 2 & 1 & 8 & 0 & 4 	&& c = lying		\\ 
+		0 & 8 & 2 & \textbf{44} & 7 & 8 & 1 & 1 & 3 	&& 3 & 5 & 2 & \textbf{52} & 9 & 3 & 1 & 1 & 1 		&& d = get things	\\ 
+		1 & 7 & 1 & 11 & \textbf{37} & 9 & 2 & 0 & 10 	&& 1 & 9 & 3 & 13 & \textbf{46} & 6 & 3 & 0 & 0 	&& e = make bread	\\ 
+		0 & 3 & 2 & 3 & 8 & \textbf{58} & 5 & 0 & 1 	&& 0 & 5 & 1 & 1 & 5 & \textbf{62} & 5 & 0 & 1 		&& f = eating		\\ 
+		0 & 8 & 3 & 0 & 6 & 4 & \textbf{67} & 0 & 1 	&& 0 & 8 & 4 & 2 & 1 & 5 & \textbf{68} & 0 & 0 		&& g = drinking		\\
 		1 & 0 & 0 & 1 & 0 & 0 & 0 & \textbf{12} & 0 	&& 1 & 0 & 0 & 1 & 0 & 0 & 0 & \textbf{13} & 0 		&& h = walking		\\
-		1 & 0 & 0 & 7 & 0 & 0 & 0 & 0 & \textbf{325} 	&& 0 & 1 & 2 & 10 & 6 & 1 & 1 & 0 & \textbf{312} 	&& i = sleeping		\\ 
+		0 & 1 & 2 & 10 & 6 & 1 & 1 & 0 & \textbf{312} 	&& 1 & 0 & 0 & 7 & 0 & 0 & 0 & 0 & \textbf{325} 	&& i = sleeping		\\ 
 		\cline{1-9} \cline{11-19} 
@@
 
-We would like to thank our evaluation participants from TU Darmstadt and Fraunhofer IGD.
+Anonymized for review.
+
+% We would like to thank our evaluation participants from TU Darmstadt and Fraunhofer IGD.
 

### 2012-06-22 Paper/Paper.tex
paper tex

--- 2012-06-21 Paper/Paper.tex
@@
 The shield is driven with the same potential as the sensing electrode, such that the the capacitance between the two electrodes is negligible.
-Using this shielding method the measured capacitance will only be slightly
-affected by the grounded underlying arm.
+Using this shielding method the measured capacitance will only be slightly affected by the grounded underlying arm.
 
@@
 \subsection{Evaluation Results}
+\label{sect:evaluation}
 
@@
 The recorded data for this activity shows strong correlations between all test persons.
+The confusion matrices show that the open door activity was often mixed up with the sitting activity because of the horizontal orientation of the hand.
+Due to the usage of the capacitive sensor, the recall for that class and confusion with the sitting activity improved.
 
@@
 For that class, the average f-measure could only be increased by 2.4 \%, from 82.1 to 84.5 \%.
+Considering some test persons, this class was confused with the make bread activity and could be improved by the proximity sensor.
+
 The participants were also asked to put food and dishes from a shelf and a locker on a table.
 The capacitive proximity sensor captured the proximity to the shelf and to the table.
-The overall f-measure for this activity is quite low, but improved by 6.7 \% from 50.4 \% to 57.1\%.
+The overall f-measure for this activity is rather low, but improved by 6.7 \% from 50.4 \% to 57.1\%.
+Reasons???
 
@@
 This activity showed a high improvement in the average f-measure by 10 \%, from 49.0 \% to 59.8 \%, when the data delivered by the capacitive proximity sensor is taken into account.
+The making bread class was often confused with the sitting class.
+Using the new input modality, this confusion could be improved.
 
@@
 While walking, the capacitance between the wristband and the leg increases when the wristband is located close to the body and decrease when the wristband moves away.
+There were problems in distinguishing this activity from get things that could be bettered by using the new input modality.
 
@@
 \end{figure}
-=======
-\subsection{Evaluation Results}
-\label{sect:evaluation}
 

### 2012-06-22 Paper/Paper.tex

--- 2012-06-21 Paper/Paper.tex
@@
 \begin{abstract}
-Activity recognition solely with an accelerometer is a common investigated research topic and enables the detection of basic activities like sitting, walking or standing.
+Activity recognition solely with an accelerometer is a common investigated research topic and enables the detection of basic activities like sit, walk or stand.
 Recent work in this area adds different sensing modalities to the inertial data to collect more information of the users environment and to boost activity recognition.
-This work presents a sensor prototype which consists of an accelerometer and a capacitive proximity sensor that senses the users activity based on the combined sensor values.
+We present a sensor set-up that consists of an accelerometer and a capacitive proximity sensor that senses the users activity based on the combined sensor values.
 We show that our proposed approach improves the recognition rate significantly for detecting activities of daily living (e.g.
-sitting, preparing dinner, eating, drinking, etc.) by adding capacitive proximity sensor data to the inertial data.
+open door, prepare dinner, etc.) by adding capacitive proximity sensor data to the inertial data.
 \keywords{ambient assisted living, capacitive proximity sensors, activity recognition, user context}
@@
 
-% This paper is structured as follows: We give a short insight in related work in chapter 2, followed by describing the hardware set-up used for this study in chapter 3.
+This paper is structured as follows: We give a short insight in related work in chapter 2, followed by describing the hardware set-up used for this study in chapter 3.
 The experiment and evaluation will be described in chapter 4, showing how data has been obtained and how activity recognition is being improved.
@@
 Section \ref{sect:hardware} will present the hardware based on which a wrist-worn sensor prototype was built, fusing together an inertial data logger and a capacitive sensor integrated in a wristband.
-The experimental setup with the scripted scenario of daily activities and the activity recognition evaluation results, showing the performance boost of the capacitive sensor unit, are given in Section \ref{sect:experiment}.
+The experimental setup, the scripted scenario with daily activities and the activity recognition evaluation results, showing the performance boost of the capacitive sensor unit, are given in Section \ref{sect:experiment}.
 The paper is wrapped up with a conclusions section enumerating the findings of the paper, as well as pointing out future research potential.
@@
 
-\begin{figure}
+\begin{figure}[htbp]
 	\centering
@@
 The recorded data for this activity shows strong correlations between all test persons.
+The confusion matrices show that the open door activity was often mixed up with the sitting activity because of the horizontal orientation of the hand.
+Due to the usage of the capacitive sensor, the recall for that class and confusion with the sitting activity improved.
 
-\begin{figure}
+\begin{figure}[htbp]
 	\centering
@@
 For that class, the average f-measure could only be increased by 2.4 \%, from 82.1 to 84.5 \%.
+Considering some test persons, this class was confused with the make bread activity and could be improved by the proximity sensor.
+
 The participants were also asked to put food and dishes from a shelf and a locker on a table.
 The capacitive proximity sensor captured the proximity to the shelf and to the table.
-The overall f-measure for this activity is quite low, but improved by 6.7 \% from 50.4 \% to 57.1\%.
+The overall f-measure for this activity is rather low, but improved by 6.7 \% from 50.4 \% to 57.1\%.
+Reasons???
 
-\begin{figure}
+\begin{figure}[htpb]
 	\centering
@@
 It is notable that the acceleration data does not seem to provide any characteristic patterns, while the proximity sensor indicates a table, plate, or other objects in immediate distance.
-This activity showed a high improvement in the average f-measure by 10 \%, from 49.0 \% to 59.8 \%, when the capacitive proximity sensor is taken into account.
+This activity showed a high improvement in the average f-measure by 10 \%, from 49.0 \% to 59.8 \%, when the data delivered by the capacitive proximity sensor is taken into account.
+The making bread class was often confused with the sitting class.
+Using the new input modality, this confusion could be improved.
 
-\begin{figure}
+\begin{figure}[htpb]
 	\centering
@@
 
-\begin{figure}
+\begin{figure}[htpb]
 	\centering
@@
 While walking, the capacitance between the wristband and the leg increases when the wristband is located close to the body and decrease when the wristband moves away.
+There were problems in distinguishing this activity from get things that could be bettered by using the new input modality.
 
-\begin{figure}
+\begin{figure}[htbp]
 	\centering
@@
 
-\subsection{Evaluation Results}
-\label{sect:evaluation}
-
 The sleeping activity was captured with an average f-measure of 83.4 \% that increased to 88.4 \% when using the capacitive proximity sensor.
@@
 
-\begin{table}
+\begin{table}[htbp]
+	\centering	\begin{tabular}{p{0.48cm}p{0.48cm}p{0.48cm}p{0.48cm}p{0.48cm}p{0.48cm}p{0.48cm}p{0.48cm}p{0.48cm}|p{0.48cm}p{0.48cm}p{0.48cm}p{0.48cm}p{0.48cm}p{0.48cm}p{0.48cm}p{0.48cm}p{0.48cm}|p{2.2cm}}
+a & b & c & d & e & f & g & h & i & a & b & c & d & e & f & g & h & i & $\leftarrow$ classified as\\
+\hline
+10 & 3 & 0 & 2 & 0 & 0 & 0 & 0 & 0 		& 9 & 5 & 0 & 0 & 1 & 0 & 0 & 0 & 0 & a = open door\\
+\hline
+2 & 259 & 4 & 2 & 4 & 2 & 4 & 0 & 0 	& 1 & 248 & 5 & 6 & 4 & 2 & 5 & 0 & 0 & b = sit\\
+\hline
+0 & 5 & 213 & 2 & 2 & 1 & 8 & 0 & 4 	& 0 & 3 & 212 & 3 & 2 & 1 & 6 & 0 & 8 & c = lying\\
+\hline
+3 & 5 & 2 & 52 & 9 & 3 & 1 & 1 & 1 		& 0 & 8 & 2 & 44 & 7 & 8 & 1 & 1 & 3 & d = get things\\
+\hline
+1 & 9 & 3 & 13 & 46 & 6 & 3 & 0 & 0 	& 1 & 7 & 1 & 11 & 37 & 9 & 2 & 0 & 10 & e = make bread\\
+\hline
+0 & 5 & 1 & 1 & 5 & 62 & 5 & 0 & 1 		& 0 & 3 & 2 & 3 & 8 & 58 & 5 & 0 & 1 & f = eat\\
+\hline
+0 & 8 & 4 & 2 & 1 & 5 & 68 & 0 & 0 		& 0 & 8 & 3 & 0 & 6 & 4 & 67 & 0 & 1 & g = drink\\
+\hline
+1 & 0 & 0 & 1 & 0 & 0 & 0 & 12 & 0 		& 1 & 0 & 0 & 1 & 0 & 0 & 0 & 13 & 0 & h = walk\\
+\hline
+1 & 0 & 0 & 7 & 0 & 0 & 0 & 0 & 325 	& 0 & 1 & 2 & 10 & 6 & 1 & 1 & 0 & 312 & i = sleep\\
+		\end{tabular}
 	\caption{Activity recognition evaluation revealing the positive impact of the capacitive proximity sensor.
@@
 
-Table \ref{tab:ConfusionMatrixForAnExemplaryUser}
+Table \ref{tab:confmatrix}
 

### 2012-06-22 Paper/Paper.tex
new images and lots of text in the paper - images now a bit smaller in height - lots of text for evaluation section

--- 2012-06-21 Paper/Paper.tex
@@
 \begin{abstract}
-Activity recognition solely with an accelerometer is a common investigated research topic and enables the detection of basic activities like sitting, walking or standing.
+Activity recognition with an accelerometer is a common investigated research topic and enables the detection of basic activities like sitting, walking or standing.
 Recent work in this area adds different sensing modalities to the inertial data to collect more information of the users environment and to boost activity recognition.
@@
 
+[ AAL scenario as motivation for research ]
+
 Sensing a persons activity is being researched for several years now, raising even more interest in it by the advanced technology of mobile phones, which include multiple sensors and therefore enable activity recognition with such a device \cite{brezmes2009activity}.
@@
 However, it is a great challenge to identify fine-grained activities just by using a single modality like the accelerometer.
+
 Capacitive proximity sensors on the other hand can indirectly measure the distance and nature of a grounded object within reach.
@@
 
+[ check the wording ! ]
+
 % While the RFID gloves or bracelets might not have reached user acceptance, and deploying tags in the user's home might appear cumbersome and still quite intrusive, our approach considers a single wrist-worn sensor unit consisting of an accelerometer and proximity sensor.
@@
 \subsection{Activity Data Logger}
-The HedgeHog sensor is a custom designed and built wearable data logger aiming at long-term deployments in activity recognition scenarios.
-Due to its small form-factor (30x25x20mm) and weight, this sensor, being worn at the wrist, is an unobtrusive way to record relevant motion data.
+The HedgeHog sensor \cite{hedgehog} is a custom designed and built wearable data logger aiming at long-term deployments in activity recognition scenarios.
+Due to its small form-factor (37x32x16mm) and weight, this sensor, being worn at the wrist, is an unobtrusive way to record relevant motion data.
 
@@
 Two electrodes are integrated into the wristband, one sensing electrode and one shielding electrode that aims to eliminate the influence of the arm that is a very proximate grounded object to the sensing electrode.
-The sensor draws a supply current of 1 mA at 3.3 V when active.
+The sensor draws a supply current of 1mA at 3.3V when active.
 This low power consumption qualifies the sensor for wearable proximity sensing applications.
@@
 
-\begin{figure}
+\begin{figure}[t]
 	\centering
@@
 
-Figure \ref{fig:pseudocode} illustrates the operations required for a measurement cycle.
+The operations required for a measurement cycle are illustrated in Figure \ref{fig:pseudocode}.
 The proximity sensor board generates a clock signal with varying frequency depending on the charge and discharge cycles.
@@
 
-The recorded test set contains the following activities: opening door, sitting on a couch, lying on a couch, grasping into a locker, sitting at the table, preparing a bread with marmalade, eating the bread, drinking water and sleeping.
+The recorded test set contains the following activities: opening door, sitting on a couch, lying on a couch, [kitchen preparations!] taking plate, cutlery, bread, marmalade from a shelf or a locker, sitting at the table, preparing a bread with marmalade, eating the bread, drinking water, walking and sleeping.
 Some of those activities are very hard to recognize without any further contextual information, when the data is limited to a single modality like a 3D accelerometer.
@@
 
-% \begin{table}
-% 	\centering
-% 	\caption{Some details on the participants and the activities which were performed during the experiment.}
-% 	\begin{tabular}{rp{.2cm}cc}
-% 		number of participants	& & 7		&	\\
-% 		average age 			& & ?		&	\\
-% 		male-to-female ratio 	& & 6/1 	&	\\
-% 	\end{tabular}
-% 	\hspace*{2cm}
-% 	\begin{tabular}{lc}
-% 		performed activities	& 	\\ \toprule
-% 		open door				& 	\\
-% 		walking					& 	\\
-% 		sitting					& 	\\
-% 		lying					& 	\\
-% 		get things				& 	\\
-% 		making bread			& 	\\
-% 		eating					& 	\\
-% 		drinking 				& 	\\
-% 		sleeping				& 	\\
-% 	\end{tabular}
-% \end{table}
+\begin{table}
+	\centering
+	\caption{Some details on the activities performed during the experiment and objects directly involved or nearby.}
+	\setlength{\tabcolsep}{12pt}
+	\begin{tabular}{lll}
+		\toprule
+		activities		& objects involved			& objects nearby			\\ 
+		\midrule
+		open door		& door knob 				& door  					\\
+		sitting			& chair or couch			& body, chair, couch, table	\\
+		lying			& couch						& body, couch, cushion		\\
+		get things		& plate, glass, cutlery, 	& shelf, locker, fridge, table \\ 
+						& bread, marmalade, bottle  & 							\\
+		making bread	& bread, knife, marmalade	& table, plate 				\\
+		eating			& bread						& table, plate, body		\\
+		drinking 		& bottle, glass 			& table, body				\\
+		sleeping		& bed, cushion, blanket		& body 						\\
+		walking			& 							& body						\\
+		 \bottomrule
+	\end{tabular}
+	\label{tab:activities}
+\end{table}
 
@@
 \subsection{Evaluation Results}
+\label{sect:evaluation}
 
 In the following, the performed activities will be analyzed in detail, stating out the influence of the capacitive proximity sensor on the classification result.
-In general, the usage of data provided by the new modality showed improvements in recognition rates by 2.4 - 14.4 \%.
+In general, the usage of data provided by the new modality showed improvements in recognition rates reaching from 2.4 up to 10.7\% for a single activity.
 
@@
 		\includegraphics[width=1.00\textwidth]{../Auswertung/images/tobias_1.pdf}
-	\caption{While entering the door, the wristband approaches the door knob twice, while opening and closing.
-Regarding the capacitive proximity sensor, this situation can be observed at the beginning and at the end of the activity.}
-	\label{fig:tobias_1}
+	\caption{When the participants entered the apartment, the wristband approached the door knob twice, at the time of opening and closing the door.
+This fact can be observed in the capacitive proximity data (upper plot) at the beginning and at the end of the activity, whereas the acceleration has no characteristic information (bottom plot).}
+	\label{fig:opendoor}
 \end{figure}
 
-The opening door activity has very poor recongition rates without the including the proximity sensor.
-The average f-measure could be increased from of 34.2 \% to 46.8 \%.
-A plot of the activity is given in Figure \ref{fig:tobias_1}.
-The capacitive proximity sensor shows two approaches to the door knob, one for opening the door (2 s) and one for closing the door (9 s).
-The acceleration sensor captures relevant data in the time in which the person moves into the room and the hand changes from the outer to the inner door knob (5 - 7 s).
+The opening door activity has very poor recognition rates without the data from the proximity sensor.
+The average f-measure could be increased from 35.5\% to 46.2\%.
+A plot of the activity is given in Figure \ref{fig:opendoor}.
+The capacitive proximity sensor shows two approaches to the door knob, one for opening the door (2s) and one for closing the door (9s).
+The acceleration sensor captures relevant data in the time in which the person moves into the room and the hand changes from the outer to the inner door knob (5 - 7s).
 The recorded data for this activity shows strong correlations between all test persons.
+The confusion matrices show that the ``open door'' activity was often confused with the ``sitting'' activity, because of the horizontal orientation of the hand.
+Due to the usage of the capacitive sensor, the recall for that class and confusion with the sitting activity improved.
+[wording, orientation vs little motion ?]
 
@@
 		\includegraphics[width=1.00\textwidth]{../Auswertung/images/eugen_2.pdf}
-	\caption{Example of a ``sitting'' activity in which user changed his position very frequently.
-Throughout the whole activity the values of the proximity sensor stay constant due to the close proximity to the couch's fabric.}
-	\label{fig:eugen_2}
+	\caption{Example of a ``sitting'' activity in which the user moved his hands quite frequently (bottom plot).
+Most of the time the values of the proximity sensor stay more or less constant, probably due to the hands position on the couch's fabric.
+The sharp peak in the capacitive sensor data (upper plot) occurred when the participant scratched the back of his head.
+}
+	\label{fig:sitting}
 \end{figure}
 
-After closing the door, the test persons were supposed to sit down on the couch.
+After closing the door, the participants were supposed to sit down on the couch.
 It turned out that there are great variations of the sitting posture and the corresponding hand positions.
-Many users tapped with their fingers or hands while sitting, and changed their sitting positions very frequently as shown in Figure \ref{fig:eugen_2}.
-In this case, it is obvious that the data from the acceleration sensor is very difficult to interprete as there are numerous changes in the axial orientation of the sensor.
+Many users tapped with their fingers or hands while sitting, changed their sitting positions very frequently, or were even talking and gesticulating, as shown in Figure \ref{fig:sitting}.
+In this case, it is obvious that the data from the acceleration sensor is very difficult to interpret as there are numerous changes in the axial orientation of the sensor.
 However, the capacitive proximity sensor is able to indicate when a hand is placed on the surface of the couch.
-Especially for this test person, the f-measure increased from 27.4 \% to 66.7 \%, while the average f-measure improved from 68.1 \% to 77.6 \%.
+Especially for this particular participant, the f-measure increased from 50.3\% to 60.4\%, while the average f-measure improved from 68.0\% to 74.4\%.
 
-In the following, the test persons were supposed to lay down on the couch.
-Again, there were great variations in performing this activity between all participants.
-For example, some of them crossed their hands under their head.
-For that class, the average f-measure could only be increased by 2.4 \%, from 82.1 to 84.5 \%.
-The participants were also asked to put food and dishes from a shelf and a locker on a table.
-The capacitive proximity sensor captured the proximity to the shelf and to the table.
-The overall f-measure for this activity is quite low, but improved by 6.7 \% from 50.4 \% to 57.1\%.
+In the following, the participant were instructed to lie  down on the couch.
+Again, there were great variations in how this activity was performed by the participants.
+For example, some of them crossed their hands under their head, or placed them on their stomach.
+For this class, the average f-measure could only be increased by 2.4\%, from 81.2 to 83.6\%.
+Considering some test persons, this class was confused with the ``making bread'' activity and could be improved by the proximity sensor.
+[ wording!
+]
+
+\begin{figure}
+	\centering
+		\includegraphics[width=\textwidth]{../Auswertung/images/tobias_4.pdf}
+	\caption{An example of the ``getting things'' activity, where the participants had to get food and dishes from shelves and lockers.
+The proximity sensor peaks in the beginning (9s and 19s) indicate immediate proximity to shelf, and to the locker (55-63s) in the kitchen (upper plot).
+The signal drop at the end results from the participant placing his hand on the table when she was finished.}
+	\label{fig:get_things}
+\end{figure}
+
+After that, the participants were asked to walk over to the kitchen and to put food and dishes from a shelf and a locker on the table.
+This activity involved direct interactions with various objects as well as proximity to furniture in the room (see Table \ref{tab:activities}).
+The capacitive proximity sensor was able to capture the proximity to the shelf and to the table.
+The average f-measure for this activity is rather low, but improved by 6.8\% from 53.8\% to 60.6\%.
+The worst performing participants for this activity reached an f-measure of 46.8\% without and 53.5\% with the capacitive sensor, while the best performing one reached 62.0\% and 64.5\% respectively.
+The quite low performance figures result from confusions with other activities, with a higher tendency to the ``prepare bread'' activity across all participants, most likely because there are also various objects involved.
+The capacitive sensor modality has a positive impact reducing the confusion with other activities.
 
@@
 The proximity sensor indicates the closeness to the table, while the acceleration sensor shows reoccurring movements.}
-	\label{fig:eugen_5}
+	\label{fig:prep_bread}
 \end{figure}
 
-Figure \ref{fig:eugen_5} shows an example instance of preparing a bread with marmelade.
+Figure \ref{fig:prep_bread} shows an example instance of preparing a bread with marmalade.
 It is notable that the acceleration data does not seem to provide any characteristic patterns, while the proximity sensor indicates a table, plate, or other objects in immediate distance.
-This activity showed a high improvement in the average f-measure by 10 \%, from 49.0 \% to 59.8 \%, when the capacitive proximity sensor is taken into account.
+This activity showed a high improvement in the average f-measure by 10\%, from 49.0\% to 59.8\%, where the data delivered by the capacitive proximity sensor is taken into account.
+The making bread class was often confused with the sitting class for some users, probably due to lots of motion during the sitting, as mentioned previously.
+For other users, making bread was confused with eating or drinking.
+Using the new input modality, confusion across users could be reduced.
 
@@
 	\caption{An eating person who took 5 bites from the bread. After each bite, the hand is placed on the table.}
-	\label{fig:eugen_6}
+	\label{fig:eating}
 \end{figure}
 
-Considering the eating activity, the influence of the capacitive proximity sensor is very low.
-Some of the participants ate their bread leaving their hand close to the mouth, while others moved their hand up and down putting their bread aside on the plate.
-A good example of an eating activity is shown in Figure \ref{fig:eugen_6} where the participant took a few bites from the bread.
-The average f-measure for that activity increased slightly from 80.3 \% to 83.8 \%.
+When considering the eating activity, the impact of the new capacitive proximity sensor on the classification performance is quite low, as the chosen features are able to distinguish it from other activities.
+Some of the participants ate their bread leaving their hand close to the mouth, while others moved their hand up and down putting their bread aside on the plate, which also results in the performance range from 71.2 to 88.1\% without and 77.5 to 90.6\% with the proximity data.
+A good example of an eating activity is shown in Figure \ref{fig:eating}, where the participant took a few bites from the bread while putting it down every time.
+The average f-measure for that activity increased slightly by 2.7\% (from 79.5\% to 82.2\%).
 
@@
 		\includegraphics[width=\textwidth]{../Auswertung/images/eugen_7.pdf}
-	\caption{An example of the ``drinking'' activity of a test person.
-The participant first pours some water into the glass and then takes three drams of water.
-After each dram, he returns his arm to the table which can be observed in the characteristic patterns of the proximity sensor.}
-	\label{fig:eugen_7}
+	\caption{An example of the ``drinking'' activity.
+The participant first pours some water into the glass and then takes three drinks of water.
+After each sip, he returns his arm to the table which can be observed in the characteristic patterns of the proximity sensor.}
+	\label{fig:drinking}
 \end{figure}
 
-The activity drinking water is depicted in Figure \ref{fig:eugen_7}.
-The participant took five nips from the glass, while leaving the hands lying on the table in between.
+The activity ``drinking water'' is depicted in Figure \ref{fig:drinking}.
+The participant took a few drinks from the glass, while leaving the hands lying on the table in between.
 These motions can be easily detected in the acceleration as well as proximity data.
 The accelerometer data shows that there periodic up- and down-movements while the capacitive proximity sensor delivers data that is associated to the proximity of the table.
+The average classification performance lies at 48.4\% without and 54.4\% with the proximity data taken into account, resulting in a gain of 6\%.
 
@@
 		\includegraphics[width=1.00\textwidth]{../Auswertung/images/marko_8.pdf}
-	\caption{An examplary instance of the class ``walking''.
-The acceleration sensor and the proximity sensor show periodic reoccuring patterns that are related to the pendulum-like arm movement and the proximity to the person's body during those movements.}
-	\label{fig:marko_8}
+	\caption{An exemplary instance of the class ``walking''.
+The acceleration sensor and the proximity sensor show periodic recurring patterns that are related to the pendulum-like arm movement and the proximity to the person's body during those movements.}
+	\label{fig:walking}
 \end{figure}
 
-Regarding the walking activity one can identify periodic changes in the acceleration as well as in the measured capacitance, illustrated in Figure \ref{fig:marko_8}.
+Regarding the walking activity one can identify periodic changes in the acceleration as well as in the measured capacitance, illustrated in Figure \ref{fig:walking}.
 While walking, the capacitance between the wristband and the leg increases when the wristband is located close to the body and decrease when the wristband moves away.
+There were problems in distinguishing this activity from get things that could be bettered by using the new input modality.
+The classification improvement for this activity accounts to 12.2\% boosting the average f-measure from 41.7\% without to 53.9\% with the new sensor.
+Due to the low performance results and the characteristic periodic signal shape, it would help to consider frequency domain features, as it is often applied in related work.
 
@@
 	\caption{During the sleeping activity the data from both sensors remains constant. One draw conclusions about the coverage of the arm with either couchons, blankets or the proximity to the mattress.}
-	\label{fig:eugen_9}
+	\label{fig:sleeping}
 \end{figure}
 
-\subsection{Evaluation Results}
-\label{sect:evaluation}
+The sleeping activity (an example shown in Figure \ref{fig:sleeping}) was classified with an average f-measure of 83.2\%, which increased to 86.9\% when using the proximity data.
+In this case, the capacitive proximity sensor is able to capture the surrounding cushions and blankets, as well as the body or head of the participant.
+The accelerometer data and the proximity data has larger periods with a constant signal.
+[further details?]
 
-The sleeping activity was captured with an average f-measure of 83.4 \% that increased to 88.4 \% when using the capacitive proximity sensor.
-In this situation the capacitive proximity sensor is able to capture the surrounding couchons and blankets.
-The accelerometer data and the proximity data is very constant during that activity.
-
-\begin{table}
-	\caption{Activity recognition evaluation revealing the positive impact of the capacitive proximity sensor.
-Here, we are comparing SVM classification presented as confusion matrices for an exemplary user, without the proximity data on the left, and with the proximity data on the right.}
-
+\begin{figure}
 	\setlength{\tabcolsep}{2.4pt}
 	\centering
-
 	\begin{scriptsize}
-	\begin{tabular}%{p{0.48cm}p{0.48cm}p{0.48cm}p{0.48cm}p{0.48cm}p{0.48cm}p{0.48cm}p{0.48cm}p{0.48cm}|p{0.48cm}p{0.48cm}p{0.48cm}p{0.48cm}p{0.48cm}p{0.48cm}p{0.48cm}p{0.48cm}p{0.48cm}|p{2.2cm}}
-	{|ccccccccc|p{0.01cm}|ccccccccc|p{0.01cm}|l}
+	\begin{tabular}{|ccccccccc|p{0.01cm}|ccccccccc|p{0.01cm}|l}
+% 		\toprule
 		\multicolumn{9}{c}{without proximity data} &\multicolumn{1}{c}{}& 
@@
 		\cline{1-9} \cline{11-19} \cline{21-21}
-% 		\textbf{10} & 3 & 0 & 2 & 0 & 0 & 0 & 0 & 0 	&& \textbf{9} & 5 & 0 & 0 & 1 & 0 & 0 & 0 & 0 		&& a = open door	\\ 
-% 		2 & \textbf{259} & 4 & 2 & 4 & 2 & 4 & 0 & 0 	&& 1 & \textbf{248} & 5 & 6 & 4 & 2 & 5 & 0 & 0 	&& b = sitting		\\ 
-% 		0 & 5 & \textbf{213} & 2 & 2 & 1 & 8 & 0 & 4 	&& 0 & 3 & \textbf{212} & 3 & 2 & 1 & 6 & 0 & 8 	&& c = lying		\\ 
-% 		3 & 5 & 2 & \textbf{52} & 9 & 3 & 1 & 1 & 1 	&& 0 & 8 & 2 & \textbf{44} & 7 & 8 & 1 & 1 & 3 		&& d = get things	\\ 
-% 		1 & 9 & 3 & 13 & \textbf{46} & 6 & 3 & 0 & 0 	&& 1 & 7 & 1 & 11 & \textbf{37} & 9 & 2 & 0 & 10 	&& e = make bread	\\ 
-% 		0 & 5 & 1 & 1 & 5 & \textbf{62} & 5 & 0 & 1 	&& 0 & 3 & 2 & 3 & 8 & \textbf{58} & 5 & 0 & 1 		&& f = eating		\\ 
-% 		0 & 8 & 4 & 2 & 1 & 5 & \textbf{68} & 0 & 0 	&& 0 & 8 & 3 & 0 & 6 & 4 & \textbf{67} & 0 & 1 		&& g = drinking		\\
+% 		\textbf{9} & 5 & 0 & 0 & 1 & 0 & 0 & 0 & 0 		&& \textbf{10} & 3 & 0 & 2 & 0 & 0 & 0 & 0 & 0 		&& a = open door	\\ 
+% 		1 & \textbf{248} & 5 & 6 & 4 & 2 & 5 & 0 & 0 	&& 2 & \textbf{259} & 4 & 2 & 4 & 2 & 4 & 0 & 0 	&& b = sitting		\\ 
+% 		0 & 3 & \textbf{212} & 3 & 2 & 1 & 6 & 0 & 8 	&& 0 & 5 & \textbf{213} & 2 & 2 & 1 & 8 & 0 & 4 	&& c = lying		\\ 
+% 		0 & 8 & 2 & \textbf{44} & 7 & 8 & 1 & 1 & 3 	&& 3 & 5 & 2 & \textbf{52} & 9 & 3 & 1 & 1 & 1 		&& d = get things	\\ 
+% 		1 & 7 & 1 & 11 & \textbf{37} & 9 & 2 & 0 & 10 	&& 1 & 9 & 3 & 13 & \textbf{46} & 6 & 3 & 0 & 0 	&& e = make bread	\\ 
+% 		0 & 3 & 2 & 3 & 8 & \textbf{58} & 5 & 0 & 1 	&& 0 & 5 & 1 & 1 & 5 & \textbf{62} & 5 & 0 & 1 		&& f = eating		\\ 
+% 		0 & 8 & 3 & 0 & 6 & 4 & \textbf{67} & 0 & 1 	&& 0 & 8 & 4 & 2 & 1 & 5 & \textbf{68} & 0 & 0 		&& g = drinking		\\
 % 		1 & 0 & 0 & 1 & 0 & 0 & 0 & \textbf{12} & 0 	&& 1 & 0 & 0 & 1 & 0 & 0 & 0 & \textbf{13} & 0 		&& h = walking		\\
-% 		1 & 0 & 0 & 7 & 0 & 0 & 0 & 0 & \textbf{325} 	&& 0 & 1 & 2 & 10 & 6 & 1 & 1 & 0 & \textbf{312} 	&& i = sleeping		\\ 
-		\textbf{9} & 5 & 0 & 0 & 1 & 0 & 0 & 0 & 0 		&& \textbf{10} & 3 & 0 & 2 & 0 & 0 & 0 & 0 & 0 		&& a = open door	\\ 
-		1 & \textbf{248} & 5 & 6 & 4 & 2 & 5 & 0 & 0 	&& 2 & \textbf{259} & 4 & 2 & 4 & 2 & 4 & 0 & 0 	&&  b = sitting		\\ 
-		0 & 3 & \textbf{212} & 3 & 2 & 1 & 6 & 0 & 8 	&& 0 & 5 & \textbf{213} & 2 & 2 & 1 & 8 & 0 & 4 	&& c = lying		\\ 
-		0 & 8 & 2 & \textbf{44} & 7 & 8 & 1 & 1 & 3 	&& 3 & 5 & 2 & \textbf{52} & 9 & 3 & 1 & 1 & 1 		&& d = get things	\\ 
-		1 & 7 & 1 & 11 & \textbf{37} & 9 & 2 & 0 & 10 	&& 1 & 9 & 3 & 13 & \textbf{46} & 6 & 3 & 0 & 0 	&& e = make bread	\\ 
-		0 & 3 & 2 & 3 & 8 & \textbf{58} & 5 & 0 & 1 	&& 0 & 5 & 1 & 1 & 5 & \textbf{62} & 5 & 0 & 1 		&& f = eating		\\ 
-		0 & 8 & 3 & 0 & 6 & 4 & \textbf{67} & 0 & 1 	&& 0 & 8 & 4 & 2 & 1 & 5 & \textbf{68} & 0 & 0 		&& g = drinking		\\
-		1 & 0 & 0 & 1 & 0 & 0 & 0 & \textbf{12} & 0 	&& 1 & 0 & 0 & 1 & 0 & 0 & 0 & \textbf{13} & 0 		&& h = walking		\\
-		0 & 1 & 2 & 10 & 6 & 1 & 1 & 0 & \textbf{312} 	&& 1 & 0 & 0 & 7 & 0 & 0 & 0 & 0 & \textbf{325} 	&& i = sleeping		\\ 
+% 		0 & 1 & 2 & 10 & 6 & 1 & 1 & 0 & \textbf{312} 	&& 1 & 0 & 0 & 7 & 0 & 0 & 0 & 0 & \textbf{325} 	&& i = sleeping		\\ 
+		\textbf{11} & 0 & 5 & 4 & 1 & 0 & 0 & 0 & 0 	&& \textbf{17} & 0 & 0 & 2 & 0 & 0 & 0 & 0 & 0		&& a = open door    \\
+		0 & \textbf{136} & 8 & 1 & 11 & 0 & 0 & 0 & 0	&& 0 & \textbf{140} & 7 & 2 & 3 & 0 & 0 & 0 & 0		&& b = sitting      \\
+		4 & 20 & \textbf{151} & 1 & 7 & 2 & 1 & 1 & 0	&& 0 & 23 & \textbf{149} & 5 & 7 & 3 & 2 & 1 & 0	&& c = lying        \\
+		3 & 6 & 5 & \textbf{116} & 21 & 9 & 4 & 4 & 2	&& 1 & 1 & 9 & \textbf{129} & 25 & 4 & 2 & 4 & 1	&& d = get things   \\
+		0 & 5 & 2 & 5 & \textbf{81} & 11 & 2 & 0 & 0	&& 0 & 3 & 1 & 15 & \textbf{83} & 6 & 1 & 0 & 1		&& e = make bread   \\
+		0 & 0 & 0 & 0 & 9 & \textbf{167} & 19 & 0 & 3	&& 0 & 0 & 0 & 4 & 8 & \textbf{158} & 26 & 0 & 3	&& f = eating       \\
+		1 & 2 & 3 & 4 & 8 & 26 & \textbf{89} & 0 & 5	&& 0 & 0 & 8 & 5 & 5 & 29 & \textbf{92} & 1 & 4		&& g = drinking     \\
+		0 & 0 & 1 & 17 & 3 & 0 & 0 & \textbf{2} & 0		&& 0 & 1 & 2 & 10 & 3 & 0 & 0 & \textbf{9} & 0		&& h = walking      \\
+		0 & 0 & 0 & 2 & 0 & 8 & 2 & 0 & \textbf{78}		&& 0 & 0 & 0 & 2 & 0 & 6 & 1 & 0 & \textbf{81}		&& i = sleeping     \\ 
 		\cline{1-9} \cline{11-19} 
+% 		\bottomrule
 	\end{tabular}
@@
 
-	\label{tab:ConfusionMatrixForAnExemplaryUser}
-\end{table}
+	\caption{Activity recognition evaluation revealing the positive impact of the capacitive proximity sensor.
+Here, we are comparing SVM classification presented as confusion matrices for an exemplary user, without the proximity data on the left, and with the proximity data on the right.
+Note: reject class is not included in the confusion matrix.}
 
-Table \ref{tab:ConfusionMatrixForAnExemplaryUser}
+	\label{fig:ConfusionMatrixForAnExemplaryUser}
+\end{figure}
+
+Figure \ref{fig:ConfusionMatrixForAnExemplaryUser} depicts an example for the SVM classification results for one of the participants.
 

### 2012-06-22 Paper/Paper.tex
added confusion matrix

@@
 % 		0 & 1 & 2 & 10 & 6 & 1 & 1 & 0 & \textbf{312} 	&& 1 & 0 & 0 & 7 & 0 & 0 & 0 & 0 & \textbf{325} 	&& i = sleeping		\\ 
-		\textbf{11} & 0 & 5 & 4 & 1 & 0 & 0 & 0 & 0 	&& \textbf{17} & 0 & 0 & 2 & 0 & 0 & 0 & 0 & 0		&& a = open door    \\
-		0 & \textbf{136} & 8 & 1 & 11 & 0 & 0 & 0 & 0	&& 0 & \textbf{140} & 7 & 2 & 3 & 0 & 0 & 0 & 0		&& b = sitting      \\
-		4 & 20 & \textbf{151} & 1 & 7 & 2 & 1 & 1 & 0	&& 0 & 23 & \textbf{149} & 5 & 7 & 3 & 2 & 1 & 0	&& c = lying        \\
-		3 & 6 & 5 & \textbf{116} & 21 & 9 & 4 & 4 & 2	&& 1 & 1 & 9 & \textbf{129} & 25 & 4 & 2 & 4 & 1	&& d = get things   \\
-		0 & 5 & 2 & 5 & \textbf{81} & 11 & 2 & 0 & 0	&& 0 & 3 & 1 & 15 & \textbf{83} & 6 & 1 & 0 & 1		&& e = make bread   \\
-		0 & 0 & 0 & 0 & 9 & \textbf{167} & 19 & 0 & 3	&& 0 & 0 & 0 & 4 & 8 & \textbf{158} & 26 & 0 & 3	&& f = eating       \\
-		1 & 2 & 3 & 4 & 8 & 26 & \textbf{89} & 0 & 5	&& 0 & 0 & 8 & 5 & 5 & 29 & \textbf{92} & 1 & 4		&& g = drinking     \\
-		0 & 0 & 1 & 17 & 3 & 0 & 0 & \textbf{2} & 0		&& 0 & 1 & 2 & 10 & 3 & 0 & 0 & \textbf{9} & 0		&& h = walking      \\
-		0 & 0 & 0 & 2 & 0 & 8 & 2 & 0 & \textbf{78}		&& 0 & 0 & 0 & 2 & 0 & 6 & 1 & 0 & \textbf{81}		&& i = sleeping     \\ 
+% 	
+%		\textbf{11} & 0 & 5 & 4 & 1 & 0 & 0 & 0 & 0 	&& \textbf{17} & 0 & 0 & 2 & 0 & 0 & 0 & 0 & 0		&& a = open door    \\
+%		0 & \textbf{136} & 8 & 1 & 11 & 0 & 0 & 0 & 0	&& 0 & \textbf{140} & 7 & 2 & 3 & 0 & 0 & 0 & 0		&& b = sitting      \\
+%		4 & 20 & \textbf{151} & 1 & 7 & 2 & 1 & 1 & 0	&& 0 & 23 & \textbf{149} & 5 & 7 & 3 & 2 & 1 & 0	&& c = lying        \\
+%		3 & 6 & 5 & \textbf{116} & 21 & 9 & 4 & 4 & 2	&& 1 & 1 & 9 & \textbf{129} & 25 & 4 & 2 & 4 & 1	&& d = get things   \\
+%		0 & 5 & 2 & 5 & \textbf{81} & 11 & 2 & 0 & 0	&& 0 & 3 & 1 & 15 & \textbf{83} & 6 & 1 & 0 & 1		&& e = make bread   \\
+%		0 & 0 & 0 & 0 & 9 & \textbf{167} & 19 & 0 & 3	&& 0 & 0 & 0 & 4 & 8 & \textbf{158} & 26 & 0 & 3	&& f = eating       \\
+%		1 & 2 & 3 & 4 & 8 & 26 & \textbf{89} & 0 & 5	&& 0 & 0 & 8 & 5 & 5 & 29 & \textbf{92} & 1 & 4		&& g = drinking     \\
+%		0 & 0 & 1 & 17 & 3 & 0 & 0 & \textbf{2} & 0		&& 0 & 1 & 2 & 10 & 3 & 0 & 0 & \textbf{9} & 0		&& h = walking      \\
+%		0 & 0 & 0 & 2 & 0 & 8 & 2 & 0 & \textbf{78}		&& 0 & 0 & 0 & 2 & 0 & 6 & 1 & 0 & \textbf{81}		&& i = sleeping     \\
+
+		\textbf{8} & 0 & 0 & 0 & 0 & 0 & 0 & 2 & 1				&& \textbf{11} & 0 & 0 & 2 & 1 & 0 & 0 & 0 & 1 		&& a = open door	\\ 
+		0 & \textbf{140} & 0 & 5 & 2 & 0 & 0 & 0 & 0  		&& 0 & \textbf{140} & 0 & 5 & 2 & 0 & 0 & 0 & 0 	&&  b = sitting		\\ 
+		0 & 0 & \textbf{294} & 1 & 9 & 10 & 1 & 1 & 0 	 	&& 0 & 0 & \textbf{309} & 1 & 0 & 4 & 0 & 0 & 1 	&& c = lying		\\ 
+		1 & 7 & 1 & \textbf{50} & 17 & 9 & 1 & 4 & 1	 		&& 2 & 6 & 0 & \textbf{67} & 14 & 9 & 1 & 2 & 1 	&& d = get things	\\ 
+		0 & 2 & 7 & 7 & \textbf{102} & 42 & 5 & 0 & 2			&& 1 & 1 & 0 & 13 & \textbf{120} & 24 & 7 & 1 & 1 && e = make bread	\\ 
+		0 & 0 & 3 & 5 & 21 & \textbf{288} & 4 & 0 & 0 	 	&& 0 & 0 & 2 & 1 & 19 & \textbf{300} & 3 & 0 & 0 	&& f = eating		\\ 
+		0 & 5 & 11 & 1 & 29 & 40 & \textbf{18} & 0 & 0	 	&& 0 & 4 & 4 & 0 & 26 & 57 & \textbf{16} & 0 & 0	&& g = drinking		\\
+		2 & 0 & 0 & 2 & 0 & 0 & 0 & \textbf{11} & 0	 			&& 2 & 0 & 0 & 0 & 0 & 0 & 0 & \textbf{14} & 0 		&& h = walking		\\
+		0 & 2 & 2 & 0 & 2 & 0 & 0 & 0 & \textbf{271}			&& 0 & 0 & 0 & 0 & 1 & 0 & 0 & 0 & \textbf{276} 		&& i = sleeping		\\
+
 		\cline{1-9} \cline{11-19} 

### 2012-06-22 Paper/Paper.tex
changed tex

@@
 sitting, preparing dinner, eating, drinking, etc.) by adding capacitive proximity sensor data to the inertial data.
-\keywords{ambient assisted living, capacitive proximity sensors, activity recognition, user context}
+\keywords{activity recognition, capacitive proximity sensors, ambient assisted living, user context}
 \end{abstract}
@@
 
-[ AAL scenario as motivation for research ]
+Persons affected from physical and mental restrictions and their care givers can profit significantly from unobtrusive activity monitoring solutions.
+For example, formal care givers could equip a person with a wearable activity monitoring solution that evaluates the course of a disease or the influence of an adapted medication.
+This scenario is especially relevant for persons suffering from dementia who have limitations in organizing their daily activities.
+A simple wearable activity monitoring solution could analyze activities performed in daily life, such as drinking, eating and sleeping habits.
 
-Sensing a persons activity is being researched for several years now, raising even more interest in it by the advanced technology of mobile phones, which include multiple sensors and therefore enable activity recognition with such a device \cite{brezmes2009activity}.
+Sensing a person's activity is an active research topic with a raising interest due to the advancement in mobile phone technology.
+These devices include multiple sensors and therefore enable the recognition of daily activities \cite{brezmes2009activity}.
 Current wearable activity recognition systems are able to unobtrusively capture and recognize a person's activities throughout the whole day.
-These systems often rely on inertial sensor data that is captured by wearable sensors embedded in a mobile device \cite{brezmes2009activity} or attached to the body \cite{Ravi2005} to identify the activities performed.
+These systems often rely on inertial sensor data that is captured by wearable sensors embedded in a mobile device \cite{brezmes2009activity} or attached to the body \cite{Ravi2005}.
 Usually, single sensor modalities are used or duplicated to detect the activities.
@@
 This means that the measurement result depends on the object's distance, its size and the material it is made of.
-In this work, we show how an accelerometer and a capacitive proximity sensor can be used to improve activity recognition in activities of daily living.
+In this work, we show that an accelerometer and a capacitive proximity sensor can be used to improve activity recognition in activities of daily living.
 % For this we obtained a wrist-worn activity data logger\footnote{HedgeHog Activity Logger -- \url{http://www.ess.tu-darmstadt.de/hedgehog}} and integrate a capacitive proximity sensor unobtrusively into the wristband.
-For this we obtained an open-hardware and open-source wrist-worn activity data logger \cite{hedgehog} and integrate a capacitive proximity sensor unobtrusively into the wristband.
+We obtained an open-hardware and open-source wrist-worn activity data logger \cite{hedgehog} and integrated a capacitive proximity sensor invisibly into the wristband.
 
 There are several intuitive examples for which a combination of accelerometer-based activity recognition with a capacitive proximity sensor reveals its strength.
-For example, it may be conducted which material is placed underneath the hand.
+For example, it may be conducted which material is placed underneath the wristband.
 The capacitive proximity sensor would return a different measurement result for a hand placed on a couch covered with fabric than a hand placed a wooden table.
-Moreover, the approximate distance to objects can be exploited to identify activities like grasping into a locker or a refrigerator to prepare food.
+Moreover, the approximate distance from the wristband to objects can be exploited to identify activities like grasping into a locker or a refrigerator to prepare food.
 
-A wristband worn capacitive proximity sensor requires a shield that eliminates the influence of the grounded arm directly underneath the sensor.
+A wrist-worn capacitive proximity sensor requires a shield that eliminates the influence of the grounded arm directly underneath the sensor.
 Using this setup, we can detect the proximity to a grounded object in the environments for distances up to 20 cm.
@@
 
-The remainder of this paper is  is structured as follows: First, Section \ref{sect:related} is dedicated to related work work in activity recognition, in particular considering approaches with multiple sensor modalities.
-Section \ref{sect:hardware} will present the hardware based on which a wrist-worn sensor prototype was built, fusing together an inertial data logger and a capacitive sensor integrated in a wristband.
-The experimental setup, the scripted scenario with daily activities and the activity recognition evaluation results, showing the performance boost of the capacitive sensor unit, are given in Section \ref{sect:experiment}.
-The paper is wrapped up with a conclusions section enumerating the findings of the paper, as well as pointing out future research potential.
+The remainder of this paper is structured as follows: First, Section \ref{sect:related} is dedicated to related work work in activity recognition, in particular considering approaches with multiple sensing modalities.
+Section \ref{sect:hardware} presents the hardware based on which a wrist-worn sensor prototype was built, fusing an inertial data logger and a capacitive sensor integrated into a wristband.
+The experimental setup, the scenario with daily activities and the activity recognition evaluation results, showing the performance boost of the capacitive sensor unit, are given in Section \ref{sect:experiment}.
+The paper is wrapped up with a conclusion enumerating the findings of the evaluation, as well as pointing out future research potential.
 
@@
 Activity recognition research relying on wearable sensors mostly considers inertial data from the participants body to infer performed activities, such as in the works of \cite{Ravi2005,Bao2004,Srinivasan2010,Amft2005}.
-The acceleration data is often augmented with data from sensors such as gyroscopes \cite{Holleczek_2010}, magnetometers \cite{Altun_2010}, ambient light \cite{Borazio2012} or ambient and skin temperature \cite{Krause_2003}, aiming at extracting more motion and environmental context of the user.
-Researchers in \cite{wyss2010recognition}, for example, use heart rate information additionally to the accelerometer to detect military related activities like lifting and lowering loads or even digging.
-In \cite{Ward_2006}, workshop assembly activities were detected by augmenting acceleration sensors with microphones.
+The acceleration data is often augmented with data from sensors such as gyroscopes \cite{Holleczek_2010}, magnetometers \cite{Altun_2010}, ambient light \cite{Borazio2012} or ambient and skin temperature \cite{Krause_2003}, aiming to extract a more detailed environmental user context.
+Researchers in \cite{wyss2010recognition} use heart rate information as an addition to the accelerometer data to detect military related activities like lifting and lowering loads or even digging.
+In \cite{Ward_2006}, workshop assembly activities are detected by augmenting acceleration sensors with microphones.
 
-The works of \cite{Fishkin_2005} and \cite{patterson2005fga} show that detecting objects the user touches or uses can be very helpful for activity recognition.
+The works of \cite{Fishkin_2005} and \cite{patterson2005fga} show that detecting touched and used objects can be very helpful for activity recognition.
 By using RFID readers that are embedded in gloves or bracelets at the wrist and RFID tags attached to various objects of interest, one can detect the object grasped and used by the user, thus aiding the activity recognition in various application scenarios, such as activities of daily living \cite{phealth:maja} and \cite{Philipose_2004}, activity tracking in car manufacturing \cite{Stiefmeier08}, or household and gardening activities \cite{berlin_laerhoven_tei_2010}.
 
-Our approach to enhance the inertial data from a wearable sensor is quite similar to the RFID scenarios just mentioned, as it also relies on a single wearable sensor and thus is aiming at an unobtrusive deployment.
-The main difference lies in the fact that we do not need accurate detection of tagged objects, but are only interested in the proximity to various objects which, as we expect, could boost activity recognition.
-
-[ check the wording ! ]
+Our approach to enhance the inertial data from a wearable sensor is comparable to the RFID scenarios just mentioned, as it also relies on a single wearable sensor and an unobtrusive deployment.
+The main difference lies in the fact that we do not consider an accurate detection of tagged objects, but the proximity to various objects and the environment.
 
@@
 
-In the field of capacitive proximity sensing, three different measurement modes shown in Figure \ref{fig:modes} were identified by Smith et al.
+In the field of capacitive proximity sensing, three different measurement modes (shown in Figure \ref{fig:modes}) were identified by Smith et al.
 \cite{Smith1999}: transmit mode, shunt mode and loading mode.
@@
 In previous works, capacitive proximity sensors were applied in various fields of human-computer-interaction.
-For example, Wimmer et al.
+Wimmer et al.
 and Grosse-Puppendahl et al.
@@
 \label{sect:hardware}
+
 This section presents the two components of the required hardware, the wrist-worn activity data logger tailored to capture acceleration data, and the capacitive proximity sensor used for distance measurements.
@@
 \subsection{Activity Data Logger}
-The HedgeHog sensor \cite{hedgehog} is a custom designed and built wearable data logger aiming at long-term deployments in activity recognition scenarios.
-Due to its small form-factor (37x32x16mm) and weight, this sensor, being worn at the wrist, is an unobtrusive way to record relevant motion data.
+
+The HedgeHog sensor \cite{hedgehog} is a custom designed wearable data logger aiming at long-term deployments in activity recognition scenarios.
+Due to its small form-factor (37x32x16mm) and weight, this wrist-worn sensor is an unobtrusive way to record relevant motion data.
 
@@
 The HedgeHog can be extended with further sensors tailoring different application scenarios.
-For our scenario, we have added a capacitive proximity sensor, which is described in detail in the next section.
+For our scenario, we have added a capacitive proximity sensor that is described in detail in the next section.
 
@@
 The capacitive proximity sensor performs measurements in loading mode.
-Two electrodes are integrated into the wristband, one sensing electrode and one shielding electrode that aims to eliminate the influence of the arm that is a very proximate grounded object to the sensing electrode.
-The sensor draws a supply current of 1mA at 3.3V when active.
-This low power consumption qualifies the sensor for wearable proximity sensing applications.
+Two electrodes are integrated into the wristband, one sensing electrode and one shielding electrode that aims to eliminate the influence of the underlying skin.
+Without shielding no meaningful measurements could be performed, since the arm's skin is a very proximate grounded object to the sensing electrode.
+The sensor draws a supply current of 1mA at 3.3V when active and qualifies it for wearable proximity sensing applications.
+In the following a virtual capacitor denotes the capacitance between the sensing electrode and the environment.
+The capacitance of the sensing electrode to environmental objects increases with closer distances.
 
@@
 
-The experiment setup aims to depict a typical scenario in daily life.
+The experiment setup aims to depict a typical scenario of person in daily life.
 Especially in the field of Ambient Assisted Living (AAL), it is desired to monitor activities like drinking, preparing lunch and sleeping.
@@
 
-The recorded test set contains the following activities: opening door, sitting on a couch, lying on a couch, [kitchen preparations!] taking plate, cutlery, bread, marmalade from a shelf or a locker, sitting at the table, preparing a bread with marmalade, eating the bread, drinking water, walking and sleeping.
-Some of those activities are very hard to recognize without any further contextual information, when the data is limited to a single modality like a 3D accelerometer.
+The recorded test set contains the following activities: opening door, sitting on a couch, lying on a couch, putting kitchen equipments from a shelf and out of a locker, preparing a bread with marmalade, eating the bread, pouring and drinking water, walking and sleeping.
+The relations of these activities to environmental objects are given in Table \ref{tab:activities}.
+Some of those activities are very hard to recognize when the data is limited to a single modality like a 3D accelerometer.
 For example, sitting at the table and sitting on a couch are very similar activities.
 We aim to show that the data basis can be significantly improved by the additional input modality.
-
-In order to evaluate if capacitive proximity sensors in wrist-bands can enhance the performance of activity recognition, we have conducted an evaluation with 7 test persons.
-All test persons received a basic plot with the activities they were supposed to perform.
-They were not given any instructions about the way they are supposed to perform the activities.
-After manual labeling, we used this test-set as ground truth and performed a 4-fold cross-validation on an support-vector machine (SVM) classifier on each user, once with and once without including the data of the capacitive proximity sensor into the feature set.
-We used an SVM classifier because of its high relevance in activity recognition and its fast performance.
 
@@
 
+In order to evaluate if capacitive proximity sensors in wrist-bands can enhance the performance of activity recognition, we have conducted an evaluation with 7 test persons.
+All test persons received a basic plot with the activities they were supposed to perform.
+They were not given any instructions about the way they are supposed to perform the activities.
+After manual labeling, we used this test-set as ground truth and performed a 4-fold cross-validation on an support-vector machine (SVM) classifier on each user.
+The cross-validation was performed once with and once without including the data of the capacitive proximity sensor into the feature set.
+We chose an SVM classifier because of its high relevance in activity recognition and its fast performance.
+
 The classifier was trained with basic features that were extracted from a sliding window of 1 second width.
-Our first tests have showed that greater window sizes do not provide better results.
+Our first tests have shown that greater window sizes do not provide better classification results.
 In order to suppress noise contained in the capacitive proximity sensing data, we applied a moving average filter with a kernel size of 10.
@@
 These simple feature types represent standard features applied in activity recognition.
+Since we aim to show an improvement using the new modality, the selection of features and classifiers does not represent the primary focus of this paper.
 

### 2012-06-22 Paper/Paper.tex
prefinal version of the paper

@@
 This work presents a sensor prototype which consists of an accelerometer and a capacitive proximity sensor that senses the users activity based on the combined sensor values.
-We show that our proposed approach improves the recognition rate significantly for detecting activities of daily living (e.g.
+We show that our proposed approach significantly improves the recognition rate for detecting activities of daily living (e.g.
 sitting, preparing dinner, eating, drinking, etc.) by adding capacitive proximity sensor data to the inertial data.
@@
 For example, formal care givers could equip a person with a wearable activity monitoring solution that evaluates the course of a disease or the influence of an adapted medication.
-This scenario is especially relevant for persons suffering from dementia who have limitations in organizing their daily activities.
+This scenario is especially relevant for people suffering from dementia who have limitations in organizing their daily activities.
 A simple wearable activity monitoring solution could analyze activities performed in daily life, such as drinking, eating and sleeping habits.
@@
 In this work, we show that an accelerometer and a capacitive proximity sensor can be used to improve activity recognition in activities of daily living.
-% For this we obtained a wrist-worn activity data logger\footnote{HedgeHog Activity Logger -- \url{http://www.ess.tu-darmstadt.de/hedgehog}} and integrate a capacitive proximity sensor unobtrusively into the wristband.
 We obtained an open-hardware and open-source wrist-worn activity data logger \cite{hedgehog} and integrated a capacitive proximity sensor invisibly into the wristband.
@@
 Moreover, the approximate distance from the wristband to objects can be exploited to identify activities like grasping into a locker or a refrigerator to prepare food.
-
-A wrist-worn capacitive proximity sensor requires a shield that eliminates the influence of the grounded arm directly underneath the sensor.
-Using this setup, we can detect the proximity to a grounded object in the environments for distances up to 20 cm.
-Especially for mobile devices, it is required that the sensor only draws a very small amount of power.
-Thus, other proximity sensing input modalities like ultrasound or optical measurements are not applicable for this type of mobile application.
-
-% This paper is structured as follows: We give a short insight in related work in chapter 2, followed by describing the hardware set-up used for this study in chapter 3.
-The experiment and evaluation will be described in chapter 4, showing how data has been obtained and how activity recognition is being improved.
-This work is concluded in chapter 5 with an outlook to future work.
 
@@
 The acceleration data is often augmented with data from sensors such as gyroscopes \cite{Holleczek_2010}, magnetometers \cite{Altun_2010}, ambient light \cite{Borazio2012} or ambient and skin temperature \cite{Krause_2003}, aiming to extract a more detailed environmental user context.
-Researchers in \cite{wyss2010recognition} use heart rate information as an addition to the accelerometer data to detect military related activities like lifting and lowering loads or even digging.
+In \cite{wyss2010recognition}, the authors use heart rate information as an addition to the accelerometer data to detect activities like lifting and lowering loads or even digging.
 In \cite{Ward_2006}, workshop assembly activities are detected by augmenting acceleration sensors with microphones.
 
-The works of \cite{Fishkin_2005} and \cite{patterson2005fga} show that detecting touched and used objects can be very helpful for activity recognition.
+The works of Fishkin et al.
+\cite{Fishkin_2005} and Patterson et al.
+\cite{patterson2005fga} show that detecting touched and used objects can be very helpful for activity recognition.
 By using RFID readers that are embedded in gloves or bracelets at the wrist and RFID tags attached to various objects of interest, one can detect the object grasped and used by the user, thus aiding the activity recognition in various application scenarios, such as activities of daily living \cite{phealth:maja} and \cite{Philipose_2004}, activity tracking in car manufacturing \cite{Stiefmeier08}, or household and gardening activities \cite{berlin_laerhoven_tei_2010}.
@@
 \cite{berlin_laerhoven_tei_2010} have also fused and benchmarked a wrist-worn accelerometer sensor node with an RFID bracelet to be able to detect objects a human is interacting with, aiming both at household as well as gardening activities.
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 
@@
 
-This section presents the two components of the required hardware, the wrist-worn activity data logger tailored to capture acceleration data, and the capacitive proximity sensor used for distance measurements.
+This section presents the two components of our hardware prototype, the wrist-worn activity data logger tailored to capture acceleration data, and the capacitive proximity sensor used for distance measurements.
 
@@
 
-The sensor node itself is built around the low-power Microchip microcontroller (PIC18F46J50) featuring an accelerometer sensor (ADXL345) to capture human motion, light and ambient temperature sensors and a microSD flash card for locally storing the sensor data.
+%  (PIC18F46J50) (ADXL345)
+
+The sensor node itself is built around the low-power Microchip microcontroller featuring an accelerometer sensor to capture human motion, light and ambient temperature sensors and a microSD flash card for locally storing the sensor data.
 The sensor is powered by a $200mAh$ lithium polymer battery, which allows for two weeks of continuous recording on a single battery charge.
@@
 setting the sensitivity of the accelerometer), to access the stored sensor data, and to recharge the battery.
-A plastic case nicely packages and protects the sensor to be worn at the wrist (as shown in Fig.
-\ref{fig:sensornode}).
+A plastic case nicely packages and protects the sensor to be worn at the wrist (Figure \ref{fig:sensornode}).
 
@@
 	\includegraphics[width=\textwidth]{Images/hardware_sensor_2.jpg}
-	\caption{The inertial data logger featuring a low-power microcontroller, a 3 axis accelerometer, a microSD flash card for storing the sensor data and a USB connector for accessing the stored data (on the right) is being powered by a small lithium polymer battery and is packaged into a plastic case to be worn at the wrist (a version with a OLED display on the left).}
+	\caption{The inertial data logger featuring a low-power microcontroller, a 3 axis accelerometer, a microSD flash card for storing the sensor data and a USB connector for accessing the data (on the right) is powered by a small lithium polymer battery and is packaged into a plastic case to be worn at the wrist (a version with a OLED display).}
 	\label{fig:sensornode}
@@
 
-The 3D accelerometer sensor is being sampled at $100Hz$, resulting in 10ms equidistant measurements.
+The 3D accelerometer sensor is being sampled at 100Hz, resulting in 10ms equidistant measurements.
 For efficiency reasons, the sensor data is run-length encoded before being stored locally to the microSD card.
@@
 
+A wrist-worn capacitive proximity sensor requires a shield that eliminates the influence of the grounded arm directly underneath the sensor.
+Using this setup, we can detect the proximity to a grounded object in the environments for distances up to 20 cm.
+Especially for mobile devices, it is required that the sensor only draws a very small amount of power.
+Thus, other proximity sensing input modalities like ultrasound or optical measurements are not applicable for this type of mobile application.
+
 The capacitive proximity sensor performs measurements in loading mode.
+Two electrodes are integrated into the wristband, one sensing electrode and one shielding electrode.
+The sensor draws a supply current of 1mA at 3.3V when active and qualifies it for wearable proximity sensing applications.
+In the following a virtual capacitor denotes the capacitance between the sensing electrode and the environment.
+The capacitance of the sensing electrode to environmental objects increases with closer distances.
+% The capacitive proximity sensor performs measurements in loading mode.
 Two electrodes are integrated into the wristband, one sensing electrode and one shielding electrode that aims to eliminate the influence of the underlying skin.
@@
 The shield is driven with the same potential as the sensing electrode, such that the the capacitance between the two electrodes is negligible.
-Using this shielding method the measured capacitance will only be slightly affected by the grounded underlying arm.
+Using this shielding method the measured capacitance will only be slightly
+affected by the grounded underlying arm.
 
-\begin{figure}
+\begin{figure}[t]
 	\centering
-		\includegraphics[width=1.00\textwidth]{Images/capacitive_sensor_wristband_2.jpg}
+		\includegraphics[width=\textwidth]{Images/capacitive_sensor_wristband_2.jpg}
 	\caption{The hardware prototype at a glance: HedgeHog activity logger at the lower right, the capacitive sensor unit at the lower left, and the wristband with the sensing and the shield electrodes on-top each other covered with adhesive tape for isolation purposes.}
@@
 \end{figure}
-
-Figure \ref{fig:cap_sensor} shows the the wrist-worn prototype used in the evaluation experiments, with the HedgeHog as the main data logger, the capacitive sensor circuit and the wristband holding the sensing and shielding electrodes.
 
@@
 \end{figure}
+
+Figure \ref{fig:cap_sensor} shows the the wrist-worn prototype used in the evaluation experiments, with the HedgeHog as the main data logger, the capacitive sensor circuit and the wristband holding the sensing and shielding electrodes.
 
@@
 The opening door activity has very poor recognition rates without the data from the proximity sensor.
-The average f-measure could be increased from 35.5\% to 46.2\%.
+The average F-measure could be increased from 35.5\% to 46.2\%.
 A plot of the activity is given in Figure \ref{fig:opendoor}.
@@
 The acceleration sensor captures relevant data in the time in which the person moves into the room and the hand changes from the outer to the inner door knob (5 - 7s).
-The recorded data for this activity shows strong correlations between all test persons.
-The confusion matrices show that the ``open door'' activity was often confused with the ``sitting'' activity, because of the horizontal orientation of the hand.
-Due to the usage of the capacitive sensor, the recall for that class and confusion with the sitting activity improved.
-[wording, orientation vs little motion ?]
+The recorded data for this activity also shows strong correlations between all experiment participants.
+The confusion matrices show that the ``open door'' activity was often confused with the ``sitting'' activity, probably because of the amount of motion on the one hand and the proximity to nearby objects (door, couch or cushions) on the other hand.
+By using the capacitive sensor data, the recall for that class and confusion with the sitting activity could be improved.
 
@@
 However, the capacitive proximity sensor is able to indicate when a hand is placed on the surface of the couch.
-Especially for this particular participant, the f-measure increased from 50.3\% to 60.4\%, while the average f-measure improved from 68.0\% to 74.4\%.
+Especially for this particular participant, the F-measure increased from 50.3\% to 60.4\%, while the average F-measure improved from 68.0\% to 74.4\%.
 
@@
 Again, there were great variations in how this activity was performed by the participants.
-For example, some of them crossed their hands under their head, or placed them on their stomach.
-For this class, the average f-measure could only be increased by 2.4\%, from 81.2 to 83.6\%.
-Considering some test persons, this class was confused with the ``making bread'' activity and could be improved by the proximity sensor.
-[ wording!
-]
+For example, some of them crossed their hands under their head, or placed them on their body.
+For this class, the average F-measure could only be increased by 2.4\%, from 81.2 to 83.6\%.
+Considering some participants, the activity was often confused with the ``making bread'' class.
+By using the proximity modality, the confusion between the two classes could be reduced.
 
@@
 The capacitive proximity sensor was able to capture the proximity to the shelf and to the table.
-The average f-measure for this activity is rather low, but improved by 6.8\% from 53.8\% to 60.6\%.
-The worst performing participants for this activity reached an f-measure of 46.8\% without and 53.5\% with the capacitive sensor, while the best performing one reached 62.0\% and 64.5\% respectively.
-The quite low performance figures result from confusions with other activities, with a higher tendency to the ``prepare bread'' activity across all participants, most likely because there are also various objects involved.
-The capacitive sensor modality has a positive impact reducing the confusion with other activities.
+The average F-measure for this activity is rather low, but improved by 6.8\% from 53.8\% to 60.6\%.
+The worst performing participants for this activity reached an F-measure of 46.8\% without and 53.5\% with the capacitive sensor, while the best performing one reached 62.0\% and 64.5\% respectively.
+The quite low performance figures result from confusions with other activities, with a higher tendency to the ``prepare bread'' activity across all participants.
+This is most likely due to various objects involved in both activities.
+The capacitive sensor modality has a much more positive impact reducing the confusion with other activities.
 
@@
 		\includegraphics[width=\textwidth]{../Auswertung/images/eugen_5.pdf}
-	\caption{An example of the ``preparing bread'' activity, where the participants had to put marmalade on a toast.
-The proximity sensor indicates the closeness to the table, while the acceleration sensor shows reoccurring movements.}
+	\caption{An example of the ``preparing bread'' activity, where the participants had to put marmalade on a slice of bread.
+The proximity sensor indicates the closeness to the table, while the acceleration sensor shows recurring hand motions.}
 	\label{fig:prep_bread}
@@
 It is notable that the acceleration data does not seem to provide any characteristic patterns, while the proximity sensor indicates a table, plate, or other objects in immediate distance.
-This activity showed a high improvement in the average f-measure by 10\%, from 49.0\% to 59.8\%, where the data delivered by the capacitive proximity sensor is taken into account.
+This activity showed a high improvement in the average F-measure by 10\%, from 49.0\% to 59.8\%, where the data delivered by the capacitive proximity sensor is taken into account.
 The making bread class was often confused with the sitting class for some users, probably due to lots of motion during the sitting, as mentioned previously.
@@
 		\includegraphics[width=\textwidth]{../Auswertung/images/eugen_6.pdf}
-	\caption{An eating person who took 5 bites from the bread. After each bite, the hand is placed on the table.}
+	\caption{An example of a participant eating a slice of bread with marmalade, taking 5 bites from it.
+After each bite, the hand is placed on the table, which can be recognized both in the acceleration as well as the proximity data plots.}
 	\label{fig:eating}
@@
 A good example of an eating activity is shown in Figure \ref{fig:eating}, where the participant took a few bites from the bread while putting it down every time.
-The average f-measure for that activity increased slightly by 2.7\% (from 79.5\% to 82.2\%).
+The average F-measure for that activity increased slightly by 2.7\% (from 79.5\% to 82.2\%).
 
-\begin{figure}
+\begin{figure}[t]
 	\centering
@@
 There were problems in distinguishing this activity from get things that could be bettered by using the new input modality.
-The classification improvement for this activity accounts to 12.2\% boosting the average f-measure from 41.7\% without to 53.9\% with the new sensor.
+The classification improvement for this activity accounts to 12.2\% boosting the average F-measure from 41.7\% without to 53.9\% with the new sensor.
 Due to the low performance results and the characteristic periodic signal shape, it would help to consider frequency domain features, as it is often applied in related work.
 
-\begin{figure}
+\begin{figure}[t]
 	\centering
 		\includegraphics[width=1.00\textwidth]{../Auswertung/images/eugen_9.pdf}
-	\caption{During the sleeping activity the data from both sensors remains constant. One draw conclusions about the coverage of the arm with either couchons, blankets or the proximity to the mattress.}
+	\caption{During the sleeping activity the data from both sensors remains constant for large time spans.
+One can draw conclusions about the coverage of the arm with either cushions, blankets or the proximity to the mattress, the head or body of the participant.}
 	\label{fig:sleeping}
@@
 
-The sleeping activity (an example shown in Figure \ref{fig:sleeping}) was classified with an average f-measure of 83.2\%, which increased to 86.9\% when using the proximity data.
+The sleeping activity (an example shown in Figure \ref{fig:sleeping}) was classified with an average F-measure of 83.2\%, which increased to 86.9\% when using the proximity data.
 In this case, the capacitive proximity sensor is able to capture the surrounding cushions and blankets, as well as the body or head of the participant.
-The accelerometer data and the proximity data has larger periods with a constant signal.
-[further details?]
+The accelerometer data and the proximity data has larger periods with a constant signal, which is a cause for confusion with the ``lying on the couch'' activity.
 
 \begin{figure}
-	\setlength{\tabcolsep}{2.4pt}
+	\setlength{\tabcolsep}{2pt}
 	\centering
@@
 		\cline{1-9} \cline{11-19} \cline{21-21}
-% 		\textbf{9} & 5 & 0 & 0 & 1 & 0 & 0 & 0 & 0 		&& \textbf{10} & 3 & 0 & 2 & 0 & 0 & 0 & 0 & 0 		&& a = open door	\\ 
-% 		1 & \textbf{248} & 5 & 6 & 4 & 2 & 5 & 0 & 0 	&& 2 & \textbf{259} & 4 & 2 & 4 & 2 & 4 & 0 & 0 	&& b = sitting		\\ 
-% 		0 & 3 & \textbf{212} & 3 & 2 & 1 & 6 & 0 & 8 	&& 0 & 5 & \textbf{213} & 2 & 2 & 1 & 8 & 0 & 4 	&& c = lying		\\ 
-% 		0 & 8 & 2 & \textbf{44} & 7 & 8 & 1 & 1 & 3 	&& 3 & 5 & 2 & \textbf{52} & 9 & 3 & 1 & 1 & 1 		&& d = get things	\\ 
-% 		1 & 7 & 1 & 11 & \textbf{37} & 9 & 2 & 0 & 10 	&& 1 & 9 & 3 & 13 & \textbf{46} & 6 & 3 & 0 & 0 	&& e = make bread	\\ 
-% 		0 & 3 & 2 & 3 & 8 & \textbf{58} & 5 & 0 & 1 	&& 0 & 5 & 1 & 1 & 5 & \textbf{62} & 5 & 0 & 1 		&& f = eating		\\ 
-% 		0 & 8 & 3 & 0 & 6 & 4 & \textbf{67} & 0 & 1 	&& 0 & 8 & 4 & 2 & 1 & 5 & \textbf{68} & 0 & 0 		&& g = drinking		\\
-% 		1 & 0 & 0 & 1 & 0 & 0 & 0 & \textbf{12} & 0 	&& 1 & 0 & 0 & 1 & 0 & 0 & 0 & \textbf{13} & 0 		&& h = walking		\\
-% 		0 & 1 & 2 & 10 & 6 & 1 & 1 & 0 & \textbf{312} 	&& 1 & 0 & 0 & 7 & 0 & 0 & 0 & 0 & \textbf{325} 	&& i = sleeping		\\ 
-% 	
-%		\textbf{11} & 0 & 5 & 4 & 1 & 0 & 0 & 0 & 0 	&& \textbf{17} & 0 & 0 & 2 & 0 & 0 & 0 & 0 & 0		&& a = open door    \\
-%		0 & \textbf{136} & 8 & 1 & 11 & 0 & 0 & 0 & 0	&& 0 & \textbf{140} & 7 & 2 & 3 & 0 & 0 & 0 & 0		&& b = sitting      \\
-%		4 & 20 & \textbf{151} & 1 & 7 & 2 & 1 & 1 & 0	&& 0 & 23 & \textbf{149} & 5 & 7 & 3 & 2 & 1 & 0	&& c = lying        \\
-%		3 & 6 & 5 & \textbf{116} & 21 & 9 & 4 & 4 & 2	&& 1 & 1 & 9 & \textbf{129} & 25 & 4 & 2 & 4 & 1	&& d = get things   \\
-%		0 & 5 & 2 & 5 & \textbf{81} & 11 & 2 & 0 & 0	&& 0 & 3 & 1 & 15 & \textbf{83} & 6 & 1 & 0 & 1		&& e = make bread   \\
-%		0 & 0 & 0 & 0 & 9 & \textbf{167} & 19 & 0 & 3	&& 0 & 0 & 0 & 4 & 8 & \textbf{158} & 26 & 0 & 3	&& f = eating       \\
-%		1 & 2 & 3 & 4 & 8 & 26 & \textbf{89} & 0 & 5	&& 0 & 0 & 8 & 5 & 5 & 29 & \textbf{92} & 1 & 4		&& g = drinking     \\
-%		0 & 0 & 1 & 17 & 3 & 0 & 0 & \textbf{2} & 0		&& 0 & 1 & 2 & 10 & 3 & 0 & 0 & \textbf{9} & 0		&& h = walking      \\
-%		0 & 0 & 0 & 2 & 0 & 8 & 2 & 0 & \textbf{78}		&& 0 & 0 & 0 & 2 & 0 & 6 & 1 & 0 & \textbf{81}		&& i = sleeping     \\
-
 \textbf{8} & 0 & 0 & 0 & 0 & 0 & 0 & 2 & 1		&& \textbf{11} & 0 & 0 & 2 & 1 & 0 & 0 & 0 & 1 	&& a = open door	\\ 
@@
 1 & 7 & 1 & \textbf{50} & 17 & 9 & 1 & 4 & 1	&& 2 & 6 & 0 & \textbf{67} & 14 & 9 & 1 & 2 & 1 && d = get things	\\ 
-		0 & 2 & 7 & 7 & \textbf{102} & 42 & 5 & 0 & 2			&& 1 & 1 & 0 & 13 & \textbf{120} & 24 & 7 & 1 & 1 && e = make bread	\\ 
+0 & 2 & 7 & 7 & \textbf{102} & 42 & 5 & 0 & 2	&& 1 & 1 & 0 & 13 & \textbf{120} & 24 & 7 & 1 & 1&& e = make bread	\\ 
 0 & 0 & 3 & 5 & 21 & \textbf{288} & 4 & 0 & 0 	&& 0 & 0 & 2 & 1 & 19 & \textbf{300} & 3 & 0 & 0 && f = eating		\\ 
@@
 0 & 2 & 2 & 0 & 2 & 0 & 0 & 0 & \textbf{271}	&& 0 & 0 & 0 & 0 & 1 & 0 & 0 & 0 & \textbf{276}  && i = sleeping	\\ 
-
 		\cline{1-9} \cline{11-19} 
@@
 Here, we are comparing SVM classification presented as confusion matrices for an exemplary user, without the proximity data on the left, and with the proximity data on the right.
-Note: reject class is not included in the confusion matrix.}
-
+Note that the reject class (background data not annotated as an activity) is not included in the confusion matrix.}
 	\label{fig:ConfusionMatrixForAnExemplaryUser}
@@
 
-Figure \ref{fig:ConfusionMatrixForAnExemplaryUser} depicts an example for the SVM classification results for one of the participants.
+Figure \ref{fig:ConfusionMatrixForAnExemplaryUser}  depicts two confusion matrices for an exemplary participant from our evaluation, once without and once with including the proximity data.
+In most activities, an enhancement in the number of correctly classified instances is observable.
+The ``lying'' activity's recognition performance could benefit a lot from the proximity data, improving both precision and recall.
+
+A better classification performance can also be observed for the ``get things'' class that includes interactions with a multitude of objects in the environment.
+Regarding the activity ``make bread'', the capacitive proximity could reduce the number of confusions with the ``eating'' class significantly.
+
+Due to the high similarity of eating and drinking (cf.
+Figure \ref{fig:eating} and \ref{fig:drinking}), the number of confusions between those two classes increases when considering the proximity data.
+However, for the ``drinking'' class the new modality limits the confusions to related activities such as ``eating'' and ``make bread'' only, while lowering the number of false recognitions for the other activities.
 
@@
 
-We have shown that current wearable inertial sensing devices for activity recognition can be significantly enhanced by integrating a capacitive proximity sensor into the wristband.
-Regarding the classification accuracy, we obtained an overall increase in correctly classified instances of TODO percent.
-Some classes, for instance ``preparing bread'' and ``drinking'', could benefit very much from proximity-related sensor data.
-For such classes the classification performance could be boosted by up to TODO percent.
+Our paper presented a wrist-worn activity data logger prototype, which consists of an accelerometer in combination with a capacitive proximity sensor integrated into the wristband.
+Our experiments with seven participants and nine basic daily activities show that this additional input modality can significantly boost the activity recognition performance.
+Regarding the classification performance, we obtained an improvement in the average F-measure of 6.3\%, from 67.2 to 73.5\%.
+Specifically, the activity classes ``walking'', ``preparing bread'' and ``open door'' could benefit a lot from proximity-related sensor data.
+For such classes, the classification performance could be boosted by 12.2, 9.0 and 10.8\% respectively.
 
-The classification results could be improved by the usage of more than one sensing electrode in the wristband.
-For example the wristband could integrate four electrodes that are placed on each side of the arm.
-However, this will lead to smaller electrode surfaces that lead to a decrease of sensing distance.
-We think that these considerations are an important subject of future work.
-The sensor's power consumption can be decreased by shorter measurement windows and the choice of more energy efficient hardware components.
-Measuring pulse width lengths instead of counting the rising edges of the sensor's signal may reduce the required time needed for a measurement.
+With this proof of concept we show that the proximity information can provide an information gain regarding the evaluated activities.
+In future work we aim at evaluating other relevant feature types, such as frequency domain features, as well as feature sets to extract the most discriminative ones.
+Additionally, using other classifiers (such as HMMs) or more fitting classifier configurations might also improve activity recognition performance.
+
+The classification results could also be improved by using more than one sensing electrode in the wristband.
+For example, the wristband could integrate up to four electrodes that are placed on each side of the arm.
+However, this will lead to smaller electrode surfaces thus resulting in a decreased sensing distance.
+The sensor's power consumption can be decreased by shorter measurement windows and the choice of more energy efficient hardware components as well as software implementation.
+Measuring pulse width lengths instead of counting the number of pulses of the sensor's signal may reduce the required time needed for a measurement, thus increasing the time the microcontroller is able to sleep.
 
@@
 
-% \newpage
-
 \bibliographystyle{splncs03}

### 2012-06-22 Paper/Paper.tex
final paper version for the submission

@@
 % a short form should be given in case it is too long for the running head
-\titlerunning{Enhancing Accelerometer-based Activity Recognition}
+\titlerunning{Enhancing Accelerometer-based Activity Recognition with Proximity Sensing}
 % (feature abused for this document to repeat the title also on left hand pages)
-\authorrunning{Enhancing Accelerometer-based Activity Recognition}
+\authorrunning{Enhancing Accelerometer-based Activity Recognition with Proximity Sensing}
 
@@
 Activity recognition with an accelerometer is a common investigated research topic and enables the detection of basic activities like sitting, walking or standing.
-Recent work in this area adds different sensing modalities to the inertial data to collect more information of the users environment and to boost activity recognition.
-This work presents a sensor prototype which consists of an accelerometer and a capacitive proximity sensor that senses the users activity based on the combined sensor values.
+Recent work in this area adds different sensing modalities to the inertial data to collect more information of the user's environment to boost activity recognition.
+This work presents a sensor prototype consisting of an accelerometer and a capacitive proximity sensor that senses the user's activities based on the combined sensor values.
 We show that our proposed approach significantly improves the recognition rate for detecting activities of daily living (e.g.
@@
 Our approach to enhance the inertial data from a wearable sensor is comparable to the RFID scenarios just mentioned, as it also relies on a single wearable sensor and an unobtrusive deployment.
-The main difference lies in the fact that we do not consider an accurate detection of tagged objects, but the proximity to various objects and the environment.
+The main difference lies in the fact that we do not consider an accurate detection of tagged objects, but the proximity to various unknown objects and the environment.
 
@@
 
-The sensor node itself is built around the low-power Microchip microcontroller featuring an accelerometer sensor to capture human motion, light and ambient temperature sensors and a microSD flash card for locally storing the sensor data.
+The sensor node itself is built around the low-power Microchip microcontroller (PIC18F46J50) featuring an accelerometer sensor (ADXL345) to capture human motion, light and ambient temperature sensors and a microSD flash card for locally storing the sensor data.
 The sensor is powered by a $200mAh$ lithium polymer battery, which allows for two weeks of continuous recording on a single battery charge.
@@
 	\includegraphics[width=\textwidth]{Images/hardware_sensor_2.jpg}
-	\caption{The inertial data logger featuring a low-power microcontroller, a 3 axis accelerometer, a microSD flash card for storing the sensor data and a USB connector for accessing the data (on the right) is powered by a small lithium polymer battery and is packaged into a plastic case to be worn at the wrist (a version with a OLED display).}
+	\caption{The inertial data logger featuring a low-power microcontroller, a 3 axis accelerometer, a microSD flash card for storing the sensor data and a USB connector for accessing the data (on the right) is powered by a small lithium polymer battery and is packaged into a plastic case to be worn at the wrist (a version with an OLED display).}
 	\label{fig:sensornode}
@@
 A wrist-worn capacitive proximity sensor requires a shield that eliminates the influence of the grounded arm directly underneath the sensor.
-Using this setup, we can detect the proximity to a grounded object in the environments for distances up to 20 cm.
-Especially for mobile devices, it is required that the sensor only draws a very small amount of power.
+Using this setup, we can detect the proximity to a grounded object in the environments for distances up to 20cm.
+Especially for mobile devices, it is required that the sensor draws a very small amount of power.
 Thus, other proximity sensing input modalities like ultrasound or optical measurements are not applicable for this type of mobile application.
@@
 In order to guard the sensor from measuring the capacitance to underlying objects, a shield electrode is placed directly underneath the measuring electrode.
-The shield is driven with the same potential as the sensing electrode, such that the the capacitance between the two electrodes is negligible.
-Using this shielding method the measured capacitance will only be slightly
-affected by the grounded underlying arm.
+The shield is driven with the same potential as the sensing electrode, such that the capacitance between the two electrodes is negligible.
+Using this shielding method, the measured capacitance will only be slightly affected by the grounded underlying arm.
 
@@
 		\includegraphics[width=\textwidth]{Images/capacitive_sensor_wristband_2.jpg}
-	\caption{The hardware prototype at a glance: HedgeHog activity logger at the lower right, the capacitive sensor unit at the lower left, and the wristband with the sensing and the shield electrodes on-top each other covered with adhesive tape for isolation purposes.}
+	\caption{The hardware prototype at a glance: HedgeHog activity logger at the lower right, the capacitive sensor unit at the lower left, and the wristband with the sensing and the shield electrodes on-top each other.
+The electrodes are covered with adhesive tape for isolation purposes.}
 	\label{fig:cap_sensor}
@@
  		\includegraphics[trim=1cm 1cm 1cm 1cm,clip,width=\textwidth]{Images/pseudocode.pdf}
-	\caption{Measurement procedure carried out by the HedgeHog sensor: using the microcontroller's Timer0 module in counting mode, the oscillating signal generated by the capacitive sensor circuit can be measured by counting the frequency pulses over a predefined gate time of approximately 9.5ms.}
+	\caption{Overview of the measurement procedure carried out by the HedgeHog sensor: using the microcontroller's Timer0 module in counting mode, the oscillating signal generated by the capacitive sensor circuit can be measured by counting the frequency pulses over a predefined gate time of approximately 9.5ms.}
 	\label{fig:pseudocode}
@@
 
-The experiment setup aims to depict a typical scenario of person in daily life.
+The experiment setup aims to depict a typical scenario of a person in daily life.
 Especially in the field of Ambient Assisted Living (AAL), it is desired to monitor activities like drinking, preparing lunch and sleeping.
@@
 		\includegraphics[width=1.00\textwidth]{../Auswertung/images/eugen_2.pdf}
-	\caption{Example of a ``sitting'' activity in which the user moved his hands quite frequently (bottom plot).
+	\caption{Example of the ``sitting'' activity in which the user moved his hands quite frequently (bottom plot).
 Most of the time the values of the proximity sensor stay more or less constant, probably due to the hands position on the couch's fabric.
@@
 The worst performing participants for this activity reached an F-measure of 46.8\% without and 53.5\% with the capacitive sensor, while the best performing one reached 62.0\% and 64.5\% respectively.
-The quite low performance figures result from confusions with other activities, with a higher tendency to the ``prepare bread'' activity across all participants.
+The low performance results from confusions with other activities, with a higher tendency to the ``prepare bread'' activity across all participants.
 This is most likely due to various objects involved in both activities.
@@
 Some of the participants ate their bread leaving their hand close to the mouth, while others moved their hand up and down putting their bread aside on the plate, which also results in the performance range from 71.2 to 88.1\% without and 77.5 to 90.6\% with the proximity data.
-A good example of an eating activity is shown in Figure \ref{fig:eating}, where the participant took a few bites from the bread while putting it down every time.
+An example of an eating activity is shown in Figure \ref{fig:eating}, where the participant took a few bites from the bread while putting it down every time.
 The average F-measure for that activity increased slightly by 2.7\% (from 79.5\% to 82.2\%).
@@
 While walking, the capacitance between the wristband and the leg increases when the wristband is located close to the body and decrease when the wristband moves away.
-There were problems in distinguishing this activity from get things that could be bettered by using the new input modality.
+There were problems distinguishing this activity from ``get things'' that could be improved by using the new input modality.
 The classification improvement for this activity accounts to 12.2\% boosting the average F-measure from 41.7\% without to 53.9\% with the new sensor.
@@
 Capacitive proximity sensors represent a suitable new input modality for future activity recognition systems.
-The low power consumption as well as the invisible integration of a sensor into the wristband meets an essential requirement of wearable applications.
+The low power consumption as well as the invisible integration of a sensor into the wristband meets an essential requirement of wearable and unobtrusive applications.
 Especially in AAL environments, these systems can help monitoring the course of chronic diseases by recognizing activities of daily life.

### 2012-08-10 Paper/Paper.tex
changed abstract and added limitations of capacitive sensors, deanonymized it

@@
 % the name(s) of the author(s) follow(s) next
-% \author{Tobias Grosse-Puppendahl\inst{1} \and Eugen Berlin\inst{2} \and Marko Borazio\inst{2} }
-\author{Author 1 \and Author 2 \and Author 3 }
+\author{Tobias Grosse-Puppendahl\inst{1} \and Eugen Berlin\inst{2} \and Marko Borazio\inst{2} }
+%\author{Author 1 \and Author 2 \and Author 3 }
 
@@
 % unless you accept that it will be published
-% \institute{
-% 	Fraunhofer IGD, Fraunhoferstr. 5, 64283 Darmstadt, Germany\\
-% 	\email{tobias.grosse-puppendahl@igd.fraunhofer.de}
-% 	\and
-% 	Technische Universit\"at Darmstadt, Hochschulstr. 10, 64289 Darmstadt, Germany\\
-% 	\email{\{berlin,borazio\}@ess.tu-darmstadt.de}
-% }
-
 \institute{
-	Anonymized for review \\
-	\email{anonymous@insitution.edu}
+ 	Fraunhofer IGD, Fraunhoferstr. 5, 64283 Darmstadt, Germany\\
+ 	\email{tobias.grosse-puppendahl@igd.fraunhofer.de}
  	\and
-	Anonymized for review \\
-	\email{anonymous@insitution.edu}
+ 	Technische Universit\"at Darmstadt, Hochschulstr. 10, 64289 Darmstadt, Germany\\
+ 	\email{\{berlin,borazio\}@ess.tu-darmstadt.de}
 }
 
-% \toctitle{Enhancing Accelerometer-based Activity Recognition with Capacitive Proximity Sensing}
-% \tocauthor{Eugen Berlin, Marko Borazio, Tobias Grosse-Puppendahl}
+%\institute{
+	%Anonymized for review \\
+	%\email{anonymous@insitution.edu}
+	%\and
+	%Anonymized for review \\
+	%\email{anonymous@insitution.edu}
+%}
+
+\toctitle{Enhancing Accelerometer-based Activity Recognition with Capacitive Proximity Sensing}
+\tocauthor{Eugen Berlin, Marko Borazio, Tobias Grosse-Puppendahl}
 
@@
 This work presents a sensor prototype consisting of an accelerometer and a capacitive proximity sensor that senses the user's activities based on the combined sensor values.
-We show that our proposed approach significantly improves the recognition rate for detecting activities of daily living (e.g.
-sitting, preparing dinner, eating, drinking, etc.) by adding capacitive proximity sensor data to the inertial data.
+We show that our proposed approach of combining both modalities significantly improves the recognition rate for detecting activities of daily living.
 \keywords{activity recognition, capacitive proximity sensors, ambient assisted living, user context}
@@
 In this work, we show that an accelerometer and a capacitive proximity sensor can be used to improve activity recognition in activities of daily living.
-We obtained an open-hardware and open-source wrist-worn activity data logger \cite{hedgehog} and integrated a capacitive proximity sensor invisibly into the wristband.
+Therefore, we obtained an open-hardware and open-source wrist-worn activity data logger \cite{hedgehog} and integrated a capacitive proximity sensor invisibly into the wristband.
 
@@
 
+Capacitive proximity sensing faces the great advantage of being robust against changing lighting conditions and occlusion.
+Moreover, sensing electrodes can be integrated invisibly into the environment.
+On the other hand, the exact distance to objects can only be approximated since the object's surface, its conductivity and grounding has influence on the measurement result.
+A single sensor will thus deliver data that has a certain degree of ambiguity.
+Due to the nature of capacitive proximity sensors, they can be prone to errors in environments with strong and rapidly changing electric fields.
+This, however, is usually not an issue when considering activities of daily living.
+
 A great variety of capacitive sensors and measurement techniques exists \cite{Smith1999}.
@@
 in \cite{Wimmer2007} who presented a toolkit for capacitive proximity sensing.
-
 In previous works, capacitive proximity sensors were applied in various fields of human-computer-interaction.
@@
 
-\begin{figure}
+\begin{figure}[h]
 	\centering
@@
 
-Anonymized for review.
-
-% We would like to thank our evaluation participants from TU Darmstadt and Fraunhofer IGD.
+We would like to thank our evaluation participants from TU Darmstadt and Fraunhofer IGD.
 

### 2012-08-23 Paper/Paper.tex
images and tables placed at the top of the pages to improve readability of the text
final version, changed *_5.pdf's and some text

@@
 \institute{
- 	Fraunhofer IGD, Fraunhoferstr. 5, 64283 Darmstadt, Germany\\
+ 	Fraunhofer IGD, Darmstadt, Germany\\
  	\email{tobias.grosse-puppendahl@igd.fraunhofer.de}
  	\and
- 	Technische Universit\"at Darmstadt, Hochschulstr. 10, 64289 Darmstadt, Germany\\
+ 	Embedded Sensing Systems, Technische Universit\"at Darmstadt, Germany\\
  	\email{\{berlin,borazio\}@ess.tu-darmstadt.de}
@@
 \begin{abstract}
-Activity recognition with an accelerometer is a common investigated research topic and enables the detection of basic activities like sitting, walking or standing.
-Recent work in this area adds different sensing modalities to the inertial data to collect more information of the user's environment to boost activity recognition.
+Activity recognition with a wearable accelerometer is a common investigated research topic and enables the detection of basic activities like sitting, walking or standing.
+Recent work in this area adds different sensing modalities to the inertial data to collect more information of the user's environment to boost activity recognition for more challenging activities.
 This work presents a sensor prototype consisting of an accelerometer and a capacitive proximity sensor that senses the user's activities based on the combined sensor values.
@@
 Current wearable activity recognition systems are able to unobtrusively capture and recognize a person's activities throughout the whole day.
-These systems often rely on inertial sensor data that is captured by wearable sensors embedded in a mobile device \cite{brezmes2009activity} or attached to the body \cite{Ravi2005}.
+These systems often rely on inertial sensor data that are captured by wearable sensors embedded in a mobile device \cite{brezmes2009activity} or attached to the body \cite{Ravi2005}.
 Usually, single sensor modalities are used or duplicated to detect the activities.
-However, it is a great challenge to identify fine-grained activities just by using a single modality like the accelerometer.
+However, it is a great challenge to identify many activities just by using a single modality like the accelerometer.
 
@@
 This means that the measurement result depends on the object's distance, its size and the material it is made of.
-In this work, we show that an accelerometer and a capacitive proximity sensor can be used to improve activity recognition in activities of daily living.
-Therefore, we obtained an open-hardware and open-source wrist-worn activity data logger \cite{hedgehog} and integrated a capacitive proximity sensor invisibly into the wristband.
+In this work, we show that an accelerometer and a capacitive proximity sensor can be used to improve activity recognition in activities of daily living that rely heavily upon object usage.
+Therefore, we obtained an open-hardware and open-source wrist-worn activity data logger \cite{hedgehog} and integrated a capacitive proximity sensor into its wristband.
 
@@
 For example, it may be conducted which material is placed underneath the wristband.
-The capacitive proximity sensor would return a different measurement result for a hand placed on a couch covered with fabric than a hand placed a wooden table.
+The capacitive proximity sensor would return a different measurement result for a hand placed on a couch covered with fabric than a hand placed on a wooden table.
 Moreover, the approximate distance from the wristband to objects can be exploited to identify activities like grasping into a locker or a refrigerator to prepare food.
@@
 have investigated the possibility of using capacitive sensors for activity recognition by measuring shape changes of muscles and skin \cite{Cheng2010}.
-To our knowledge, capacitive proximity sensors have not been embedded into a wearable device to enhance the performance of activity recognition.
+%To our knowledge, capacitive proximity sensors have not been embedded into a wearable device to enhance the performance of activity recognition.
 
@@
 setting the sensitivity of the accelerometer), to access the stored sensor data, and to recharge the battery.
-A plastic case nicely packages and protects the sensor to be worn at the wrist (Figure \ref{fig:sensornode}).
+A plastic case packages and protects the sensor to be worn at the wrist (Figure \ref{fig:sensornode}).
 
@@
 Two electrodes are integrated into the wristband, one sensing electrode and one shielding electrode.
-The sensor draws a supply current of 1mA at 3.3V when active and qualifies it for wearable proximity sensing applications.
+The sensor draws a supply current of 1mA at 3.3V when active which qualifies it for wearable proximity sensing applications.
 In the following a virtual capacitor denotes the capacitance between the sensing electrode and the environment.
@@
 During that counting phase, the microcontroller is sent to sleep in order to reduce power consumption.
-In the following, the HedgeHog applies run-length encoding on the measured data to reduce overhead and periodically logs the data to the integrated microSD card.
+%In the following, the HedgeHog applies run-length encoding on the measured data to reduce overhead and periodically logs the data to the integrated microSD card.
 
@@
 
-The recorded test set contains the following activities: opening door, sitting on a couch, lying on a couch, putting kitchen equipments from a shelf and out of a locker, preparing a bread with marmalade, eating the bread, pouring and drinking water, walking and sleeping.
+The recorded test set contains the following activities: opening door, sitting on a couch, lying on a couch, putting kitchen equipments from a shelf and out of a locker, making a marmalade sandwich, eating the sandwich, pouring and drinking water, walking and sleeping.
 The relations of these activities to environmental objects are given in Table \ref{tab:activities}.
@@
 						& bread, marmalade, bottle  & 							\\
-		making bread	& bread, knife, marmalade	& table, plate 				\\
-		eating			& bread						& table, plate, body		\\
+		make sandwich	& bread, knife, marmalade	& table, plate 				\\
+		eating			& marmalade sandwich		& table, plate, body		\\
 		drinking 		& bottle, glass 			& table, body				\\
@@
 In order to evaluate if capacitive proximity sensors in wrist-bands can enhance the performance of activity recognition, we have conducted an evaluation with 7 test persons.
-All test persons received a basic plot with the activities they were supposed to perform.
+All test persons received a basic script with the activities they were supposed to perform.
 They were not given any instructions about the way they are supposed to perform the activities.
-After manual labeling, we used this test-set as ground truth and performed a 4-fold cross-validation on an support-vector machine (SVM) classifier on each user.
+After manual labelling, we used this test-set as ground truth and performed a 4-fold cross-validation on an support-vector machine (SVM) classifier on each user.
 The cross-validation was performed once with and once without including the data of the capacitive proximity sensor into the feature set.
@@
 
-In the following, the performed activities will be analyzed in detail, stating out the influence of the capacitive proximity sensor on the classification result.
-In general, the usage of data provided by the new modality showed improvements in recognition rates reaching from 2.4 up to 10.7\% for a single activity.
+In the following, the performed activities will be analyzed in detail, stating the influence of the capacitive proximity sensor on the classification result.
+In general, the usage of data provided by the new modality showed improvements in recognition rates reaching from 2.4\% up to 10.7\% for single activities.
 
@@
 		\includegraphics[width=1.00\textwidth]{../Auswertung/images/tobias_1.pdf}
-	\caption{When the participants entered the apartment, the wristband approached the door knob twice, at the time of opening and closing the door.
-This fact can be observed in the capacitive proximity data (upper plot) at the beginning and at the end of the activity, whereas the acceleration has no characteristic information (bottom plot).}
+	\caption{When the participants entered the apartment, the wrist approached the door knob twice, at the time of opening and closing the door.
+This fact can be observed in the capacitive proximity data (upper plot) at the beginning and at the end of the activity, whereas the acceleration has little characteristic information (bottom plot).}
 	\label{fig:opendoor}
@@
 
-The opening door activity has very poor recognition rates without the data from the proximity sensor.
+The ''opening door'' activity has very poor recognition rates without the data from the proximity sensor.
 The average F-measure could be increased from 35.5\% to 46.2\%.
@@
 For this class, the average F-measure could only be increased by 2.4\%, from 81.2 to 83.6\%.
-Considering some participants, the activity was often confused with the ``making bread'' class.
+Considering some participants, the activity was often confused with the ``make sandwich'' class.
 By using the proximity modality, the confusion between the two classes could be reduced.
 
-\begin{figure}
+\begin{figure}[t]
 	\centering
@@
 The worst performing participants for this activity reached an F-measure of 46.8\% without and 53.5\% with the capacitive sensor, while the best performing one reached 62.0\% and 64.5\% respectively.
-The low performance results from confusions with other activities, with a higher tendency to the ``prepare bread'' activity across all participants.
-This is most likely due to various objects involved in both activities.
+The low performance results from confusions with other activities, with a higher tendency to the ``make sandwich'' activity across all participants.
+This is most likely due to various objects involved in both activities, and the fact that 1-second features are obtained.
 The capacitive sensor modality has a much more positive impact reducing the confusion with other activities.
@@
 		\includegraphics[width=\textwidth]{../Auswertung/images/eugen_5.pdf}
-	\caption{An example of the ``preparing bread'' activity, where the participants had to put marmalade on a slice of bread.
+	\caption{An example of the ``make sandwich'' activity, where the participants had to put marmalade on a slice of bread.
 The proximity sensor indicates the closeness to the table, while the acceleration sensor shows recurring hand motions.}
@@
 This activity showed a high improvement in the average F-measure by 10\%, from 49.0\% to 59.8\%, where the data delivered by the capacitive proximity sensor is taken into account.
-The making bread class was often confused with the sitting class for some users, probably due to lots of motion during the sitting, as mentioned previously.
-For other users, making bread was confused with eating or drinking.
+The ''make sandwich'' class was often confused with the ''sitting'' class for some users, probably due to lots of motion during the sitting, as mentioned previously.
+For other users, ''make sandwich'' was confused with ''eating'' or ''drinking''.
 Using the new input modality, confusion across users could be reduced.
@@
 		\includegraphics[width=\textwidth]{../Auswertung/images/eugen_6.pdf}
-	\caption{An example of a participant eating a slice of bread with marmalade, taking 5 bites from it.
+	\caption{An example of a participant eating a marmalade sandwich, taking 5 bites from it.
 After each bite, the hand is placed on the table, which can be recognized both in the acceleration as well as the proximity data plots.}
@@
 When considering the eating activity, the impact of the new capacitive proximity sensor on the classification performance is quite low, as the chosen features are able to distinguish it from other activities.
-Some of the participants ate their bread leaving their hand close to the mouth, while others moved their hand up and down putting their bread aside on the plate, which also results in the performance range from 71.2 to 88.1\% without and 77.5 to 90.6\% with the proximity data.
-An example of an eating activity is shown in Figure \ref{fig:eating}, where the participant took a few bites from the bread while putting it down every time.
+Some of the participants ate their sandwich leaving their hand close to the mouth, while others moved their hand up and down putting their sandwich aside on the plate, which also results in the performance range from 71.2 to 88.1\% without and 77.5 to 90.6\% with the proximity data.
+An example of an eating activity is shown in Figure \ref{fig:eating}, where the participant took a few bites from the sandwich while putting it down every time.
 The average F-measure for that activity increased slightly by 2.7\% (from 79.5\% to 82.2\%).
@@
 The activity ``drinking water'' is depicted in Figure \ref{fig:drinking}.
-The participant took a few drinks from the glass, while leaving the hands lying on the table in between.
+The participant took a few sips from the glass, with leaving the hands positioned.
 These motions can be easily detected in the acceleration as well as proximity data.
 The accelerometer data shows that there periodic up- and down-movements while the capacitive proximity sensor delivers data that is associated to the proximity of the table.
-The average classification performance lies at 48.4\% without and 54.4\% with the proximity data taken into account, resulting in a gain of 6\%.
+The F-measure lies at 48.4\% without and 54.4\% with the proximity data taken into account, resulting in a gain of 6\%.
 
@@
 Regarding the walking activity one can identify periodic changes in the acceleration as well as in the measured capacitance, illustrated in Figure \ref{fig:walking}.
-While walking, the capacitance between the wristband and the leg increases when the wristband is located close to the body and decrease when the wristband moves away.
+While walking, the capacitance between the wristband and the leg increases when the wristband is located close to the body and decreases when the wristband moves away.
 There were problems distinguishing this activity from ``get things'' that could be improved by using the new input modality.
@@
 In this case, the capacitive proximity sensor is able to capture the surrounding cushions and blankets, as well as the body or head of the participant.
-The accelerometer data and the proximity data has larger periods with a constant signal, which is a cause for confusion with the ``lying on the couch'' activity.
+The accelerometer data and the proximity data have larger periods of low variance, which is a cause for confusion with the ``lying on the couch'' activity.
 
-\begin{figure}
-	\setlength{\tabcolsep}{2pt}
+\begin{figure}[t]
+	\setlength{\tabcolsep}{1.7pt}
 	\centering
@@
 1 & 7 & 1 & \textbf{50} & 17 & 9 & 1 & 4 & 1	&& 2 & 6 & 0 & \textbf{67} & 14 & 9 & 1 & 2 & 1 && d = get things	\\ 
-0 & 2 & 7 & 7 & \textbf{102} & 42 & 5 & 0 & 2	&& 1 & 1 & 0 & 13 & \textbf{120} & 24 & 7 & 1 & 1&& e = make bread	\\ 
+0 & 2 & 7 & 7 & \textbf{102} & 42 & 5 & 0 & 2	&& 1 & 1 & 0 & 13 & \textbf{120} & 24 & 7 & 1 & 1&& e = make sandwich	\\ 
 0 & 0 & 3 & 5 & 21 & \textbf{288} & 4 & 0 & 0 	&& 0 & 0 & 2 & 1 & 19 & \textbf{300} & 3 & 0 & 0 && f = eating		\\ 
@@
 A better classification performance can also be observed for the ``get things'' class that includes interactions with a multitude of objects in the environment.
-Regarding the activity ``make bread'', the capacitive proximity could reduce the number of confusions with the ``eating'' class significantly.
+Regarding the activity ``make sandwich'', the capacitive proximity could reduce the number of confusions with the ``eating'' class significantly.
 
@@
 Figure \ref{fig:eating} and \ref{fig:drinking}), the number of confusions between those two classes increases when considering the proximity data.
-However, for the ``drinking'' class the new modality limits the confusions to related activities such as ``eating'' and ``make bread'' only, while lowering the number of false recognitions for the other activities.
+However, for the ``drinking'' class the new modality limits the confusions to related activities such as ``eating'' and ``make sandwich'' only, while lowering the number of false recognitions for the other activities.
 
@@
 Our paper presented a wrist-worn activity data logger prototype, which consists of an accelerometer in combination with a capacitive proximity sensor integrated into the wristband.
-Our experiments with seven participants and nine basic daily activities show that this additional input modality can significantly boost the activity recognition performance.
+Our experiments with seven participants and nine daily activities show that this additional input modality can significantly boost the activity recognition performance.
 Regarding the classification performance, we obtained an improvement in the average F-measure of 6.3\%, from 67.2 to 73.5\%.
-Specifically, the activity classes ``walking'', ``preparing bread'' and ``open door'' could benefit a lot from proximity-related sensor data.
-For such classes, the classification performance could be boosted by 12.2, 9.0 and 10.8\% respectively.
+Specifically, the activity classes ``walking'', ``make sandwich'' and ``open door'' could benefit a lot from proximity-related sensor data.
+For such classes, the classification performance could be boosted by 12.2\%, 9.0\% and 10.8\% respectively.
 
@@
 In future work we aim at evaluating other relevant feature types, such as frequency domain features, as well as feature sets to extract the most discriminative ones.
-Additionally, using other classifiers (such as HMMs) or more fitting classifier configurations might also improve activity recognition performance.
+Additionally, using other classifiers (such as HMMs) might also improve activity recognition performance.
 
@@
 Capacitive proximity sensors represent a suitable new input modality for future activity recognition systems.
-The low power consumption as well as the invisible integration of a sensor into the wristband meets an essential requirement of wearable and unobtrusive applications.
+The low power consumption as well as the unobtrusive integration of a sensor into the wristband meets an essential requirement of wearable applications.
 Especially in AAL environments, these systems can help monitoring the course of chronic diseases by recognizing activities of daily life.
@@
 \bibliographystyle{splncs03}
-\bibliography{references}
+%\bibliography{references}
 

### 2012-08-24 Paper/Paper.tex
last changes in the text, added a reference

@@
 Persons affected from physical and mental restrictions and their care givers can profit significantly from unobtrusive activity monitoring solutions.
-For example, formal care givers could equip a person with a wearable activity monitoring solution that evaluates the course of a disease or the influence of an adapted medication.
+For example, formal care givers could equip a person with a wearable activity monitoring system that evaluates the course of a disease or the influence of an adapted medication \cite{Muhlsteff2004a}.
 This scenario is especially relevant for people suffering from dementia who have limitations in organizing their daily activities.
@@
 
+\begin{figure}[h]
+	\centering
+	\includegraphics[width=\textwidth]{Images/hardware_sensor_2.jpg}
+	\caption{The inertial data logger featuring a low-power microcontroller, a 3 axis accelerometer, a microSD flash card for storing the sensor data and a USB connector for accessing the data (on the right) is powered by a small lithium polymer battery and is packaged into a plastic case to be worn at the wrist (a version with an OLED display).}
+	\label{fig:sensornode}
+\end{figure}
+
 The sensor node itself is built around the low-power Microchip microcontroller (PIC18F46J50) featuring an accelerometer sensor (ADXL345) to capture human motion, light and ambient temperature sensors and a microSD flash card for locally storing the sensor data.
@@
 A plastic case packages and protects the sensor to be worn at the wrist (Figure \ref{fig:sensornode}).
-
-\begin{figure}
-	\centering
-	\includegraphics[width=\textwidth]{Images/hardware_sensor_2.jpg}
-	\caption{The inertial data logger featuring a low-power microcontroller, a 3 axis accelerometer, a microSD flash card for storing the sensor data and a USB connector for accessing the data (on the right) is powered by a small lithium polymer battery and is packaged into a plastic case to be worn at the wrist (a version with an OLED display).}
-	\label{fig:sensornode}
-\end{figure}
 
@@
 
-In the following, the participant were instructed to lie  down on the couch.
+In the following, the participants were instructed to lie  down on the couch.
 Again, there were great variations in how this activity was performed by the participants.
@@
 
-\begin{figure}[t]
+\begin{figure}[!t]
 	\centering
@@
 This activity involved direct interactions with various objects as well as proximity to furniture in the room (see Table \ref{tab:activities}).
-The capacitive proximity sensor was able to capture the proximity to the shelf and to the table.
+The capacitive proximity sensor was able to capture the proximity to the shelf and to the table (see Figure \ref{fig:get_things}).
 The average F-measure for this activity is rather low, but improved by 6.8\% from 53.8\% to 60.6\%.
@@
 This is most likely due to various objects involved in both activities, and the fact that 1-second features are obtained.
-The capacitive sensor modality has a much more positive impact reducing the confusion with other activities.
+The capacitive sensor modality has a more positive impact reducing the confusion with other activities.
 
-\begin{figure}[t]
+\begin{figure}[h]
 	\centering
@@
 An example of an eating activity is shown in Figure \ref{fig:eating}, where the participant took a few bites from the sandwich while putting it down every time.
-The average F-measure for that activity increased slightly by 2.7\% (from 79.5\% to 82.2\%).
+The average F-measure increased slightly by 2.7\% (from 79.5\% to 82.2\%).
 
@@
 	\caption{During the sleeping activity the data from both sensors remains constant for large time spans.
-One can draw conclusions about the coverage of the arm with either cushions, blankets or the proximity to the mattress, the head or body of the participant.}
+The capacitive sensor shows the coverage of the arm with either cushions, blankets or the proximity to the mattress, the head or body of the participant.}
 	\label{fig:sleeping}
@@
 
-\section*{Acknowledgments}
+\section*{Acknowledgements}
 
@@
 \bibliographystyle{splncs03}
-%\bibliography{references}
+\bibliography{references}
 

### 2012-08-24 Paper/Paper.tex
final camera ready version

@@
 
-\begin{figure}[!t]
+\begin{figure}[t]
 	\centering
 		\includegraphics[width=\textwidth]{../Auswertung/images/tobias_4.pdf}
-	\caption{An example of the ``getting things'' activity, where the participants had to get food and dishes from shelves and lockers.
+	\caption{An example of the ``get things'' activity, where the participants had to get food and dishes from shelves and lockers.
 The proximity sensor peaks in the beginning (9s and 19s) indicate immediate proximity to shelf, and to the locker (55-63s) in the kitchen (upper plot).
@@
 	\caption{During the sleeping activity the data from both sensors remains constant for large time spans.
-The capacitive sensor shows the coverage of the arm with either cushions, blankets or the proximity to the mattress, the head or body of the participant.}
+The capacitive sensor plot shows the coverage of the arm with either cushions, blankets or the proximity to the mattress, other parts of the bed, the head or body of the participant.}
 	\label{fig:sleeping}
@@
 In this case, the capacitive proximity sensor is able to capture the surrounding cushions and blankets, as well as the body or head of the participant.
-The accelerometer data and the proximity data have larger periods of low variance, which is a cause for confusion with the ``lying on the couch'' activity.
+The accelerometer data and the capacitive proximity data have larger periods of low variance, which is a cause for confusion with the ``lying on the couch'' activity.
 
@@
 
-We would like to thank our evaluation participants from TU Darmstadt and Fraunhofer IGD.
+We would like to thank our evaluation participants from the Technische Universit\"at Darmstadt and Fraunhofer IGD.
 
